### 2017-04-25 README.Rmd
First 80%

---
title: "Git and GitHub for statistical practice and teaching"
author: "Jenny Bryan"
date: "`r format(Sys.Date())`"
output: github_document
---

*Alternative title: Excuse me, do you have a moment to talk about Git and GitHub?*

## Why and who?

Why would a practicing statistician use version control, such as Git? And what is the point of hosting your work online, e.g., on GitHub? Could the gains possibly justify the inevitable pain?

I say yes, with the zeal of the converted.

The target reader is anyone who does statistical research, analysis, or instruction. Those whose work is some combination of these three may find the work style described here especially rewarding.

Benefits of using hosted version control as your *modus operandi*:

  * Doing your work becomes tightly integrated with organizing, recording, and disseminating it.
  * The marginal effort required to create a web presence for a project is negligible.
  * GitHub makes a fantastic course management system for courses that use R.
  * By using common mechanics across work modes, you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.

This article does not provide step-by-step instructions on how to use Git and GitHub. This format would not be effective, but we do provide links to such resources. Instead, I'll convey what the workflow feels like and what the payoffs are. The goal is to help the Git-curious generate the activation energy needed to get started.

## Git

[Git](http://git-scm.com) is a __version control system__. Its original purpose was to help groups of developers work collaboratively on big software projects. Git manages the evolution of a set of files -- called a __repository__ -- in a sane, highly structured way. It is a bit like the "Track Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up to multiple files.

Git has been re-purposed by the data science community. In addition to using it for source code, we use it to manage the motley collection of files that make up typical data analytical projects, which often consist of data, figures, reports, and, yes, source code. Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.

A lone ranger, working on a single computer, can benefit from adopting version control. But not nearly enough to justify the pain of installation and workflow upheaval. There are much easier ways to get versioned back ups of files, if that's all you're worried about.

In my opinion, **for new users**, the pros of Git only outweigh the cons when you factor in the overhead of communicating and collaborating with other people, including your future self. And who among us does not need to do that? Your life is much easier if this is baked into your workflow, as opposed to being a separate process that you dread or neglect. Communication and collaboration are the killer apps of version control, which brings me to my next point.

## GitHub

[GitHub](https://github.com) is currently the most popular Git hosting service. Others include [GitLab](https://about.gitlab.com) and [Bitbucket](https://bitbucket.org). These services provide a home for Git-based projects on the internet. It is a bit like DropBox or Google Drive, but more structured, powerful, and programmatic.

The remote host acts as a distribution channel or clearinghouse for a Git-managed project. This allows other people to see the project files, sync up with the current version, and perhaps even make or propose changes. These hosting providers offer well-designed web interfaces that are a dramatic improvement over traditional Unix Git servers. Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files.

Even for private solo projects, there are two advantages to keeping a synced copy on GitHub:

  1. When you are new with Git (or, frankly, even when you're not), it's common to goof up the Git infrastructure for a project. Note that your files can absolutely be intact and safe, even while the Git tracking is a bit confused. Of course there are official Git remedies, but sometimes the easiest fix is to clone a fresh copy from GitHub, patch things up with the changes that only exist locally, and move on with your life. This workaround obviously requires the existence of a recent copy on GitHub.
  2. The highly functional web interfaces mentioned above are often the most pleasant and natural way to navigate and search your files, even though all the same information exists locally. It is a pleasure to browse through your own work, across multiple projects or files and across time, as if it's a well-designed website. You must push your work to GitHub to enjoy this.

## Don't get caught up on these things

I target [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity. However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms. I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today. I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.

Don't fret too much about public versus private repositories at this point. All the major hosting providers offer private repositories with flexible control over who can read or write to the repo. There are many ways to get private repositories for low or no cost, especially for academics. Just get started and figure out if and how Git/GitHub is going to work for you. If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem. You can either pay for a higher level of service, self-host one of these platforms, or connect with / advocate for organization-wide solutions. You can also leave more things public than you think, because one's early days with Git and GitHub generally do not attract a keen audience. It is also easy to delete repos or clear the history.

## Is this going to hurt?

Yes.

Git was built neither for the exact usage described here, nor for broad usability. You will undoubtedly notice this, so, no, it's not you. Happily there are many helpful tools that mediate your interactions with Git. GitHub itself is a fine example. In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.

General recommendations for agony reduction:

  * Consider installing a graphical front-end for Git, a.k.a. a "Git client", on your computer, versus restricting yourself to the command line interface.
  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
  * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.

## Initial setup

This is one-time or once-per-computer setup.

  * Register for a free account with GitHub.

  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to do this.

  * Install a local Git client. Optional but highly recommended. A Git client is software that provides a graphical user interface for Git, which is otherwise command-line only. If you are an R user, you will find that RStudio provides a great deal of this functionality. There are some notable gaps, however, so you might still choose to install a dedicated and comprehensive Git client such as SourceTree or GitKraken. Git is a file-based system, so you can do some operations from RStudio, others from SourceTree, and others from the shell.

  * Confirm, with a practice repository, that local Git can send and receive the current version of the repository on GitHub, known as *pushing* and *pulling*, respectively. This will require authenticating yourself with GitHub from a local shell or Git client. At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.

## Repositories and workflow

For new or existing projects, you will:

  * Dedicate a directory (a.k.a "folder") to it.
  * Make it an RStudio Project. Optional but recommended; obviously only applies to projects involving R and users of RStudio.
  * Make it a Git repository.

This is once-per-project setup and can happen at project inception or at any later point. Chances are your project already lives in a dedicated directory. Making this directory an RStudio Project and Git repository boils down to allowing those applications to leave notes for themselves in hidden files or directories. The project is still a regular directory on your computer, that you can locate, name, and interact with as you wish.

Here is the daily workflow:

  * Go about your usual business, e.g. writing R scripts or authoring reports in LaTeX or R Markdown. But instead of only *saving* individual files, periodically you make a **commit**, which takes a multi-file snapshot of the entire project.
    - Have you ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531)? That is effectively a **commit**, albeit only for a single file: it is a version that is significant to you and that you might want to inspect or revert to later.
  * Push commits to GitHub periodically.
    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment. It signals you're ready to make your work visible to others and invite comment or edits.

This is a moderate change to your normal, daily workflow. It feels weird at first, but quickly becomes second nature. FWIW, [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub. This is a major topic in class and office hours for the first two weeks. Then we practically never discuss it again.

## Commits, diffs, and tags

The *commit* is one of the most fundamental concepts in Git, along with *repository*. A commit functions like a snapshot of all the files in the repo, at a specific moment. Under the hood, that is not exactly how Git implements things. Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.

*this is really crying out for an example and/or diagrams*

Consider version A of a file and then version B, which comes from modifying version A. Assume that version A was part of a Git commit and version B was part of the next commit. The set of differences between A and B is called a *diff* and Git users contemplate diffs a lot. Diff inspection is how you re-explain to yourself how version A differs from version B. Diff inspection is not limited to adjacent commits. You can inspect the diffs between any two commits.

In fact, Git's notion of version B of your file is as an accumulation of diffs. At some earlier point, the file was created in the first place. That version of the file was part of a commit and, therefore, a diff. Git stores Version A of the file as the initial version, plus all the intervening diffs in the history that affect the file. And version B simply includes one more. We'll set these internal details aside now, but understanding the importance of these deltas will eventually make Git's operations less baffling.

So, it's easy to see how two snapshots differ, but what about the why?

Every time you make a commit you must also write a short *commit message*. Ideally, this conveys the *motivation* for the change. Remember, the diff will show the content. When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the *history*, by reading commit messages and skimming through diffs, is an extremely efficient way to get back up to speed.

Every commit needs some sort of nickname, so you can identify it. Git does this automatically, assigning each commit what is called a *SHA*, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a checksum hash and is technically a SHA-1). Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice. You can also designate certain snapshots as special with a *tag*, which is a name of your choosing. In a software project, it is typical to tag a release with its version, e.g., "v1.0.3". For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.

Many people who don't use Git are unwittingly creating a poor man's version of it. They create long chains of copies of an important file, decorating the name with initials, dates and other descriptors. The Git way is to track the evolution of that file, through a series of commits, each equipped with an explanatory message and a nickname. Especially important versions are given a human-readable tag, to signal a meaningful milestone. Yes, there is some pain in adopting the formalism of Git, but it is worth it.

## Markdown is special

This may seem unrelated to Git and R, but it's time to talk about Markdown. Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible. The goal is still to separate form and content, but also to preserve human-readability, even at the cost of fancy features. Markdown is in wide use on sites like Wikipedia, WordPress, StackOverflow, and, yes, GitHub. Do not build this up into some heroic, LaTeX-level learning task, for it is not. If you can write an email, you can write Markdown.

Any file written in Markdown is rendered in an HTML-like way on GitHub. In particular, formatting and links "just work". This is the last piece we needed to seal my claim that merely pushing your project to GitHub gives it a web presence for zero extra work. If you make even a modest effort to embed a few explanatory Markdown files in your repo, you will get an automatically-updated project website for free.

Markdown is extra special for R users because of R Markdown, which is just Markdown that includes chunks of R code. Again, do not build this up into something you must clear your schedule to learn. If you can write email and a bit of R code, you can write R Markdown. The rmarkdown package converts R Markdown (`.Rmd` files) to Markdown (`.md` files), running the code and inserting the results, including figures, into the document. This is powered by another package, knitr, under the hood. This process is made very easy in RStudio, but is by no means limited to users of that application.

These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described. The overall effect is that a directory that is an RStudio project and GitHub-synced Git repo can simultaneously be the code-heavy back end of a project and the outward-facing front end.

One last point about R Markdown more generally: the rmarkdown package can render `.Rmd` files to many more formats than Markdown. In addition to knitr, rmarkdown uses Pandoc and, optionally, LaTeX, to convert `.Rmd` to a wide array of output formats, including HTML, PDF, and Word (`.docx`). The Markdown must be created regardless. Sometimes it is all you need, such as on GitHub, and, in other contexts, it is just a necessary intermediate and may even be discarded.

## Which files to commit

The files in a project play different roles and arise in different ways. Let's have a few examples in mind for this discussion:

  * R markdown `.Rmd` --> markdown `.md`
  * R markdown `.Rmd` --> markdown `.md` --> `.html` or `.pdf` or `.docx`
  * R script `.R` --> results as `.csv` or `.rds` and figures `.png`
  * LaTeX `.tex` --> WHAT GOES HERE NOW? --> `.pdf`

The files at the far left are clearly source files. In the case of an R script, this is literally true, but it's morally true for R markdown and LaTeX files too. These are files that you directly create and edit "by hand".

The files at the far right are clearly derived and are often described as "targets". These files are programmatically generated from source files (and possibly other inputs). These files are the product and they have external value, often for communicating ideas and results.

The files in the middle are intermediates. Like targets, they are programmatically generated, but, unlike targets, no one necessarily cares about them. However, note that intermediate Markdown `.md` is an exception, since it is extremely useful on GitHub -- much more so than `.html`, `.pdf`, or `.docx` -- and is more like an additional target.

A critical issue for workflow happiness is figuring out how to manage the production and storage of source, intermediates, and targets with respect to Git. You can direct Git to ignore specific files or certain types of files, such as autosaves created by your editor. This reduces clutter in your project: Git will not pester you to commit changes to these files and they will not appear in the associated GitHub repository. A file that Git does not ignore is said to be *tracked*.

The only point on which there is consensus is that source files should absolutely be tracked. The status of intermediates and targets with respect to Git is much less clear cut.

Therefore, the main message for intermediates and targets is that you can pick a policy that works for you and adapt as your needs change. There is no right answer. I suggest erring on the side of committing everything at first.

What are the main considerations when deciding whether to track a derived file or file type?

*Is it immediately useful to someone?* If so, that is a reason to track it and push it to GitHub. There is a taboo against committing derived products, inherited from Git's software development roots. The reasoning is that compiled programs are platform-specific and, therefore, people are better served by getting current source from Git and compiling themselves. I think data analytic targets, like figures and rendered reports, are very different beasts and it's misguided to reflexively exclude them from version control. They are immediately useful, especially to consumers of a project (versus the makers). To the extent that a GitHub repo is meant for communication, there is no reason to burden every consumer with unnecessary friction. Most simply will not bother to clone the repo, install all the necessary dependencies, and remake the products. Make them readily available.

*Is it available elsewhere?* If so, then perhaps you don't need to track it and push it to GitHub. Many people who have a policy of not tracking derived products also have a system that makes these available elsewhere, such as on a separate website. There are ways to automate this via GitHub, but that is beyond the scope of this article and not recommended for your early days with Git and GitHub. Beware those who recommend the exclusion of derived products without offering any practical solution for making them available some other way.

*Is it huge or changing often? Is it a format that is of little use on GitHub?* These are all good reasons to not track a file with Git. They can make your repository bloated and slow down pushes and pulls. If a file is binary, such as a Word document or Excel spreadsheet, Git and Git clients will not be able to provide human-readable diffs. Neither will GitHub be able to directly display this file in the browser. Be aware, however, that Git/GitHub-friendliness is more complicated than "is it plain text?". GitHub has excellent support for non-code files, such as image formats (PNG, JPG, GIF, PSD, and SVG) and PDFs. And, while HTML is plain text, it is of little direct use on GitHub, because it is not rendered.

*Will diffs be useful to you?* Some derived files are simply too big or miserable to read casually, such as `.csv` files of processed results or `.html` derived from `.Rmd`. But they may still be worth tracking with Git, because the diffs are often modest and quite interesting. I have caught unexpected changes in analytical results and student-facing webpages this way. When you re-run an analysis with updated input data or after updating R packages, the diffs presented by Git help you quickly pinpoint the downstream consequences.

In summary, I recommend you default to including a file in your Git repository, unless there's a specific reason not to. But good reasons absolutely do exist.

## Collaboration

Collaboration is probably the most compelling reason to manage a project with Git and GitHub. I have a broad definition of collaboration here, that includes hands-on participation by multiple people as well as a more asymmetric model, in which some people are active makers and others only read or review.

Consider two different models of collaboration on a document:

  * **Edit, save, attach.** In this workflow, everyone has one (or more!) copies of the document and they circulate via email attachment, accumulating initials and dates in the filename. Which one is "master"? Does this question even make sense anymore? How do different versions of the document relate to each other? If you want to see a version combining the best versions of each section, how would you reconcile the different copies into one?  All of this usually gets sorted out by social contract, a fairly manual process, and at least one miserable person.

  * **Google Doc.** In this workflow, there is only one copy of the document and it lives in the cloud. Anyone can access the most recent version on demand. Anyone can edit or comment or propose a change and this is immediately available to everyone else. Anyone can see who's been editing the document and how and, if disaster strikes, can revert to a previous version. A great deal of ambiguity and annoying reconciliation work has been designed away.

Managing a project via Git/GitHub is much more like the Google Doc scenario and enjoys many of the same advantages. It is definitely more complicated than collaborating on a Google Doc, but also more powerful. How does collaboration work?

Git is a decentralized version control system, meaning each collaborator has their own complete copy of the repo and its history. Everyone can work offline and/or simultaneously. GitHub plays the role of another collaborator, but a very special one. By convention, everyone agrees that it keeps the master copy of the project. GitHub is the clearinghouse. The joke is that GitHub puts the "central" in decentralized version control. You pull regularly from GitHub, to receive and integrate changes made by your collaborators. You also push regularly to GitHub, to return the favor, and to maintain its status as the comprehensive, authoritative version of the project.

*These two points were written as comparisons with Google Docs, but now there are intervening paragraphs. Is it too disconnected?*

*Manage multiple files*. A Git repository is inherently multi-file and therefore well suited to projects comprised of many files, evolving in a coordinated fashion. Examples include a data analysis, a course website, a blog, an R package, or a book. If there is any way to proactively check or enforce their joint functionality, this is something you could verify manually prior to a commit or at certain milestones. In the case of a website, you might choose to rebuild the site prior to a commit. In the case of an R package, `R CMD check` is one of the easiest things to automate on GitHub.

*Diffs and time travel*. Google Docs are fantastic for simple collaborative work, when you don't need detailed access to the history. But the version control offered by Google Drive is very limited compared to Git. You can't compare versions at arbitrary points in time, temporarily checkout previous versions, or maintain two lines of development.

*Needs a segue, which depends on what happens above.*

Merge conflicts are the most frustrating thing about using Git and GitHub. You can avoid them if you only work alone, on one computer, but I've also said that collaboration is the best reason to use GitHub! So this problem must be confronted.

What is a merge conflict? Your first encounter will probably look like this: your collaborator makes a change to a file, commits it locally, and pushes to GitHub. Meanwhile you also make a different change to the same file and also commit locally. When you try to push your commit to GitHub, it will fail because there are commits on GitHub that you do not have. You must pull from GitHub. The good news is that quite often, this will "just work", i.e. the GitHub version and your version will merge cleanly. Git is quite clever at reconciliation and changes to different files or distinct parts of the same file will merge.

But sometimes it's not clear how to reconcile your changes with the new ones from GitHub and you get a *merge conflict*. You must inspect the locations of conflict, where the two competing versions will have been left for you, and reconcile them. Then, and only then, will you be able to move forward and push a version integrating your recent changes to GitHub.

The best way to deal with merge conflicts is to avoid them altogether. This is another reason for all parties to commit, pull, and push often. Small changes, being integrated frequently, are easy for Git to merge, usually without conflict. The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work. So keep it small with lots of small commits and regular syncing with GitHub.

## GitHub as course management system

*adapt from here: <http://happygitwithr.com/classroom-overview.html>*

*Want to convey utility for managing the course materials/website, discussion board via issues, student repositories for coursework.*

## R specific stuff

*Much of this appears elsewhere in some form. Perhaps I should find a home for what's not covered and kill this section?*

  * The active R package development community on GitHub. The read-only mirrors of R source and all of CRAN. Coupled with GitHub search features, you can answer alot of your own questions this way.
  * Specific workflows for R and Rmd destined to GitHub. `github_document`!!! `keep_md = TRUE`!!!
  * Git- and GitHub-related features of the [RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).

## Special features of GitHub

*at this point, just fodder mined from various locations in Happy Git*

In addition to a well-designed user interface, GitHub offers two especially important features:

  * **Issues.** Remember how we're high-jacking software development tools? Well, this is the bug tracker. It's a list of things ... bugs, feature requests, to dos, whatever.
    - Issues are tightly integrated with email and therefore allow you to copy/embed important conversations in the associated repo.
    - Issues can be assigned to people (e.g., to dos) and tagged ("bug" or "progress-report").
    - Issues are tightly integrated with commits and therefore allow you to record *that the changes in this commit solve that problem which was discussed in that issue*.
    - As a new user of GitHub, one of the most productive things you can do is to use GitHub issues to provide a clear bug report or feature request for a package you use.
 * **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch. These are technical Git terms but hopefully also make sense on their own. A pull request is a formal proposal that says: "Here are some changes I would like to make." It might be linked to a specific issue: "Related to #14." or "Fixes #56". GitHub facilitates and preserves the discussion of the proposal, holistically and line-by-line.

**Be a keener!** If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub. You can watch the repository to get notified of major activity. You can fork it to keep your own copy. You can modify your fork to add features or fix bugs and send them back to the owner as a proposed change.

### GitHub as web presence

Simply having a project on GitHub gives it a web presence.

If someone needs to see your work or if you want them to try out your code, they can easily get it from GitHub. If they use Git, they can clone or fork your repository. If they don't use Git, they can still browse your project on GitHub like a normal website and even grab everything by downloading a zip archive.

There are many things you can do a "working" repository to make it double as a project website.

  * [Make a GitHub repo browsable](http://happygitwithr.com/repo-browsability.html)

If you want a more proper website, GitHub, rmarkdown, and, especially, the combination provide several options:

  * https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3
  * https://help.github.com/categories/github-pages-basics/
  * http://rmarkdown.rstudio.com/rmarkdown_websites.html
  * https://bookdown.org/yihui/bookdown/
  * https://bookdown.org/yihui/blogdown/
  * https://hadley.github.io/pkgdown/

## Random things lying around

need to add a ton of links, citations

make some bespoke examples or diagrams?

Doing = documenting = sharing

<https://help.github.com/categories/working-with-non-code-files/>
<https://help.github.com/articles/rendering-and-diffing-images/>

make statistical thought and implementation available

Transparency about process and product is increasingly important in science. The SOMETHING for reproducibility is well accepted. A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.

Who can do what?

A public repository is readable by the world. The owner can grant higher levels of permission to others, such as the ability to push commits.

A private repository is invisible to the world. The owner can grant read, write (push), or admin access to others.

There is also a formal notion of an organization, which can be useful for managing repository permissions for entire teams of people.

### 2017-05-13 README.Rmd
Newline and 2FA test
Respond to @nicholasjhorton comment 1; closes #4

@@
 ---
-title: "Git and GitHub for statistical practice and teaching"
+title: "Excuse me, do you have a moment to talk about version control?"
 author: "Jenny Bryan"
@@
 
-*Alternative title: Excuse me, do you have a moment to talk about Git and GitHub?*
-
-## Why and who?
+## Why Git?
 
@@
 
-The target reader is anyone who does statistical research, analysis, or instruction. Those whose work is some combination of these three may find the work style described here especially rewarding.
+Benefits of using hosted version control in your statistical practice:
 
-Benefits of using hosted version control as your *modus operandi*:
+  * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
+  * The marginal effort required to create a web presence for a project is negligible. It's not a separate, burdensome task you are tempted to neglect.
+  * GitHub makes a fantastic course management system for courses that use R. You and your students can exchange actual working code and explore the associated results.
+  * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.
 
-  * Doing your work becomes tightly integrated with organizing, recording, and disseminating it.
-  * The marginal effort required to create a web presence for a project is negligible.
-  * GitHub makes a fantastic course management system for courses that use R.
-  * By using common mechanics across work modes, you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.
-
-This article does not provide step-by-step instructions on how to use Git and GitHub.
-This format would not be effective, but we do provide links to such resources.
-Instead, I'll convey what the workflow feels like and what the payoffs are.
-The goal is to help the Git-curious generate the activation energy needed to get started.
-
-## Git
+## What is Git?
 
@@
 And who among us does not need to do that?
-Your life is much easier if this is baked into your workflow, as opposed to being a separate process that you dread or neglect.
-Communication and collaboration are the killer apps of version control, which brings me to my next point.
+Your life is much easier if this is baked into your workflow, as opposed to being a separate process that you dread and avoid.
+Communication and collaboration are the killer apps of version control.
+Git enforces a painfully rigorous model of file management, but it is critical for the distribution of files across different people, computers, and time.
+
+Many people who don't use Git unwittingly re-invent a poor man's version of it.
+They take an important file and distribute it via email.
+All parties make changes, decorating the file name with initials, dates, and other descriptors.
+Before you know it, the original file is the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve.
+
+The Git way is to track the evolution of that file, through a series of commits, each equipped with an explanatory message and a nickname.
+All collaborators sync regularly to a common version, acknowledging that the difficulty of merging goes up faster than the size of the difference (cite good enough).
+Especially important versions get a human-readable tag, to signal a meaningful milestone.
+Yes, there is some pain in adopting the formalism of Git, but it is worth it.
+
+## Who should read this and what to expect
+
+The target reader is anyone who does statistical research, analysis, or instruction. Those whose work is some combination of these three may find the work style described here especially rewarding.
+
+This article does not provide step-by-step instructions on how to use Git and GitHub.
+This format would not be effective, but we do provide links to such resources.
+Instead, I'll convey what the workflow feels like and what the payoffs are.
+The goal is to help the Git-curious generate the activation energy needed to get started.
 
@@
 For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.
-
-Many people who don't use Git are unwittingly creating a poor man's version of it.
-They create long chains of copies of an important file, decorating the name with initials, dates and other descriptors.
-The Git way is to track the evolution of that file, through a series of commits, each equipped with an explanatory message and a nickname.
-Especially important versions are given a human-readable tag, to signal a meaningful milestone.
-Yes, there is some pain in adopting the formalism of Git, but it is worth it.
 
@@
 
+Does creation and presence in GitHub repo make the project more readable and navigable? If so DO IT.
+
 need to add a ton of links, citations
@@
 <https://help.github.com/categories/working-with-non-code-files/>
+
 <https://help.github.com/articles/rendering-and-diffing-images/>
@@
 There is also a formal notion of an organization, which can be useful for managing repository permissions for entire teams of people.
+
+why sharing intermediate md is better than html (not readable) or pdf (merge conflicts).
+markdown is the magical intersection of git-friendly (plain text) and human-friendly (becasue GitHub renders it nicely)
+
+possible github explainer
+https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/
+
+http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst
+
+my tips around md are the large concept of "make things navigable/readable" of GitHub, particularized to R code
+
+https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem

### 2017-05-13 README.Rmd
Move GitHub issues up; closes #5

@@
 Communication and collaboration are the killer apps of version control.
-Git enforces a painfully rigorous model of file management, but it is critical for the distribution of files across different people, computers, and time.
+Git enforces a rigorous model of file management, but it is critical for the distribution of files across different people, computers, and time.
 
@@
 They take an important file and distribute it via email.
-All parties make changes, decorating the file name with initials, dates, and other descriptors.
+Various parties make changes, decorating the file name with initials, dates, and other descriptors.
 Before you know it, the original file is the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve.
@@
 This format would not be effective, but we do provide links to such resources.
-Instead, I'll convey what the workflow feels like and what the payoffs are.
+Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statsitics and R context.
 The goal is to help the Git-curious generate the activation energy needed to get started.
 
-## GitHub
+## What is GitHub?
 
@@
 You must push your work to GitHub to enjoy this.
+
+GitHub __issues__ are another powerful feature of the platform.
+Recall that we are repurposing Git, a tool designed to facilitate software development.
+The issues for a project are its bug tracker.
+For projects that are not pure software development, we use this machinery to organize our to do list more generally.
+The basic unit is an issue and you can interact with them in two ways.
+First, issues are integrated into the project's web interface on GitHub, with a rich set of options for linking to project files and incremental changes.
+Second, issues and their associated comment threads appear in your email, just like regular messages (this can, of course, be configured).
+The result is that all correspondence about a project comes through your normal channels, but is also tracked inside the project itself, with excellent navigability and search capabilities.
+
+Issues can be assigned to specific people and they can be labelled, e.g.
+"bug", "simulation-study", or "final-exam".
+Coupled with the ability to cross-link issues and the project files or file changes, you have extraordinary power to document why things have happened in the past and to organize what needs to happen in the future.
+
+## Initial Git setup
+
+This is one-time or once-per-computer setup.
+
+  * Register for a free account with GitHub.
+
+  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to do this.
+
+  * Install a local Git client.
+Optional but highly recommended.
+A Git client is software that provides a graphical user interface for Git, which is otherwise command-line only.
+If you are an R user, you will find that RStudio provides a great deal of this functionality.
+There are some notable gaps, however, so you might still choose to install a dedicated and comprehensive Git client such as SourceTree or GitKraken.
+Git is a file-based system, so you can do some operations from RStudio, others from SourceTree, and others from the shell.
+
+  * Confirm, with a practice repository, that local Git can send and receive the current version of the repository on GitHub, known as *pushing* and *pulling*, respectively.
+This will require authenticating yourself with GitHub from a local shell or Git client.
+At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.
 
@@
   * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.
-
-## Initial setup
-
-This is one-time or once-per-computer setup.
-
-  * Register for a free account with GitHub.
-
-  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to do this.
-
-  * Install a local Git client.
-Optional but highly recommended.
-A Git client is software that provides a graphical user interface for Git, which is otherwise command-line only.
-If you are an R user, you will find that RStudio provides a great deal of this functionality.
-There are some notable gaps, however, so you might still choose to install a dedicated and comprehensive Git client such as SourceTree or GitKraken.
-Git is a file-based system, so you can do some operations from RStudio, others from SourceTree, and others from the shell.
-
-  * Confirm, with a practice repository, that local Git can send and receive the current version of the repository on GitHub, known as *pushing* and *pulling*, respectively.
-This will require authenticating yourself with GitHub from a local shell or Git client.
-At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.
 
@@
 
-In addition to a well-designed user interface, GitHub offers two especially important features:
-
-  * **Issues.** Remember how we're high-jacking software development tools? Well, this is the bug tracker. It's a list of things ... bugs, feature requests, to dos, whatever.
-    - Issues are tightly integrated with email and therefore allow you to copy/embed important conversations in the associated repo.
-    - Issues can be assigned to people (e.g., to dos) and tagged ("bug" or "progress-report").
-    - Issues are tightly integrated with commits and therefore allow you to record *that the changes in this commit solve that problem which was discussed in that issue*.
-    - As a new user of GitHub, one of the most productive things you can do is to use GitHub issues to provide a clear bug report or feature request for a package you use.
  * **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch.

### 2017-05-13 README.Rmd
More on issues
Kill 'special features of github' section

@@
 The basic unit is an issue and you can interact with them in two ways.
+
 First, issues are integrated into the project's web interface on GitHub, with a rich set of options for linking to project files and incremental changes.
@@
 The result is that all correspondence about a project comes through your normal channels, but is also tracked inside the project itself, with excellent navigability and search capabilities.
+For software, issues are used to track bugs and feature requests.
+In a data analysis project, you might open an issue for a specific sub-analysis or complicated figure that needs to be made.
+In a course, we use them to manage homework submission, marking, and peer review.
 
@@
 
-## Don't get caught up on these things
+## Things you can tackle later
 
@@
 It is also easy to delete repos or clear the history.
+
+*text pulled from elsewhere but might fit here?* **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch.
+These are technical Git terms but hopefully also make sense on their own.
+A pull request is a formal proposal that says: "Here are some changes I would like to make." It might be linked to a specific issue: "Related to #14." or "Fixes #56".
+GitHub facilitates and preserves the discussion of the proposal, holistically and line-by-line.
 
@@
 
-## Special features of GitHub
-
-*at this point, just fodder mined from various locations in Happy Git*
-
- * **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch.
-These are technical Git terms but hopefully also make sense on their own.
-A pull request is a formal proposal that says: "Here are some changes I would like to make." It might be linked to a specific issue: "Related to #14." or "Fixes #56".
-GitHub facilitates and preserves the discussion of the proposal, holistically and line-by-line.
-
   **Be a keener!** If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub.

### 2017-05-13 README.Rmd
Move 'things you can tackle later'
Mention basic, mission critical config; closes #2

@@
 The issues for a project are its bug tracker.
-For projects that are not pure software development, we use this machinery to organize our to do list more generally.
+For projects that are not pure software development, we coopt this machinery to organize our to do list more generally.
 The basic unit is an issue and you can interact with them in two ways.
@@
 
-## Initial Git setup
+## Initial system setup
 
@@
 
-  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to do this.
+  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to do this. Some basic configuration is also critical, such as setting your username and email.
 
@@
 At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.
-
-## Things you can tackle later
-
-I target [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity.
-However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms.
-I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today.
-I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.
-
-Don't fret too much about public versus private repositories at this point.
-All the major hosting providers offer private repositories with flexible control over who can read or write to the repo.
-There are many ways to get private repositories for low or no cost, especially for academics.
-Just get started and figure out if and how Git/GitHub is going to work for you.
-If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem.
-You can either pay for a higher level of service, self-host one of these platforms, or connect with / advocate for organization-wide solutions.
-You can also leave more things public than you think, because one's early days with Git and GitHub generally do not attract a keen audience.
-It is also easy to delete repos or clear the history.
-
-*text pulled from elsewhere but might fit here?* **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch.
-These are technical Git terms but hopefully also make sense on their own.
-A pull request is a formal proposal that says: "Here are some changes I would like to make." It might be linked to a specific issue: "Related to #14." or "Fixes #56".
-GitHub facilitates and preserves the discussion of the proposal, holistically and line-by-line.
 
@@
 
+## Things you can tackle later
+
+I target [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity.
+However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms.
+I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today.
+I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.
+
+Don't fret too much about public versus private repositories at this point.
+All the major hosting providers offer private repositories with flexible control over who can read or write to the repo.
+There are many ways to get private repositories for low or no cost, especially for academics.
+Just get started and figure out if and how Git/GitHub is going to work for you.
+If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem.
+You can either pay for a higher level of service, self-host one of these platforms, or connect with / advocate for organization-wide solutions.
+You can also leave more things public than you think, because one's early days with Git and GitHub generally do not attract a keen audience.
+It is also easy to delete repos or clear the history.
+
+*text pulled from elsewhere but might fit here?* **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch.
+These are technical Git terms but hopefully also make sense on their own.
+A pull request is a formal proposal that says: "Here are some changes I would like to make." It might be linked to a specific issue: "Related to #14." or "Fixes #56".
+GitHub facilitates and preserves the discussion of the proposal, holistically and line-by-line.
+
 ## Random things lying around

### 2017-05-17 README.Rmd
Monster commit :japanese_ogre:

--- 2017-05-13 README.Rmd
@@
 date: "`r format(Sys.Date())`"
-output: github_document
+output:
+  github_document:
+    toc: true
 ---
@@
   * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
-  * The marginal effort required to create a web presence for a project is negligible. It's not a separate, burdensome task you are tempted to neglect.
+  * The marginal effort required to create a web presence for a project is negligible.
   * GitHub makes a fantastic course management system for courses that use R. You and your students can exchange actual working code and explore the associated results.
@@
 
+This has an implication for selecting your first Git projects: in order to see substantial gain for your pain, you need to pick projects that require you to share rapidly evolving files with others.
+It is tempting to pick a quiet little project that you work on privately, but you risk missing the main point of formal version control.
+
 Many people who don't use Git unwittingly re-invent a poor man's version of it.
@@
 Before you know it, the original file is the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve.
+
+*here's one way I see phylogeny in my mind*
+![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)
+
+*make such a phylogeny with a set of crazy filenames like this*
+![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)
+
+*or riff on the famous PhD Comic*
+![](http://phdcomics.com/comics/archive/phd101212s.gif)
 
@@
 
+*another good image to emulate, from the Bartlett Git for Designers talk*
+![](bartlett-copies-vs-commits.png)
+
 ## Who should read this and what to expect
@@
 This article does not provide step-by-step instructions on how to use Git and GitHub.
-This format would not be effective, but we do provide links to such resources.
-Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statsitics and R context.
+This format would not be effective, but we do provide annotated links to such resources in ?the appendix or online supplement?.
+Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context.
 The goal is to help the Git-curious generate the activation energy needed to get started.
@@
 
-The remote host acts as a distribution channel or clearinghouse for a Git-managed project.
+The remote host acts as the clearinghouse for a Git-managed project.
 This allows other people to see the project files, sync up with the current version, and perhaps even make or propose changes.
@@
 Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files.
+These hosts also offer granular control over who can see, edit, and administer a project.
 
@@
 The issues for a project are its bug tracker.
-For projects that are not pure software development, we use this machinery to organize our to do list more generally.
+For projects that are not pure software development, we co-opt this machinery to organize our to do list more generally.
 The basic unit is an issue and you can interact with them in two ways.
@@
 For software, issues are used to track bugs and feature requests.
-In a data analysis project, you might open an issue for a specific sub-analysis or complicated figure that needs to be made.
+In a data analysis project, you might open an issue to flesh out a specific sub-analysis or to develop a complicated figure.
 In a course, we use them to manage homework submission, marking, and peer review.
@@
 
-  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to do this.
+  * Install Git.
+Depending on your OS, Git might already be installed.
+But many of us will need to install it or might choose to update to the most recent version.
+Some basic configuration is critical, such as setting your username and email.
 
@@
 
-## Things you can tackle later
-
-I target [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity.
-However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms.
-I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today.
-I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.
-
-Don't fret too much about public versus private repositories at this point.
-All the major hosting providers offer private repositories with flexible control over who can read or write to the repo.
-There are many ways to get private repositories for low or no cost, especially for academics.
-Just get started and figure out if and how Git/GitHub is going to work for you.
-If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem.
-You can either pay for a higher level of service, self-host one of these platforms, or connect with / advocate for organization-wide solutions.
-You can also leave more things public than you think, because one's early days with Git and GitHub generally do not attract a keen audience.
-It is also easy to delete repos or clear the history.
-
-*text pulled from elsewhere but might fit here?* **Pull requests.** Git allows a project to have multiple, independent branches of development, with the notion that some should eventually be merged back into the main development branch.
-These are technical Git terms but hopefully also make sense on their own.
-A pull request is a formal proposal that says: "Here are some changes I would like to make." It might be linked to a specific issue: "Related to #14." or "Fixes #56".
-GitHub facilitates and preserves the discussion of the proposal, holistically and line-by-line.
+Once this setup is done, you are ready to start using Git and GitHub with your projects.
 
@@
 Git was built neither for the exact usage described here, nor for broad usability.
-You will undoubtedly notice this, so, no, it's not you.
+You will undoubtedly notice this, so it can be helpful to know this in advance.
 Happily there are many helpful tools that mediate your interactions with Git.
@@
 
-  * Consider installing a graphical front-end for Git, a.k.a. a "Git client", on your computer, versus restricting yourself to the command line interface.
+  * I repeat: consider using a graphical front-end for Git, a.k.a. a "Git client", versus restricting yourself to the command line interface.
   * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
@@
 
-  * Dedicate a directory (a.k.a "folder") to it.
+  * Dedicate a local directory (a.k.a "folder") to it.
   * Make it an RStudio Project. Optional but recommended; obviously only applies to projects involving R and users of RStudio.
@@
 Making this directory an RStudio Project and Git repository boils down to allowing those applications to leave notes for themselves in hidden files or directories.
-The project is still a regular directory on your computer, that you can locate, name, and interact with as you wish.
+The project is still a regular directory on your computer, that you can locate, name, move, and generally interact with as you wish.
+You don't have to handle it with special gloves!
 
@@
 writing R scripts or authoring reports in LaTeX or R Markdown.
-But instead of only *saving* individual files, periodically you make a **commit**, which takes a multi-file snapshot of the entire project.
+But instead of only *saving* individual files, periodically you make a **commit**, which takes a snapshot of all the files in the entire project.
     - Have you ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531)?
@@
   * Push commits to GitHub periodically.
-    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment. It signals you're ready to make your work visible to others and invite comment or edits.
+    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment.
+By pushing to GitHub, you make your work and all your accumulated progress accessible to others.
 
@@
 It feels weird at first, but quickly becomes second nature.
-FWIW, [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub.
-This is a major topic in class and office hours for the first two weeks.
+In [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub, starting in week one.
+Most have never seen Git before and do not identify as "programmers".
+It is a major topic in class and office hours for the first two weeks.
 Then we practically never discuss it again.
+
+*I am not saying anything about pulling and collaboration here, which seems a bit extreme, even if I can't say much*
 
@@
 
-The *commit* is one of the most fundamental concepts in Git, along with *repository*.
+We now explore the fundamental concepts of Git and connect them to the data science workflow:
+
+  * repository
+  * commit
+  * diff
+
+Recall that a repository -- or "repo" -- is just a directory of files that Git is going to manage holistically.
 A commit functions like a snapshot of all the files in the repo, at a specific moment.
@@
 
-*this is really crying out for an example and/or diagrams*
+*this is really crying out for an example and/or diagrams, something that shows commits unfolding and let's you illustrate a diff, sort of like this from the barlett talk*
 
-Consider version A of a file and then version B, which comes from modifying version A.
+![](bartlett-commit-history.png)
+
+Consider version A of a file and a modified version, version B.
 Assume that version A was part of a Git commit and version B was part of the next commit.
-The set of differences between A and B is called a *diff* and Git users contemplate diffs a lot.
+The set of differences between A and B is called a "diff" and Git users contemplate diffs a lot.
 Diff inspection is how you re-explain to yourself how version A differs from version B.
@@
 
-So, it's easy to see how two snapshots differ, but what about the why?
+So, by looking at diffs, it's easy to see how two snapshots differ, but what about the why?
 
@@
 Remember, the diff will show the content.
-When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the *history*, by reading commit messages and skimming through diffs, is an extremely efficient way to get back up to speed.
+When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the *history*, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.
 
 Every commit needs some sort of nickname, so you can identify it.
-Git does this automatically, assigning each commit what is called a *SHA*, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a checksum hash and is technically a SHA-1).
+Git does this automatically, assigning each commit what is called a *SHA*, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a SHA-1 checksum hash of the commit).
 Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice.
@@
 
-## Markdown is special
+## Markdown is special on GitHub
 
-This may seem unrelated to Git and R, but it's time to talk about Markdown.
+This may seem unrelated to Git, GitHub, and R, but we need to talk about Markdown.
 Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible.
-The goal is still to separate form and content, but also to preserve human-readability, even at the cost of fancy features.
+The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features.
 Markdown is in wide use on sites like Wikipedia, WordPress, StackOverflow, and, yes, GitHub.
+These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting.
 Do not build this up into some heroic, LaTeX-level learning task, for it is not.
@@
 In particular, formatting and links "just work".
-This is the last piece we needed to seal my claim that merely pushing your project to GitHub gives it a web presence for zero extra work.
+This is the last piece we need to seal my claim that merely pushing your project to GitHub gives it a web presence for zero extra work.
 If you make even a modest effort to embed a few explanatory Markdown files in your repo, you will get an automatically-updated project website for free.
+In particular, if a directory has a `README.md` file, GitHub renders it like a home page or "index.html" when people visit that directory in the browser.
+It is very common for a repo to have a top-level `README.md`, but each subdirectory can have its own as well.
+
+## Markdown is special for R users
 
 Markdown is extra special for R users because of R Markdown, which is just Markdown that includes chunks of R code.
-Again, do not build this up into something you must clear your schedule to learn.
+Again, do not regard R Markdown as   something you must clear your schedule to learn.
 If you can write email and a bit of R code, you can write R Markdown.
@@
 This is powered by another package, knitr, under the hood.
-This process is made very easy in RStudio, but is by no means limited to users of that application.
+This process is made especially easy in RStudio, but is by no means limited to users of that application.
 
-These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described.
-The overall effect is that a directory that is an RStudio project and GitHub-synced Git repo can simultaneously be the code-heavy back end of a project and the outward-facing front end.
+*this too is crying out for a visual Rmd -> md -> rendered thing looking good*
 
-One last point about R Markdown more generally: the rmarkdown package can render `.Rmd` files to many more formats than Markdown.
-In addition to knitr, rmarkdown uses Pandoc and, optionally, LaTeX, to convert `.Rmd` to a wide array of output formats, including HTML, PDF, and Word (`.docx`).
+These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described above.
+Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, **and view the results of running that code**, including figures.
+It is the best of all worlds, because the code is revealed and, by definition, is the code that produced the results.
+And yet a reader can gaze upon the product in a web browser, without needing to download the code, install all necessary dependencies, and run it.
+
+The overall effect is that a directory that is a GitHub-synced Git repo can simultaneously be the code-heavy back end of a project and an outward-facing front end.
+
+You do not, in fact, even need to work in R Markdown to exploit this.
+It works with plain R scripts as well.
+You can use exactly the same machinery to prepare a rendered version of an R script, i.e.
+to go from `.R` to `.md`.
+Again, RStudio makes this especially easy, but this is not limited to RStudio.
+Once the Markdown file is pushed to GitHub, it is as if the reader has run your code or is able to look over your shoulder at your R session.
+This provides an lightweight system for exposing work-in-progress to collaborators, without slowing down to create separate reports.
+Comment lines that begin with `#'` are elevated to top-level prose, providing a way to make the document more welcoming for a reader.
+Once there are many prose comments, you might decide to switch from `.R` to `.Rmd`, have proper top-level prose, and move the code down into chunks.
+
+*figure showing the yaml needed for this? so the use of `github_document` or `keep_md = TRUE`; try to make a figure that does double or triple duty, i.e. illustrates this and points made elsewhere*
+
+Let's zoom back out again and consider R Markdown and the rmarkdown package more generally.
+I want to point out that R Markdown can be rendered to many more formats than Markdown!
+I have emphasized the production of Markdown here because it is extremely useful in GitHub-hosted projects and that seems to be underappreciated.
+But the rmarkdown package can convert `.Rmd` to a wide array of output formats, including HTML, PDF, and Word (`.docx`).
 The Markdown must be created regardless.
-Sometimes it is all you need, such as on GitHub, and, in other contexts, it is just a necessary intermediate and may even be discarded.
+Sometimes it is all you need, such as on GitHub.
+In other contexts, it is just a necessary intermediate and may even be discarded.
 
@@
 
-The only point on which there is consensus is that source files should absolutely be tracked. The status of intermediates and targets with respect to Git is much less clear cut.
+The only point on which there is consensus is that source files should absolutely be tracked. The best treatment of intermediates and targets with respect to Git is much less clear cut.
 
@@
 They are immediately useful, especially to consumers of a project (versus the makers).
-To the extent that a GitHub repo is meant for communication, there is no reason to burden every consumer with unnecessary friction.
+To the extent that a GitHub repo is meant for dissemination, there is no reason to burden every consumer with unnecessary friction.
 Most simply will not bother to clone the repo, install all the necessary dependencies, and remake the products.
@@
 Neither will GitHub be able to directly display this file in the browser.
-Be aware, however, that Git/GitHub-friendliness is more complicated than "is it plain text?".
-GitHub has excellent support for non-code files, such as image formats (PNG, JPG, GIF, PSD, and SVG) and PDFs.
-And, while HTML is plain text, it is of little direct use on GitHub, because it is not rendered.
+Be aware, however, GitHub-friendliness does not just boil down to "is it plain text?".
+GitHub has excellent support for non-code files, such as image formats (PNG, JPG, GIF, and SVG) and PDFs.
+It even provides visual diffs, which are extremely useful for understanding changes in figures.
+Finally, even though HTML is plain text, it is of little direct use on GitHub, because, unlike Markdown, it is not rendered.
 
@@
 
+*Will it make collaboration harder?* *talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>*
+
 In summary, I recommend you default to including a file in your Git repository, unless there's a specific reason not to. But good reasons absolutely do exist.
@@
 Collaboration is probably the most compelling reason to manage a project with Git and GitHub.
-I have a broad definition of collaboration here, that includes hands-on participation by multiple people as well as a more asymmetric model, in which some people are active makers and others only read or review.
+I have a broad definition of collaboration here, that includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
 
@@
 
-Managing a project via Git/GitHub is much more like the Google Doc scenario and enjoys many of the same advantages.
+Managing a project via Git/GitHub is much more like the Google Doc scenario, but also offers some of the attractive features of "edit, save, attach".
+With Git/GitHub, collaborators can work offline and there can be independent lines of development.
+The real power comes from regular and structured reconciliation of all the files in the project via Git/GitHub.
 It is definitely more complicated than collaborating on a Google Doc, but also more powerful.
+
 How does collaboration work?
@@
 Git is a decentralized version control system, meaning each collaborator has their own complete copy of the repo and its history.
-Everyone can work offline and/or simultaneously.
+Everyone can work offline and/or simultaneously, but with regular syncing to GitHub.
 GitHub plays the role of another collaborator, but a very special one.
-By convention, everyone agrees that it keeps the master copy of the project.
+By convention, everyone agrees that GitHub keeps the master copy of the project.
 GitHub is the clearinghouse.
@@
 
-*These two points were written as comparisons with Google Docs, but now there are intervening paragraphs. Is it too disconnected?*
+*These two points were originally written as comparisons with Google Docs, but now there are intervening paragraphs. Is it too disconnected? Is it worth weaving in?*
 
@@
 
-*Needs a segue, which depends on what happens above.*
+*May need a segue ... depends on the fate of the two paragraphs above.*
 
@@
 What is a merge conflict?
-Your first encounter will probably look like this: your collaborator makes a change to a file, commits it locally, and pushes to GitHub.
+Here is a typical first encounter: your collaborator makes a change to a file, commits it locally, and pushes to GitHub.
 Meanwhile you also make a different change to the same file and also commit locally.
@@
 the GitHub version and your version will merge cleanly.
-Git is quite clever at reconciliation and changes to different files or distinct parts of the same file will merge.
+Git is quite clever at reconciliation and changes to different files or even distinct parts of the same file will merge.
+If this pull and merge goes smoothly, you'll able to push your changes and the cycle goes on.
 
 But sometimes it's not clear how to reconcile your changes with the new ones from GitHub and you get a *merge conflict*.
-You must inspect the locations of conflict, where the two competing versions will have been left for you, and reconcile them.
-Then, and only then, will you be able to move forward and push a version integrating your recent changes to GitHub.
+You must inspect the locations of conflict, which Git marks for you.
+You will pick one version or the other -- or create a hybrid -- at each location of conflict and mark it as resolved.
+Once you've resolved all conflicts, you will be able to push a version integrating your recent changes to GitHub.
 
@@
 This is another reason for all parties to commit, pull, and push often.
-Small changes, being integrated frequently, are easy for Git to merge, usually without conflict.
+Small changes, being integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you.
 The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work.
-So keep it small with lots of small commits and regular syncing with GitHub.
+The presence of frequently-changing binary files also increases the burden.
+So make lots of small commits, sync regularly with GitHub, and only track binary files with good reason.
+
+*once I understand Nick's point about binary files causing merge conflicts <https://github.com/dsscollection/git-github-for-stats/issues/6> I may need to reword the above*
+
+*something I would like to write about, but have no space for here: different models of collaboration, i.e.
+everyone working on one repo & one branch vs.
+making pull requests from branches in the main repo or from forks*
+
+*inspiration for possible figures on collaboration/merging and the role of a remote, from the barlett talk*
+
+![](bartlett-merge-commit.png)
+
+![](bartlett-pull.png)
 
@@
 
-*adapt from here: <http://happygitwithr.com/classroom-overview.html>*
+[STAT 545](http://stat545.com) is a data wrangling and analysis course at the University of British Columbia.
+I was the instructor in charge for several years, which coincided with my own adoption of Git/GitHub.
+GitHub is used to manage the development of course material, to serve the course website, to create a discussion forum, and to host all student-submitted work.
 
-*Want to convey utility for managing the course materials/website, discussion board via issues, student repositories for coursework.*
+Given that students must submit their work and provide peer review of others' work via GitHub, the use of hosted version control is an explicit, though modest, part of the course.
+The website [Happy Git and GitHub for the useR](http://happygitwithr.com) holds our battle-tested instructions for setup and early usage.
+The students achieve basic competence quite quickly and find it gratifying to see their formatted, figure-rich R Markdown reports up on the internet.
+Since it’s easy to expose their work within the class, we conduct peer review, which helps expertise to spread quickly through the group.
 
-## R specific stuff
+### Use a GitHub Organization
 
-*Much of this appears elsewhere in some form. Perhaps I should find a home for what's not covered and kill this section?*
+[GitHub Organizations](https://help.github.com/articles/differences-between-user-and-organization-accounts/) are "shared accounts where groups of people can collaborate across many projects at once".
+This is the most appropriate structure for stewarding course resources, since I can grant TAs and students different levels of access to various repositories.
+Access can be controlled at the individual user level or, more conveniently, for entire [Teams](https://help.github.com/articles/setting-up-teams/).
+The TA Team shares write access with me on a private repository for internal matters.
+I provide each student with their own private repository for coursework and grant other members of the Students Team read access, in order to facilitate peer review.
+There is a public repository that underpins the course website (see below).
+We have one other public repository that exists solely so the [Issues](https://github.com/STAT545-UBC/Discussion/issues) can be used as a discussion forum.
 
-  * The active R package development community on GitHub. The read-only mirrors of R source and all of CRAN. Coupled with GitHub search features, you can answer alot of your own questions this way.
-  * Specific workflows for R and Rmd destined to GitHub. `github_document`!!! `keep_md = TRUE`!!!
-  * Git- and GitHub-related features of the [RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).
+GitHub actively encourages the use of its platform in teaching.
+As an instructor you can request a [free Organization account](https://help.github.com/articles/discounted-organization-accounts/) that provides features normally available only on paid plans, such as private repositories.
+In fact, GitHub provides tooling for specific teaching workflows via [GitHub Classroom](https://classroom.github.com/), although I do not use it.
+That is not an intentional knock on their tools.
+I started teaching with GitHub several years before this existed and developed a different way of using the platform.
+I also find the [GitHub Education](https://education.github.com) resources to be geared more towards computer science than data science.
 
-  **Be a keener!** If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub.
-You can watch the repository to get notified of major activity.
-You can fork it to keep your own copy.
-You can modify your fork to add features or fix bugs and send them back to the owner as a proposed change.
+### GitHub Pages for course website
 
-### GitHub as web presence
+All course content is provided on the [STAT 545 website](http://stat545.com).
+Each page is generated from an R Markdown document that is rendered to HTML locally using the rmarkdown package, retaining the intermediate Markdown.
+These pages are a mix of prose and rendered R code, reflecting the live coding done in class.
+All of these files and their history can be explored in the [source repository](https://github.com/STAT545-UBC/STAT545-UBC.github.io).
+The TA team has permission to write to this repo, meaning they can (and do!) help me maintain the website.
+I rejoice that I am no longer the webmaster.
+We also get typo corrections and other input from the world at large, since this is entirely public.
 
-Simply having a project on GitHub gives it a web presence.
+If I were starting from scratch today, I would continue to use R Markdown, RStudio, and GitHub Pages (see below), but would upgrade to a more modern, automated approach to rendering the pages.
+I now recommend [R Markdown websites](http://rmarkdown.rstudio.com/rmarkdown_websites.html), [bookdown](https://bookdown.org), or [blogdown](https://bookdown.org/yihui/blogdown/) to manage the process of creating a static website from a large and inter-related set of `.Rmd` files.
 
-If someone needs to see your work or if you want them to try out your code, they can easily get it from GitHub.
-If they use Git, they can clone or fork your repository.
-If they don't use Git, they can still browse your project on GitHub like a normal website and even grab everything by downloading a zip archive.
+GitHub offers several ways to host a website directly from a repository, collectively known as [GitHub Pages](https://help.github.com/categories/github-pages-basics/).
+The STAT 545 website is a very simple [Organization Page](https://help.github.com/articles/user-organization-and-project-pages/) that uses a [custom domain](https://help.github.com/articles/custom-domain-redirects-for-github-pages-sites/), `stat545.com`, instead of the default `orgname.github.io`.
 
-There are many things you can do a "working" repository to make it double as a project website.
+This system for managing course content is a great example of integrating the doing of work and the sharing of it.
+We analyze data live in class, using R, based on the scripts on the website.
+I re-render the associated `.R` or `.Rmd`, commit the changed files, push, and see it reflected right away on http://stat545.com.
+There is no separation between having an idea, implementing it, and posting on the website.
 
-  * [Make a GitHub repo browsable](http://happygitwithr.com/repo-browsability.html)
+### Student-specific private repos
 
-If you want a more proper website, GitHub, rmarkdown, and, especially, the combination provide several options:
+Early in the course I elicit GitHub usernames for registered students, via a [Shiny app](https://shiny.rstudio.com), and invite them to join the course Organization.
+I then create one private repository per student, in the STAT 545 Organization.
+The targeted student has write access and the other students have read access.
+This is somewhat controversial, due to the possibility of cheating, but I have seen more pros than cons for this setup, in the STAT 545 context.
+In other settings, I have also used one repo per student *per homework assignment*, which allows you to keep the repos completely private until homework submission, then increase their visibility during marking and peer review.
+Some courses will work better with one model or the other.
 
-  * https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3
-  * https://help.github.com/categories/github-pages-basics/
-  * http://rmarkdown.rstudio.com/rmarkdown_websites.html
-  * https://bookdown.org/yihui/bookdown/
-  * https://bookdown.org/yihui/blogdown/
-  * https://hadley.github.io/pkgdown/
+Each student does their work in this repo, submitting a major assignment approximately once a week.
+The first assignment is simply to claim the repository and create a README, which proves they have all the relevant software setup and they can write a little Markdown.
+Each week we tackle some new data analysis or wrangling task, with increasing latitude for independence.
+Homework is implemented in R Markdown documents, rendered to Markdown, and pushed to GitHub.
+Students submit their work by opening an issue in their repo, naming the assignment in the title, providing the SHA of the associated final commit, and linking to the main `.md` file.
+We leave feedback as comments in the issue thread or, occasionally, propose changes to code via "pull requests".
+Two peers are selected at random to review each assignment, a process that we also implement via GitHub Issues.
+
+At the end of term, the student (and their instructor!) can visit the repo to find an organized, navigable sequence of ~10 assignments.
+Each student leaves with self-written documentation of everything they've done, ready to consult in future projects.
+The last assignments require writing an R package or Shiny app, which they generally do in public repositories under their own accounts.
+They finish STAT 545 with several months of Git/GitHub experience and the start of a data science portfolio.
+
+## GitHub as web presence
+
+Simply having a project on GitHub gives it a web presence!
+Non-users of Git/GitHub can visit the project in the browser and interact with it like a webpage.
+They can grab a snapshot of all the files as a ZIP archive by simply clicking a button.
+People with GitHub accounts have even more options.
+They can clone or fork the repository to get their own copy, which also makes it easy for them to stay current on future changes.
+
+As described above, GitHub Pages offer various ways to serve proper websites, even quite sophisticated ones, directly out of a GitHub repo.
+But before you even worry about that, certain practices can make a GitHub repository much more browsable.
+For many projects, this is more than sufficient for granting people access to your work.
+
+**Be savvy about file formats.** Keep files in the plainest, web-friendliest form that is compatible with your main goals.
+As explained above, Markdown is the ideal format for prose, because it is just plain text with some markup, but will be displayed like HTML on GitHub.
+Files named `README.md` are extra special, acting as the index or landing page for their host directory.
+[CSV and TSV](https://help.github.com/articles/rendering-csv-and-tsv-data/) files also get special treatment, including an attractive grid layout and search.
+GitHub has excellent support for displaying and diffing common image formats.
+
+**Use conventional file extensions.** GitHub is very code-focused and will apply proper syntax high-lighting for almost any language you can think of, if you use one of the standard file extensions.
+This also has advantages for people searching GitHub and trying to filter by language.
+
+**Use internal links.** `README.md` is a great place to explain how your project fits together.
+Any Markdown file can include relative links to other files in the repo.
+Embedded images are also displayed.
+For figures produced by R code, these links are part of what rmarkdown takes care of for you, but there's no reason you can't do the same yourself for any Markdown file.
+
+*link to http://happygitwithr.com/repo-browsability.html*
+
+## More resources
+
+I've tried to convey the main points about the use of Git and GitHub in statistical and data analytical settings, but I've had to leave many things out.
+There are more advanced topics that will come up once your use of Git becomes more sophisticated and there are topics that are only relevant to certain types of reader.
+
+I targeted [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity.
+However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms.
+I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today.
+I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.
+For example, we host our own instance of GitHub Enterprise at UBC to support our Master of Data Science program
+
+Don't fret too much about public versus private repositories at this point.
+All the major hosting providers offer private repositories with flexible control over who can read or write to the repo.
+There are many ways to get private repositories for low or no cost, especially for academics.
+If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem.
+You can either pay for a higher level of service, self-host one of these platforms, or advocate for organization-wide solutions.
+
+Branches and pull requests are an extremely powerful feature of Git/GitHub and should be your first foray beyond the basics of commit, push, and pull.
+A branch is an independent line of development within a repo, with the notion it will eventually be merged back into the main branch, a.k.a.
+"master".
+In a course website, you might work in a branch to update a series of lessons, while leaving the current version intact in the meantime.
+A pull request is a specific and formal GitHub process for merging one branch into another.
+You can make a pull request between two branches in the same repo or between two copies of a repo.
+This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages *link to a search?*.
+
+## Conclusion
+
+*NEEDS ONE ... BUT SHORT ... SOME FODDER*
+
+Transparency about process and product is increasingly important in science.
+The SOMETHING for reproducibility is well accepted.
+A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.
 
@@
 
-<https://help.github.com/categories/working-with-non-code-files/>
-
-<https://help.github.com/articles/rendering-and-diffing-images/>
-
 make statistical thought and implementation available
-
-Transparency about process and product is increasingly important in science.
-The SOMETHING for reproducibility is well accepted.
-A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.
-
-Who can do what?
-
-A public repository is readable by the world. The owner can grant higher levels of permission to others, such as the ability to push commits.
-
-A private repository is invisible to the world. The owner can grant read, write (push), or admin access to others.
-
-There is also a formal notion of an organization, which can be useful for managing repository permissions for entire teams of people.
-
-why sharing intermediate md is better than html (not readable) or pdf (merge conflicts).
-markdown is the magical intersection of git-friendly (plain text) and human-friendly (becasue GitHub renders it nicely)
 
@@
 
-my tips around md are the large concept of "make things navigable/readable" of GitHub, particularized to R code
+https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem
 
-https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem
+https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3
+
+The active R package development community on GitHub.
+
+  * If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub.
+You can watch the repository to get notified of major activity.
+You can fork it to keep your own copy.
+You can modify your fork to add features or fix bugs and send them back to the owner as a proposed change.
+
+The read-only mirrors of R source and all of CRAN. Coupled with GitHub search features, you can answer alot of your own questions this way.
+
+[RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).

### 2017-05-18 README.Rmd
Intro tweaks

@@
 
-Benefits of using hosted version control in your statistical practice:
+There many benefits of using hosted version control in your statistical practice:
 
   * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
+  * Collaboration becomes much more structured, with powerful tools for asynchronous work and managing versions.
   * The marginal effort required to create a web presence for a project is negligible.
@@
 Its original purpose was to help groups of developers work collaboratively on big software projects.
-Git manages the evolution of a set of files -- called a __repository__ -- in a sane, highly structured way.
+Git manages the evolution of a set of files -- called a __repository__ or __repo__ -- in a sane, highly structured way.
 It is a bit like the "Track Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up to multiple files.

### 2017-05-18 README.Rmd
First success using the rticles templagte :tada:

@@
 title: "Excuse me, do you have a moment to talk about version control?"
-author: "Jenny Bryan"
-date: "`r format(Sys.Date())`"
+authors:
+- affiliation: RStudio and the Department of Statistics, University of British Columbia
+  name: Jennifer Bryan
+  thanks: The author gratefully acknowledges the constructive feedback from reviewers
+    Nicholas Horton and Colin Rundel.
+blinded: 0
 output:
+  rticles::asa_article:
+    pandoc_args: --listings
   github_document:
-    toc: true
+    toc: yes
+keywords: version control
+abstract: Abstract abstract abstract.
 ---
+
+```{r}
+is_github_document <-
+  knitr::opts_knit$get("rmarkdown.pandoc.to") == "markdown_github"
+```
 
@@
 
-There many benefits of using hosted version control in your statistical practice:
+There are many benefits of using hosted version control in your statistical practice:
 
@@
 
-*here's one way I see phylogeny in my mind*
-![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)
+*If you're looking at the PDF draft, you're missing out on some figures around here in the GitHub Markdown version that hint at the figure I plan to embed.*
 
-*make such a phylogeny with a set of crazy filenames like this*
-![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)
+```{r eval = is_github_document, echo = FALSE, results = 'asis'}
+cat("*here's one way I see phylogeny in my mind*\n")
+cat("![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)")
 
-*or riff on the famous PhD Comic*
-![](http://phdcomics.com/comics/archive/phd101212s.gif)
+cat("*make such a phylogeny with a set of crazy filenames like this*\n")
+cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")
+
+cat("*or riff on the famous PhD Comic*")
+cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
+```
 

### 2017-05-18 README.Rmd
Don't include the output-format-detecting chunk

@@
 output:
+  github_document:
+    toc: yes
   rticles::asa_article:
     pandoc_args: --listings
-  github_document:
-    toc: yes
 keywords: version control
@@
 
-```{r}
+```{r include = FALSE}
 is_github_document <-

### 2017-05-18 README.Rmd
Work on the yaml

--- 2017-05-18 README.Rmd
@@
 authors:
-- affiliation: RStudio and the Department of Statistics, University of British Columbia
-  name: Jennifer Bryan
-  thanks: The author gratefully acknowledges the constructive feedback from reviewers
+- name: Jennifer Bryan
+  affiliation: |
+    RStudio and the Department of Statistics, University of British Columbia
+  thanks: |
+    The author gratefully acknowledges the constructive feedback from reviewers
     Nicholas Horton and Colin Rundel.
@@
 output:
-  rticles::asa_article:
-    pandoc_args: --listings
+  rticles::asa_article: null
   github_document:
@@
 
-```{r}
+```{r include = FALSE}
+## I temporarily embed some figures from the internet and LaTeX just. can't.
+## deal. with. it.
 is_github_document <-
@@
 
-Why would a practicing statistician use version control, such as Git? And what is the point of hosting your work online, e.g., on GitHub? Could the gains possibly justify the inevitable pain?
+Why would a statistician use version control, such as Git? And what is the point of hosting your work online, e.g., on GitHub? Could the gains possibly justify the inevitable pain?
 
@@
   * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
-  * Collaboration becomes much more structured, with powerful tools for asynchronous work and managing versions.
+  * Collaboration is much more structured, with powerful tools for asynchronous work and managing versions.
   * The marginal effort required to create a web presence for a project is negligible.

### 2017-05-19 README.Rmd
I. HAVE. MADE. A. BIBLIOGRAPHY. 🔥
Make both formats play nice re: bibliography

@@
 abstract: Abstract abstract abstract.
+bibliography: git-github-for-stats.bib
 ---
@@
 
-Transparency about process and product is increasingly important in science.
+Transparency about [@greenwade93] process and product is increasingly important in science.
 The SOMETHING for reproducibility is well accepted.
@@
 [RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).
+
+```{r eval = is_github_document, echo = FALSE, results = 'asis'}
+## surely there is a better way?
+cat("## References\n")
+```

### 2017-05-19 README.tex
Make both formats play nice re: bibliography

--- 2017-05-18 README.Rmd
@@
----
-title: "Excuse me, do you have a moment to talk about version control?"
-authors:
-- name: Jennifer Bryan
-  affiliation: |
-    RStudio and the Department of Statistics, University of British Columbia
-  thanks: |
-    The author gratefully acknowledges the constructive feedback from reviewers
-    Nicholas Horton and Colin Rundel.
-blinded: 0
-output:
-  rticles::asa_article: null
-  github_document:
-    toc: yes
-keywords: version control
-abstract: Abstract abstract abstract.
----
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{listings}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
 
-```{r include = FALSE}
-## I temporarily embed some figures from the internet and LaTeX just. can't.
-## deal. with. it.
-is_github_document <-
-  knitr::opts_knit$get("rmarkdown.pandoc.to") == "markdown_github"
-```
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-## Why Git?
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
 
-Why would a statistician use version control, such as Git? And what is the point of hosting your work online, e.g., on GitHub? Could the gains possibly justify the inevitable pain?
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Excuse me, do you have a moment to talk about version control?}
+
+  \author{
+        Jennifer Bryan \thanks{The author gratefully acknowledges the constructive feedback from
+reviewers Nicholas Horton and Colin Rundel.} \\
+    RStudio and the Department of Statistics, University of British Columbia\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Excuse me, do you have a moment to talk about version control?}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+Abstract abstract abstract.
+\end{abstract}
+
+\noindent%
+{\it Keywords:} version control
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
+
+\subsection{Why Git?}\label{why-git}
+
+Why would a statistician use version control, such as Git? And what is
+the point of hosting your work online, e.g., on GitHub? Could the gains
+possibly justify the inevitable pain?
 
@@
 
-There are many benefits of using hosted version control in your statistical practice:
+There are many benefits of using hosted version control in your
+statistical practice:
 
-  * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
-  * Collaboration is much more structured, with powerful tools for asynchronous work and managing versions.
-  * The marginal effort required to create a web presence for a project is negligible.
-  * GitHub makes a fantastic course management system for courses that use R. You and your students can exchange actual working code and explore the associated results.
-  * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.
+\begin{itemize}
+\tightlist
+\item
+  Doing your work becomes tightly integrated with organizing, recording,
+  and disseminating it. It's not a separate, burdensome task you are
+  tempted to neglect.
+\item
+  Collaboration is much more structured, with powerful tools for
+  asynchronous work and managing versions.
+\item
+  The marginal effort required to create a web presence for a project is
+  negligible.
+\item
+  GitHub makes a fantastic course management system for courses that use
+  R. You and your students can exchange actual working code and explore
+  the associated results.
+\item
+  By using common mechanics across work modes (research, teaching,
+  analysis), you achieve basic competence quickly and avoid the
+  demoralizing forget-relearn cycle.
+\end{itemize}
 
-## What is Git?
+\subsection{What is Git?}\label{what-is-git}
 
-[Git](http://git-scm.com) is a __version control system__.
-Its original purpose was to help groups of developers work collaboratively on big software projects.
-Git manages the evolution of a set of files -- called a __repository__ or __repo__ -- in a sane, highly structured way.
-It is a bit like the "Track Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up to multiple files.
+\href{http://git-scm.com}{Git} is a \textbf{version control system}. Its
+original purpose was to help groups of developers work collaboratively
+on big software projects. Git manages the evolution of a set of files --
+called a \textbf{repository} or \textbf{repo} -- in a sane, highly
+structured way. It is a bit like the ``Track Changes'' feature from
+Microsoft Word, but more rigorous, powerful, and scaled up to multiple
+files.
 
-Git has been re-purposed by the data science community.
-In addition to using it for source code, we use it to manage the motley collection of files that make up typical data analytical projects, which often consist of data, figures, reports, and, yes, source code.
-Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.
+Git has been re-purposed by the data science community. In addition to
+using it for source code, we use it to manage the motley collection of
+files that make up typical data analytical projects, which often consist
+of data, figures, reports, and, yes, source code. Even those who
+identify more as statistician than data scientist generally have a
+similar mix of files that are the artifacts of a project.
 
-A lone ranger, working on a single computer, can benefit from adopting version control.
-But not nearly enough to justify the pain of installation and workflow upheaval.
-There are much easier ways to get versioned back ups of files, if that's all you're worried about.
+A lone ranger, working on a single computer, can benefit from adopting
+version control. But not nearly enough to justify the pain of
+installation and workflow upheaval. There are much easier ways to get
+versioned back ups of files, if that's all you're worried about.
 
-In my opinion, **for new users**, the pros of Git only outweigh the cons when you factor in the overhead of communicating and collaborating with other people, including your future self.
-And who among us does not need to do that?
-Your life is much easier if this is baked into your workflow, as opposed to being a separate process that you dread and avoid.
-Communication and collaboration are the killer apps of version control.
-Git enforces a rigorous model of file management, but it is critical for the distribution of files across different people, computers, and time.
+In my opinion, \textbf{for new users}, the pros of Git only outweigh the
+cons when you factor in the overhead of communicating and collaborating
+with other people, including your future self. And who among us does not
+need to do that? Your life is much easier if this is baked into your
+workflow, as opposed to being a separate process that you dread and
+avoid. Communication and collaboration are the killer apps of version
+control. Git enforces a rigorous model of file management, but it is
+critical for the distribution of files across different people,
+computers, and time.
 
-This has an implication for selecting your first Git projects: in order to see substantial gain for your pain, you need to pick projects that require you to share rapidly evolving files with others.
-It is tempting to pick a quiet little project that you work on privately, but you risk missing the main point of formal version control.
+This has an implication for selecting your first Git projects: in order
+to see substantial gain for your pain, you need to pick projects that
+require you to share rapidly evolving files with others. It is tempting
+to pick a quiet little project that you work on privately, but you risk
+missing the main point of formal version control.
 
-Many people who don't use Git unwittingly re-invent a poor man's version of it.
-They take an important file and distribute it via email.
-Various parties make changes, decorating the file name with initials, dates, and other descriptors.
-Before you know it, the original file is the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve.
+Many people who don't use Git unwittingly re-invent a poor man's version
+of it. They take an important file and distribute it via email. Various
+parties make changes, decorating the file name with initials, dates, and
+other descriptors. Before you know it, the original file is the root of
+a complicated phylogeny that no amount of ``Track changes'' and good
+intentions can resolve.
 
-*If you're looking at the PDF draft, you're missing out on some figures around here in the GitHub Markdown version that hint at the figure I plan to embed.*
+\emph{If you're looking at the PDF draft, you're missing out on some
+figures around here in the GitHub Markdown version that hint at the
+figure I plan to embed.}
 
-```{r eval = is_github_document, echo = FALSE, results = 'asis'}
-cat("*here's one way I see phylogeny in my mind*\n")
-cat("![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)")
+The Git way is to track the evolution of that file, through a series of
+commits, each equipped with an explanatory message and a nickname. All
+collaborators sync regularly to a common version, acknowledging that the
+difficulty of merging goes up faster than the size of the difference
+(cite good enough). Especially important versions get a human-readable
+tag, to signal a meaningful milestone. Yes, there is some pain in
+adopting the formalism of Git, but it is worth it.
 
-cat("*make such a phylogeny with a set of crazy filenames like this*\n")
-cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")
+\emph{another good image to emulate, from the Bartlett Git for Designers
+talk} \includegraphics{bartlett-copies-vs-commits.png}
 
-cat("*or riff on the famous PhD Comic*")
-cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
-```
+\subsection{Who should read this and what to
+expect}\label{who-should-read-this-and-what-to-expect}
 
-The Git way is to track the evolution of that file, through a series of commits, each equipped with an explanatory message and a nickname.
-All collaborators sync regularly to a common version, acknowledging that the difficulty of merging goes up faster than the size of the difference (cite good enough).
-Especially important versions get a human-readable tag, to signal a meaningful milestone.
-Yes, there is some pain in adopting the formalism of Git, but it is worth it.
+The target reader is anyone who does statistical research, analysis, or
+instruction. Those whose work is some combination of these three may
+find the work style described here especially rewarding.
 
-*another good image to emulate, from the Bartlett Git for Designers talk*
-![](bartlett-copies-vs-commits.png)
+This article does not provide step-by-step instructions on how to use
+Git and GitHub. This format would not be effective, but we do provide
+annotated links to such resources in ?the appendix or online
+supplement?. Instead, I'll convey what the workflow feels like and what
+the payoffs are, with special attention to the statistics and R context.
+The goal is to help the Git-curious generate the activation energy
+needed to get started.
 
-## Who should read this and what to expect
+\subsection{What is GitHub?}\label{what-is-github}
 
-The target reader is anyone who does statistical research, analysis, or instruction. Those whose work is some combination of these three may find the work style described here especially rewarding.
-
-This article does not provide step-by-step instructions on how to use Git and GitHub.
-This format would not be effective, but we do provide annotated links to such resources in ?the appendix or online supplement?.
-Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context.
-The goal is to help the Git-curious generate the activation energy needed to get started.
-
-## What is GitHub?
-
-[GitHub](https://github.com) is currently the most popular Git hosting service.
-Others include [GitLab](https://about.gitlab.com) and [Bitbucket](https://bitbucket.org).
-These services provide a home for Git-based projects on the internet.
-It is a bit like DropBox or Google Drive, but more structured, powerful, and programmatic.
+\href{https://github.com}{GitHub} is currently the most popular Git
+hosting service. Others include \href{https://about.gitlab.com}{GitLab}
+and \href{https://bitbucket.org}{Bitbucket}. These services provide a
+home for Git-based projects on the internet. It is a bit like DropBox or
+Google Drive, but more structured, powerful, and programmatic.
 
 The remote host acts as the clearinghouse for a Git-managed project.
-This allows other people to see the project files, sync up with the current version, and perhaps even make or propose changes.
-These hosting providers offer well-designed web interfaces that are a dramatic improvement over traditional Unix Git servers.
-Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files.
-These hosts also offer granular control over who can see, edit, and administer a project.
+This allows other people to see the project files, sync up with the
+current version, and perhaps even make or propose changes. These hosting
+providers offer well-designed web interfaces that are a dramatic
+improvement over traditional Unix Git servers. Many operations can be
+done entirely in the browser, such as viewing current or past versions
+of a file, commenting on a recent change, and editing or adding files.
+These hosts also offer granular control over who can see, edit, and
+administer a project.
 
-Even for private solo projects, there are two advantages to keeping a synced copy on GitHub:
+Even for private solo projects, there are two advantages to keeping a
+synced copy on GitHub:
 
-  1.
-When you are new with Git (or, frankly, even when you're not), it's common to goof up the Git infrastructure for a project.
-Note that your files can absolutely be intact and safe, even while the Git tracking is a bit confused.
-Of course there are official Git remedies, but sometimes the easiest fix is to clone a fresh copy from GitHub, patch things up with the changes that only exist locally, and move on with your life.
-This workaround obviously requires the existence of a recent copy on GitHub.
-  2.
-The highly functional web interfaces mentioned above are often the most pleasant and natural way to navigate and search your files, even though all the same information exists locally.
-It is a pleasure to browse through your own work, across multiple projects or files and across time, as if it's a well-designed website.
-You must push your work to GitHub to enjoy this.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  When you are new with Git (or, frankly, even when you're not), it's
+  common to goof up the Git infrastructure for a project. Note that your
+  files can absolutely be intact and safe, even while the Git tracking
+  is a bit confused. Of course there are official Git remedies, but
+  sometimes the easiest fix is to clone a fresh copy from GitHub, patch
+  things up with the changes that only exist locally, and move on with
+  your life. This workaround obviously requires the existence of a
+  recent copy on GitHub.
+\item
+  The highly functional web interfaces mentioned above are often the
+  most pleasant and natural way to navigate and search your files, even
+  though all the same information exists locally. It is a pleasure to
+  browse through your own work, across multiple projects or files and
+  across time, as if it's a well-designed website. You must push your
+  work to GitHub to enjoy this.
+\end{enumerate}
 
-GitHub __issues__ are another powerful feature of the platform.
-Recall that we are repurposing Git, a tool designed to facilitate software development.
-The issues for a project are its bug tracker.
-For projects that are not pure software development, we co-opt this machinery to organize our to do list more generally.
-The basic unit is an issue and you can interact with them in two ways.
+GitHub \textbf{issues} are another powerful feature of the platform.
+Recall that we are repurposing Git, a tool designed to facilitate
+software development. The issues for a project are its bug tracker. For
+projects that are not pure software development, we co-opt this
+machinery to organize our to do list more generally. The basic unit is
+an issue and you can interact with them in two ways.
 
-First, issues are integrated into the project's web interface on GitHub, with a rich set of options for linking to project files and incremental changes.
-Second, issues and their associated comment threads appear in your email, just like regular messages (this can, of course, be configured).
-The result is that all correspondence about a project comes through your normal channels, but is also tracked inside the project itself, with excellent navigability and search capabilities.
-For software, issues are used to track bugs and feature requests.
-In a data analysis project, you might open an issue to flesh out a specific sub-analysis or to develop a complicated figure.
-In a course, we use them to manage homework submission, marking, and peer review.
+First, issues are integrated into the project's web interface on GitHub,
+with a rich set of options for linking to project files and incremental
+changes. Second, issues and their associated comment threads appear in
+your email, just like regular messages (this can, of course, be
+configured). The result is that all correspondence about a project comes
+through your normal channels, but is also tracked inside the project
+itself, with excellent navigability and search capabilities. For
+software, issues are used to track bugs and feature requests. In a data
+analysis project, you might open an issue to flesh out a specific
+sub-analysis or to develop a complicated figure. In a course, we use
+them to manage homework submission, marking, and peer review.
 
 Issues can be assigned to specific people and they can be labelled, e.g.
-"bug", "simulation-study", or "final-exam".
-Coupled with the ability to cross-link issues and the project files or file changes, you have extraordinary power to document why things have happened in the past and to organize what needs to happen in the future.
+``bug'', ``simulation-study'', or ``final-exam''. Coupled with the
+ability to cross-link issues and the project files or file changes, you
+have extraordinary power to document why things have happened in the
+past and to organize what needs to happen in the future.
 
-## Initial system setup
+\subsection{Initial system setup}\label{initial-system-setup}
 
@@
 
-  * Register for a free account with GitHub.
+\begin{itemize}
+\item
+  Register for a free account with GitHub.
+\item
+  Install Git. Depending on your OS, Git might already be installed. But
+  many of us will need to install it or might choose to update to the
+  most recent version. Some basic configuration is critical, such as
+  setting your username and email.
+\item
+  Install a local Git client. Optional but highly recommended. A Git
+  client is software that provides a graphical user interface for Git,
+  which is otherwise command-line only. If you are an R user, you will
+  find that RStudio provides a great deal of this functionality. There
+  are some notable gaps, however, so you might still choose to install a
+  dedicated and comprehensive Git client such as SourceTree or
+  GitKraken. Git is a file-based system, so you can do some operations
+  from RStudio, others from SourceTree, and others from the shell.
+\item
+  Confirm, with a practice repository, that local Git can send and
+  receive the current version of the repository on GitHub, known as
+  \emph{pushing} and \emph{pulling}, respectively. This will require
+  authenticating yourself with GitHub from a local shell or Git client.
+  At this point, most people elect to do a bit of extra setup to ensure
+  that they are not repeatedly challenged for their GitHub credentials
+  going forward.
+\end{itemize}
 
-  * Install Git.
-Depending on your OS, Git might already be installed.
-But many of us will need to install it or might choose to update to the most recent version.
-Some basic configuration is critical, such as setting your username and email.
+Once this setup is done, you are ready to start using Git and GitHub
+with your projects.
 
-  * Install a local Git client.
-Optional but highly recommended.
-A Git client is software that provides a graphical user interface for Git, which is otherwise command-line only.
-If you are an R user, you will find that RStudio provides a great deal of this functionality.
-There are some notable gaps, however, so you might still choose to install a dedicated and comprehensive Git client such as SourceTree or GitKraken.
-Git is a file-based system, so you can do some operations from RStudio, others from SourceTree, and others from the shell.
-
-  * Confirm, with a practice repository, that local Git can send and receive the current version of the repository on GitHub, known as *pushing* and *pulling*, respectively.
-This will require authenticating yourself with GitHub from a local shell or Git client.
-At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.
-
-Once this setup is done, you are ready to start using Git and GitHub with your projects.
-
-## Is this going to hurt?
+\subsection{Is this going to hurt?}\label{is-this-going-to-hurt}
 
@@
 
-Git was built neither for the exact usage described here, nor for broad usability.
-You will undoubtedly notice this, so it can be helpful to know this in advance.
-Happily there are many helpful tools that mediate your interactions with Git.
-GitHub itself is a fine example.
-In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.
+Git was built neither for the exact usage described here, nor for broad
+usability. You will undoubtedly notice this, so it can be helpful to
+know this in advance. Happily there are many helpful tools that mediate
+your interactions with Git. GitHub itself is a fine example. In addition
+to pointing out tools that soften Git's sharpest edges, I recommend
+specific habits and attitudes that reduce frustration.
 
@@
 
-  * I repeat: consider using a graphical front-end for Git, a.k.a. a "Git client", versus restricting yourself to the command line interface.
-  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
-  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
-  * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.
+\begin{itemize}
+\tightlist
+\item
+  I repeat: consider using a graphical front-end for Git, a.k.a. a ``Git
+  client'', versus restricting yourself to the command line interface.
+\item
+  Establish confidence in the basics (e.g.~make a change, commit it,
+  push it) before wading into more advanced usage (e.g.~branching).
+\item
+  Commit yourself to Git usage on a project that will provide sustained
+  practice over several months. Usage in a course is great, because it
+  provides a relentless stream of small deadlines.
+\item
+  Realize that no one is giving out Git style points. It's ok to
+  ``power-cycle'', i.e.~re-initialize the Git repository, to get
+  unstuck.
+\end{itemize}
 
-## Repositories and workflow
+\subsection{Repositories and workflow}\label{repositories-and-workflow}
 
@@
 
-  * Dedicate a local directory (a.k.a "folder") to it.
-  * Make it an RStudio Project. Optional but recommended; obviously only applies to projects involving R and users of RStudio.
-  * Make it a Git repository.
+\begin{itemize}
+\tightlist
+\item
+  Dedicate a local directory (a.k.a ``folder'') to it.
+\item
+  Make it an RStudio Project. Optional but recommended; obviously only
+  applies to projects involving R and users of RStudio.
+\item
+  Make it a Git repository.
+\end{itemize}
 
-This is once-per-project setup and can happen at project inception or at any later point.
-Chances are your project already lives in a dedicated directory.
-Making this directory an RStudio Project and Git repository boils down to allowing those applications to leave notes for themselves in hidden files or directories.
-The project is still a regular directory on your computer, that you can locate, name, move, and generally interact with as you wish.
-You don't have to handle it with special gloves!
+This is once-per-project setup and can happen at project inception or at
+any later point. Chances are your project already lives in a dedicated
+directory. Making this directory an RStudio Project and Git repository
+boils down to allowing those applications to leave notes for themselves
+in hidden files or directories. The project is still a regular directory
+on your computer, that you can locate, name, move, and generally
+interact with as you wish. You don't have to handle it with special
+gloves!
 
@@
 
-  * Go about your usual business, e.g.
-writing R scripts or authoring reports in LaTeX or R Markdown.
-But instead of only *saving* individual files, periodically you make a **commit**, which takes a snapshot of all the files in the entire project.
-    - Have you ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531)?
-That is effectively a **commit**, albeit only for a single file: it is a version that is significant to you and that you might want to inspect or revert to later.
-  * Push commits to GitHub periodically.
-    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment.
-By pushing to GitHub, you make your work and all your accumulated progress accessible to others.
+\begin{itemize}
+\tightlist
+\item
+  Go about your usual business, e.g.~writing R scripts or authoring
+  reports in LaTeX or R Markdown. But instead of only \emph{saving}
+  individual files, periodically you make a \textbf{commit}, which takes
+  a snapshot of all the files in the entire project.
 
-This is a moderate change to your normal, daily workflow.
-It feels weird at first, but quickly becomes second nature.
-In [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub, starting in week one.
-Most have never seen Git before and do not identify as "programmers".
-It is a major topic in class and office hours for the first two weeks.
-Then we practically never discuss it again.
+  \begin{itemize}
+  \tightlist
+  \item
+    Have you ever versioned a file
+    \href{http://www.phdcomics.com/comics/archive.php?comicid=1531}{by
+    adding your initials or the date}? That is effectively a
+    \textbf{commit}, albeit only for a single file: it is a version that
+    is significant to you and that you might want to inspect or revert
+    to later.
+  \end{itemize}
+\item
+  Push commits to GitHub periodically.
 
-*I am not saying anything about pulling and collaboration here, which seems a bit extreme, even if I can't say much*
+  \begin{itemize}
+  \tightlist
+  \item
+    This is like sharing a document with colleagues on DropBox or
+    sending it out as an email attachment. By pushing to GitHub, you
+    make your work and all your accumulated progress accessible to
+    others.
+  \end{itemize}
+\end{itemize}
 
-## Commits, diffs, and tags
+This is a moderate change to your normal, daily workflow. It feels weird
+at first, but quickly becomes second nature. In
+\href{http://stat545.com}{STAT 545} students are required to submit all
+coursework via GitHub, starting in week one. Most have never seen Git
+before and do not identify as ``programmers''. It is a major topic in
+class and office hours for the first two weeks. Then we practically
+never discuss it again.
 
-We now explore the fundamental concepts of Git and connect them to the data science workflow:
+\emph{I am not saying anything about pulling and collaboration here,
+which seems a bit extreme, even if I can't say much}
 
-  * repository
-  * commit
-  * diff
+\subsection{Commits, diffs, and tags}\label{commits-diffs-and-tags}
 
-Recall that a repository -- or "repo" -- is just a directory of files that Git is going to manage holistically.
-A commit functions like a snapshot of all the files in the repo, at a specific moment.
-Under the hood, that is not exactly how Git implements things.
-Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.
+We now explore the fundamental concepts of Git and connect them to the
+data science workflow:
 
-*this is really crying out for an example and/or diagrams, something that shows commits unfolding and let's you illustrate a diff, sort of like this from the barlett talk*
+\begin{itemize}
+\tightlist
+\item
+  repository
+\item
+  commit
+\item
+  diff
+\end{itemize}
 
-![](bartlett-commit-history.png)
+Recall that a repository -- or ``repo'' -- is just a directory of files
+that Git is going to manage holistically. A commit functions like a
+snapshot of all the files in the repo, at a specific moment. Under the
+hood, that is not exactly how Git implements things. Mental models don't
+have to be accurate in order to be useful, but in this case there's some
+value in aligning the two.
 
-Consider version A of a file and a modified version, version B.
-Assume that version A was part of a Git commit and version B was part of the next commit.
-The set of differences between A and B is called a "diff" and Git users contemplate diffs a lot.
-Diff inspection is how you re-explain to yourself how version A differs from version B.
-Diff inspection is not limited to adjacent commits.
-You can inspect the diffs between any two commits.
+\emph{this is really crying out for an example and/or diagrams,
+something that shows commits unfolding and let's you illustrate a diff,
+sort of like this from the barlett talk}
 
-In fact, Git's notion of version B of your file is as an accumulation of diffs.
-At some earlier point, the file was created in the first place.
+\begin{figure}
+\centering
+\includegraphics{bartlett-commit-history.png}
+\caption{}
+\end{figure}
+
+Consider version A of a file and a modified version, version B. Assume
+that version A was part of a Git commit and version B was part of the
+next commit. The set of differences between A and B is called a ``diff''
+and Git users contemplate diffs a lot. Diff inspection is how you
+re-explain to yourself how version A differs from version B. Diff
+inspection is not limited to adjacent commits. You can inspect the diffs
+between any two commits.
+
+In fact, Git's notion of version B of your file is as an accumulation of
+diffs. At some earlier point, the file was created in the first place.
 That version of the file was part of a commit and, therefore, a diff.
-Git stores Version A of the file as the initial version, plus all the intervening diffs in the history that affect the file.
-And version B simply includes one more.
-We'll set these internal details aside now, but understanding the importance of these deltas will eventually make Git's operations less baffling.
+Git stores Version A of the file as the initial version, plus all the
+intervening diffs in the history that affect the file. And version B
+simply includes one more. We'll set these internal details aside now,
+but understanding the importance of these deltas will eventually make
+Git's operations less baffling.
 
-So, by looking at diffs, it's easy to see how two snapshots differ, but what about the why?
+So, by looking at diffs, it's easy to see how two snapshots differ, but
+what about the why?
 
-Every time you make a commit you must also write a short *commit message*.
-Ideally, this conveys the *motivation* for the change.
-Remember, the diff will show the content.
-When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the *history*, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.
+Every time you make a commit you must also write a short \emph{commit
+message}. Ideally, this conveys the \emph{motivation} for the change.
+Remember, the diff will show the content. When you revisit a project
+after a break or need to digest recent changes made by a colleague,
+looking at the \emph{history}, by reading commit messages and skimming
+through diffs, is an extremely efficient way to get up to speed.
 
-Every commit needs some sort of nickname, so you can identify it.
-Git does this automatically, assigning each commit what is called a *SHA*, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a SHA-1 checksum hash of the commit).
-Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice.
-You can also designate certain snapshots as special with a *tag*, which is a name of your choosing.
-In a software project, it is typical to tag a release with its version, e.g., "v1.0.3".
-For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.
+Every commit needs some sort of nickname, so you can identify it. Git
+does this automatically, assigning each commit what is called a
+\emph{SHA}, a seemingly random string of 40 letters and numbers (it is
+not, in fact, random but is a SHA-1 checksum hash of the commit). Though
+you will be exposed to these, you don't have to handle them directly
+very often and, when you do, usually the first 7 characters suffice. You
+can also designate certain snapshots as special with a \emph{tag}, which
+is a name of your choosing. In a software project, it is typical to tag
+a release with its version, e.g., ``v1.0.3''. For a manuscript or
+analytical project, you might tag the version submitted to a journal or
+transmitted to external collaborators.
 
-## Markdown is special on GitHub
+\subsection{Markdown is special on
+GitHub}\label{markdown-is-special-on-github}
 
-This may seem unrelated to Git, GitHub, and R, but we need to talk about Markdown.
-Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible.
-The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features.
-Markdown is in wide use on sites like Wikipedia, WordPress, StackOverflow, and, yes, GitHub.
-These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting.
-Do not build this up into some heroic, LaTeX-level learning task, for it is not.
-If you can write an email, you can write Markdown.
+This may seem unrelated to Git, GitHub, and R, but we need to talk about
+Markdown. Markdown is a markup language, like HTML and LaTeX, but
+designed to be as lightweight as possible. The goal is still to separate
+form and content, but also to prioritize human-readability, even at the
+cost of fancy features. Markdown is in wide use on sites like Wikipedia,
+WordPress, StackOverflow, and, yes, GitHub. These sites use Markdown
+because it allows a diverse population of site users to author
+decent-looking web content, with hyperlinks and some formatting. Do not
+build this up into some heroic, LaTeX-level learning task, for it is
+not. If you can write an email, you can write Markdown.
 
 Any file written in Markdown is rendered in an HTML-like way on GitHub.
-In particular, formatting and links "just work".
-This is the last piece we need to seal my claim that merely pushing your project to GitHub gives it a web presence for zero extra work.
-If you make even a modest effort to embed a few explanatory Markdown files in your repo, you will get an automatically-updated project website for free.
-In particular, if a directory has a `README.md` file, GitHub renders it like a home page or "index.html" when people visit that directory in the browser.
-It is very common for a repo to have a top-level `README.md`, but each subdirectory can have its own as well.
+In particular, formatting and links ``just work''. This is the last
+piece we need to seal my claim that merely pushing your project to
+GitHub gives it a web presence for zero extra work. If you make even a
+modest effort to embed a few explanatory Markdown files in your repo,
+you will get an automatically-updated project website for free. In
+particular, if a directory has a \texttt{README.md} file, GitHub renders
+it like a home page or ``index.html'' when people visit that directory
+in the browser. It is very common for a repo to have a top-level
+\texttt{README.md}, but each subdirectory can have its own as well.
 
-## Markdown is special for R users
+\subsection{Markdown is special for R
+users}\label{markdown-is-special-for-r-users}
 
-Markdown is extra special for R users because of R Markdown, which is just Markdown that includes chunks of R code.
-Again, do not regard R Markdown as   something you must clear your schedule to learn.
-If you can write email and a bit of R code, you can write R Markdown.
-The rmarkdown package converts R Markdown (`.Rmd` files) to Markdown (`.md` files), running the code and inserting the results, including figures, into the document.
-This is powered by another package, knitr, under the hood.
-This process is made especially easy in RStudio, but is by no means limited to users of that application.
+Markdown is extra special for R users because of R Markdown, which is
+just Markdown that includes chunks of R code. Again, do not regard R
+Markdown as something you must clear your schedule to learn. If you can
+write email and a bit of R code, you can write R Markdown. The rmarkdown
+package converts R Markdown (\texttt{.Rmd} files) to Markdown
+(\texttt{.md} files), running the code and inserting the results,
+including figures, into the document. This is powered by another
+package, knitr, under the hood. This process is made especially easy in
+RStudio, but is by no means limited to users of that application.
 
-*this too is crying out for a visual Rmd -> md -> rendered thing looking good*
+\emph{this too is crying out for a visual Rmd -\textgreater{} md
+-\textgreater{} rendered thing looking good}
 
-These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described above.
-Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, **and view the results of running that code**, including figures.
-It is the best of all worlds, because the code is revealed and, by definition, is the code that produced the results.
-And yet a reader can gaze upon the product in a web browser, without needing to download the code, install all necessary dependencies, and run it.
+These R-derived Markdown files, if committed and pushed, then enjoy the
+usual privileged treatment on GitHub already described above. Once an
+\texttt{.Rmd} file has been rendered to \texttt{.md}, anyone viewing it
+on GitHub can read the prose, study the R code, \textbf{and view the
+results of running that code}, including figures. It is the best of all
+worlds, because the code is revealed and, by definition, is the code
+that produced the results. And yet a reader can gaze upon the product in
+a web browser, without needing to download the code, install all
+necessary dependencies, and run it.
 
-The overall effect is that a directory that is a GitHub-synced Git repo can simultaneously be the code-heavy back end of a project and an outward-facing front end.
+The overall effect is that a directory that is a GitHub-synced Git repo
+can simultaneously be the code-heavy back end of a project and an
+outward-facing front end.
 
-You do not, in fact, even need to work in R Markdown to exploit this.
-It works with plain R scripts as well.
-You can use exactly the same machinery to prepare a rendered version of an R script, i.e.
-to go from `.R` to `.md`.
-Again, RStudio makes this especially easy, but this is not limited to RStudio.
-Once the Markdown file is pushed to GitHub, it is as if the reader has run your code or is able to look over your shoulder at your R session.
-This provides an lightweight system for exposing work-in-progress to collaborators, without slowing down to create separate reports.
-Comment lines that begin with `#'` are elevated to top-level prose, providing a way to make the document more welcoming for a reader.
-Once there are many prose comments, you might decide to switch from `.R` to `.Rmd`, have proper top-level prose, and move the code down into chunks.
+You do not, in fact, even need to work in R Markdown to exploit this. It
+works with plain R scripts as well. You can use exactly the same
+machinery to prepare a rendered version of an R script, i.e.~to go from
+\texttt{.R} to \texttt{.md}. Again, RStudio makes this especially easy,
+but this is not limited to RStudio. Once the Markdown file is pushed to
+GitHub, it is as if the reader has run your code or is able to look over
+your shoulder at your R session. This provides an lightweight system for
+exposing work-in-progress to collaborators, without slowing down to
+create separate reports. Comment lines that begin with
+\texttt{\#\textquotesingle{}} are elevated to top-level prose, providing
+a way to make the document more welcoming for a reader. Once there are
+many prose comments, you might decide to switch from \texttt{.R} to
+\texttt{.Rmd}, have proper top-level prose, and move the code down into
+chunks.
 
-*figure showing the yaml needed for this? so the use of `github_document` or `keep_md = TRUE`; try to make a figure that does double or triple duty, i.e. illustrates this and points made elsewhere*
+\emph{figure showing the yaml needed for this? so the use of
+\texttt{github\_document} or \texttt{keep\_md\ =\ TRUE}; try to make a
+figure that does double or triple duty, i.e.~illustrates this and points
+made elsewhere}
 
-Let's zoom back out again and consider R Markdown and the rmarkdown package more generally.
-I want to point out that R Markdown can be rendered to many more formats than Markdown!
-I have emphasized the production of Markdown here because it is extremely useful in GitHub-hosted projects and that seems to be underappreciated.
-But the rmarkdown package can convert `.Rmd` to a wide array of output formats, including HTML, PDF, and Word (`.docx`).
-The Markdown must be created regardless.
-Sometimes it is all you need, such as on GitHub.
-In other contexts, it is just a necessary intermediate and may even be discarded.
+Let's zoom back out again and consider R Markdown and the rmarkdown
+package more generally. I want to point out that R Markdown can be
+rendered to many more formats than Markdown! I have emphasized the
+production of Markdown here because it is extremely useful in
+GitHub-hosted projects and that seems to be underappreciated. But the
+rmarkdown package can convert \texttt{.Rmd} to a wide array of output
+formats, including HTML, PDF, and Word (\texttt{.docx}). The Markdown
+must be created regardless. Sometimes it is all you need, such as on
+GitHub. In other contexts, it is just a necessary intermediate and may
+even be discarded.
 
-## Which files to commit
+\subsection{Which files to commit}\label{which-files-to-commit}
 
-The files in a project play different roles and arise in different ways. Let's have a few examples in mind for this discussion:
+The files in a project play different roles and arise in different ways.
+Let's have a few examples in mind for this discussion:
 
-  * R markdown `.Rmd` --> markdown `.md`
-  * R markdown `.Rmd` --> markdown `.md` --> `.html` or `.pdf` or `.docx`
-  * R script `.R` --> results as `.csv` or `.rds` and figures `.png`
-  * LaTeX `.tex` --> WHAT GOES HERE NOW? --> `.pdf`
+\begin{itemize}
+\tightlist
+\item
+  R markdown \texttt{.Rmd} --\textgreater{} markdown \texttt{.md}
+\item
+  R markdown \texttt{.Rmd} --\textgreater{} markdown \texttt{.md}
+  --\textgreater{} \texttt{.html} or \texttt{.pdf} or \texttt{.docx}
+\item
+  R script \texttt{.R} --\textgreater{} results as \texttt{.csv} or
+  \texttt{.rds} and figures \texttt{.png}
+\item
+  LaTeX \texttt{.tex} --\textgreater{} WHAT GOES HERE NOW?
+  --\textgreater{} \texttt{.pdf}
+\end{itemize}
 
-The files at the far left are clearly source files.
-In the case of an R script, this is literally true, but it's morally true for R markdown and LaTeX files too.
-These are files that you directly create and edit "by hand".
+The files at the far left are clearly source files. In the case of an R
+script, this is literally true, but it's morally true for R markdown and
+LaTeX files too. These are files that you directly create and edit ``by
+hand''.
 
-The files at the far right are clearly derived and are often described as "targets".
-These files are programmatically generated from source files (and possibly other inputs).
-These files are the product and they have external value, often for communicating ideas and results.
+The files at the far right are clearly derived and are often described
+as ``targets''. These files are programmatically generated from source
+files (and possibly other inputs). These files are the product and they
+have external value, often for communicating ideas and results.
 
-The files in the middle are intermediates.
-Like targets, they are programmatically generated, but, unlike targets, no one necessarily cares about them.
-However, note that intermediate Markdown `.md` is an exception, since it is extremely useful on GitHub -- much more so than `.html`, `.pdf`, or `.docx` -- and is more like an additional target.
+The files in the middle are intermediates. Like targets, they are
+programmatically generated, but, unlike targets, no one necessarily
+cares about them. However, note that intermediate Markdown \texttt{.md}
+is an exception, since it is extremely useful on GitHub -- much more so
+than \texttt{.html}, \texttt{.pdf}, or \texttt{.docx} -- and is more
+like an additional target.
 
-A critical issue for workflow happiness is figuring out how to manage the production and storage of source, intermediates, and targets with respect to Git.
-You can direct Git to ignore specific files or certain types of files, such as autosaves created by your editor.
-This reduces clutter in your project: Git will not pester you to commit changes to these files and they will not appear in the associated GitHub repository.
-A file that Git does not ignore is said to be *tracked*.
+A critical issue for workflow happiness is figuring out how to manage
+the production and storage of source, intermediates, and targets with
+respect to Git. You can direct Git to ignore specific files or certain
+types of files, such as autosaves created by your editor. This reduces
+clutter in your project: Git will not pester you to commit changes to
+these files and they will not appear in the associated GitHub
+repository. A file that Git does not ignore is said to be
+\emph{tracked}.
 
-The only point on which there is consensus is that source files should absolutely be tracked. The best treatment of intermediates and targets with respect to Git is much less clear cut.
+The only point on which there is consensus is that source files should
+absolutely be tracked. The best treatment of intermediates and targets
+with respect to Git is much less clear cut.
 
-Therefore, the main message for intermediates and targets is that you can pick a policy that works for you and adapt as your needs change.
-There is no right answer.
-I suggest erring on the side of committing everything at first.
+Therefore, the main message for intermediates and targets is that you
+can pick a policy that works for you and adapt as your needs change.
+There is no right answer. I suggest erring on the side of committing
+everything at first.
 
-What are the main considerations when deciding whether to track a derived file or file type?
+What are the main considerations when deciding whether to track a
+derived file or file type?
 
-*Is it immediately useful to someone?* If so, that is a reason to track it and push it to GitHub.
-There is a taboo against committing derived products, inherited from Git's software development roots.
-The reasoning is that compiled programs are platform-specific and, therefore, people are better served by getting current source from Git and compiling themselves.
-I think data analytic targets, like figures and rendered reports, are very different beasts and it's misguided to reflexively exclude them from version control.
-They are immediately useful, especially to consumers of a project (versus the makers).
-To the extent that a GitHub repo is meant for dissemination, there is no reason to burden every consumer with unnecessary friction.
-Most simply will not bother to clone the repo, install all the necessary dependencies, and remake the products.
-Make them readily available.
+\emph{Is it immediately useful to someone?} If so, that is a reason to
+track it and push it to GitHub. There is a taboo against committing
+derived products, inherited from Git's software development roots. The
+reasoning is that compiled programs are platform-specific and,
+therefore, people are better served by getting current source from Git
+and compiling themselves. I think data analytic targets, like figures
+and rendered reports, are very different beasts and it's misguided to
+reflexively exclude them from version control. They are immediately
+useful, especially to consumers of a project (versus the makers). To the
+extent that a GitHub repo is meant for dissemination, there is no reason
+to burden every consumer with unnecessary friction. Most simply will not
+bother to clone the repo, install all the necessary dependencies, and
+remake the products. Make them readily available.
 
-*Is it available elsewhere?* If so, then perhaps you don't need to track it and push it to GitHub.
-Many people who have a policy of not tracking derived products also have a system that makes these available elsewhere, such as on a separate website.
-There are ways to automate this via GitHub, but that is beyond the scope of this article and not recommended for your early days with Git and GitHub.
-Beware those who recommend the exclusion of derived products without offering any practical solution for making them available some other way.
+\emph{Is it available elsewhere?} If so, then perhaps you don't need to
+track it and push it to GitHub. Many people who have a policy of not
+tracking derived products also have a system that makes these available
+elsewhere, such as on a separate website. There are ways to automate
+this via GitHub, but that is beyond the scope of this article and not
+recommended for your early days with Git and GitHub. Beware those who
+recommend the exclusion of derived products without offering any
+practical solution for making them available some other way.
 
-*Is it huge or changing often?
-Is it a format that is of little use on GitHub?* These are all good reasons to not track a file with Git.
-They can make your repository bloated and slow down pushes and pulls.
-If a file is binary, such as a Word document or Excel spreadsheet, Git and Git clients will not be able to provide human-readable diffs.
-Neither will GitHub be able to directly display this file in the browser.
-Be aware, however, GitHub-friendliness does not just boil down to "is it plain text?".
-GitHub has excellent support for non-code files, such as image formats (PNG, JPG, GIF, and SVG) and PDFs.
-It even provides visual diffs, which are extremely useful for understanding changes in figures.
-Finally, even though HTML is plain text, it is of little direct use on GitHub, because, unlike Markdown, it is not rendered.
+\emph{Is it huge or changing often? Is it a format that is of little use
+on GitHub?} These are all good reasons to not track a file with Git.
+They can make your repository bloated and slow down pushes and pulls. If
+a file is binary, such as a Word document or Excel spreadsheet, Git and
+Git clients will not be able to provide human-readable diffs. Neither
+will GitHub be able to directly display this file in the browser. Be
+aware, however, GitHub-friendliness does not just boil down to ``is it
+plain text?''. GitHub has excellent support for non-code files, such as
+image formats (PNG, JPG, GIF, and SVG) and PDFs. It even provides visual
+diffs, which are extremely useful for understanding changes in figures.
+Finally, even though HTML is plain text, it is of little direct use on
+GitHub, because, unlike Markdown, it is not rendered.
 
-*Will diffs be useful to you?* Some derived files are simply too big or miserable to read casually, such as `.csv` files of processed results or `.html` derived from `.Rmd`.
-But they may still be worth tracking with Git, because the diffs are often modest and quite interesting.
-I have caught unexpected changes in analytical results and student-facing webpages this way.
-When you re-run an analysis with updated input data or after updating R packages, the diffs presented by Git help you quickly pinpoint the downstream consequences.
+\emph{Will diffs be useful to you?} Some derived files are simply too
+big or miserable to read casually, such as \texttt{.csv} files of
+processed results or \texttt{.html} derived from \texttt{.Rmd}. But they
+may still be worth tracking with Git, because the diffs are often modest
+and quite interesting. I have caught unexpected changes in analytical
+results and student-facing webpages this way. When you re-run an
+analysis with updated input data or after updating R packages, the diffs
+presented by Git help you quickly pinpoint the downstream consequences.
 
-*Will it make collaboration harder?* *talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>*
+\emph{Will it make collaboration harder?} \emph{talk about the scenario
+Nick brought up re: binary files like PDF being a common source of merge
+conflict
+\url{https://github.com/dsscollection/git-github-for-stats/issues/6}}
 
-In summary, I recommend you default to including a file in your Git repository, unless there's a specific reason not to. But good reasons absolutely do exist.
+In summary, I recommend you default to including a file in your Git
+repository, unless there's a specific reason not to. But good reasons
+absolutely do exist.
 
-## Collaboration
+\subsection{Collaboration}\label{collaboration}
 
-Collaboration is probably the most compelling reason to manage a project with Git and GitHub.
-I have a broad definition of collaboration here, that includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
+Collaboration is probably the most compelling reason to manage a project
+with Git and GitHub. I have a broad definition of collaboration here,
+that includes hands-on participation by multiple people as well as an
+asymmetric model, in which some people are active makers and others only
+read or review.
 
@@
 
-  * **Edit, save, attach.** In this workflow, everyone has one (or more!) copies of the document and they circulate via email attachment, accumulating initials and dates in the filename.
-Which one is "master"?
-Does this question even make sense anymore?
-How do different versions of the document relate to each other?
-If you want to see a version combining the best versions of each section, how would you reconcile the different copies into one?
-All of this usually gets sorted out by social contract, a fairly manual process, and at least one miserable person.
+\begin{itemize}
+\item
+  \textbf{Edit, save, attach.} In this workflow, everyone has one (or
+  more!) copies of the document and they circulate via email attachment,
+  accumulating initials and dates in the filename. Which one is
+  ``master''? Does this question even make sense anymore? How do
+  different versions of the document relate to each other? If you want
+  to see a version combining the best versions of each section, how
+  would you reconcile the different copies into one? All of this usually
+  gets sorted out by social contract, a fairly manual process, and at
+  least one miserable person.
+\item
+  \textbf{Google Doc.} In this workflow, there is only one copy of the
+  document and it lives in the cloud. Anyone can access the most recent
+  version on demand. Anyone can edit or comment or propose a change and
+  this is immediately available to everyone else. Anyone can see who's
+  been editing the document and how and, if disaster strikes, can revert
+  to a previous version. A great deal of ambiguity and annoying
+  reconciliation work has been designed away.
+\end{itemize}
 
-  * **Google Doc.** In this workflow, there is only one copy of the document and it lives in the cloud.
-Anyone can access the most recent version on demand.
-Anyone can edit or comment or propose a change and this is immediately available to everyone else.
-Anyone can see who's been editing the document and how and, if disaster strikes, can revert to a previous version.
-A great deal of ambiguity and annoying reconciliation work has been designed away.
-
-Managing a project via Git/GitHub is much more like the Google Doc scenario, but also offers some of the attractive features of "edit, save, attach".
-With Git/GitHub, collaborators can work offline and there can be independent lines of development.
-The real power comes from regular and structured reconciliation of all the files in the project via Git/GitHub.
-It is definitely more complicated than collaborating on a Google Doc, but also more powerful.
+Managing a project via Git/GitHub is much more like the Google Doc
+scenario, but also offers some of the attractive features of ``edit,
+save, attach''. With Git/GitHub, collaborators can work offline and
+there can be independent lines of development. The real power comes from
+regular and structured reconciliation of all the files in the project
+via Git/GitHub. It is definitely more complicated than collaborating on
+a Google Doc, but also more powerful.
 
@@
 
-Git is a decentralized version control system, meaning each collaborator has their own complete copy of the repo and its history.
-Everyone can work offline and/or simultaneously, but with regular syncing to GitHub.
+Git is a decentralized version control system, meaning each collaborator
+has their own complete copy of the repo and its history. Everyone can
+work offline and/or simultaneously, but with regular syncing to GitHub.
 GitHub plays the role of another collaborator, but a very special one.
-By convention, everyone agrees that GitHub keeps the master copy of the project.
-GitHub is the clearinghouse.
-The joke is that GitHub puts the "central" in decentralized version control.
-You pull regularly from GitHub, to receive and integrate changes made by your collaborators.
-You also push regularly to GitHub, to return the favor, and to maintain its status as the comprehensive, authoritative version of the project.
+By convention, everyone agrees that GitHub keeps the master copy of the
+project. GitHub is the clearinghouse. The joke is that GitHub puts the
+``central'' in decentralized version control. You pull regularly from
+GitHub, to receive and integrate changes made by your collaborators. You
+also push regularly to GitHub, to return the favor, and to maintain its
+status as the comprehensive, authoritative version of the project.
 
-*These two points were originally written as comparisons with Google Docs, but now there are intervening paragraphs. Is it too disconnected? Is it worth weaving in?*
+\emph{These two points were originally written as comparisons with
+Google Docs, but now there are intervening paragraphs. Is it too
+disconnected? Is it worth weaving in?}
 
-*Manage multiple files*.
-A Git repository is inherently multi-file and therefore well suited to projects comprised of many files, evolving in a coordinated fashion.
-Examples include a data analysis, a course website, a blog, an R package, or a book.
-If there is any way to proactively check or enforce their joint functionality, this is something you could verify manually prior to a commit or at certain milestones.
-In the case of a website, you might choose to rebuild the site prior to a commit.
-In the case of an R package, `R CMD check` is one of the easiest things to automate on GitHub.
+\emph{Manage multiple files}. A Git repository is inherently multi-file
+and therefore well suited to projects comprised of many files, evolving
+in a coordinated fashion. Examples include a data analysis, a course
+website, a blog, an R package, or a book. If there is any way to
+proactively check or enforce their joint functionality, this is
+something you could verify manually prior to a commit or at certain
+milestones. In the case of a website, you might choose to rebuild the
+site prior to a commit. In the case of an R package,
+\texttt{R\ CMD\ check} is one of the easiest things to automate on
+GitHub.
 
-*Diffs and time travel*.
-Google Docs are fantastic for simple collaborative work, when you don't need detailed access to the history.
-But the version control offered by Google Drive is very limited compared to Git.
-You can't compare versions at arbitrary points in time, temporarily checkout previous versions, or maintain two lines of development.
+\emph{Diffs and time travel}. Google Docs are fantastic for simple
+collaborative work, when you don't need detailed access to the history.
+But the version control offered by Google Drive is very limited compared
+to Git. You can't compare versions at arbitrary points in time,
+temporarily checkout previous versions, or maintain two lines of
+development.
 
-*May need a segue ... depends on the fate of the two paragraphs above.*
+\emph{May need a segue \ldots{} depends on the fate of the two
+paragraphs above.}
 
-Merge conflicts are the most frustrating thing about using Git and GitHub.
-You can avoid them if you only work alone, on one computer, but I've also said that collaboration is the best reason to use GitHub!
-So this problem must be confronted.
+Merge conflicts are the most frustrating thing about using Git and
+GitHub. You can avoid them if you only work alone, on one computer, but
+I've also said that collaboration is the best reason to use GitHub! So
+this problem must be confronted.
 
-What is a merge conflict?
-Here is a typical first encounter: your collaborator makes a change to a file, commits it locally, and pushes to GitHub.
-Meanwhile you also make a different change to the same file and also commit locally.
-When you try to push your commit to GitHub, it will fail because there are commits on GitHub that you do not have.
-You must pull from GitHub.
-The good news is that quite often, this will "just work", i.e.
-the GitHub version and your version will merge cleanly.
-Git is quite clever at reconciliation and changes to different files or even distinct parts of the same file will merge.
-If this pull and merge goes smoothly, you'll able to push your changes and the cycle goes on.
+What is a merge conflict? Here is a typical first encounter: your
+collaborator makes a change to a file, commits it locally, and pushes to
+GitHub. Meanwhile you also make a different change to the same file and
+also commit locally. When you try to push your commit to GitHub, it will
+fail because there are commits on GitHub that you do not have. You must
+pull from GitHub. The good news is that quite often, this will ``just
+work'', i.e.~the GitHub version and your version will merge cleanly. Git
+is quite clever at reconciliation and changes to different files or even
+distinct parts of the same file will merge. If this pull and merge goes
+smoothly, you'll able to push your changes and the cycle goes on.
 
-But sometimes it's not clear how to reconcile your changes with the new ones from GitHub and you get a *merge conflict*.
-You must inspect the locations of conflict, which Git marks for you.
-You will pick one version or the other -- or create a hybrid -- at each location of conflict and mark it as resolved.
-Once you've resolved all conflicts, you will be able to push a version integrating your recent changes to GitHub.
+But sometimes it's not clear how to reconcile your changes with the new
+ones from GitHub and you get a \emph{merge conflict}. You must inspect
+the locations of conflict, which Git marks for you. You will pick one
+version or the other -- or create a hybrid -- at each location of
+conflict and mark it as resolved. Once you've resolved all conflicts,
+you will be able to push a version integrating your recent changes to
+GitHub.
 
@@
 This is another reason for all parties to commit, pull, and push often.
-Small changes, being integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you.
-The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work.
-The presence of frequently-changing binary files also increases the burden.
-So make lots of small commits, sync regularly with GitHub, and only track binary files with good reason.
+Small changes, being integrated frequently, in non-binary files, are the
+easiest for Git to automatically merge for you. The difficulty of
+merging (by Git or by you) is proportional to the evolutionary distance
+between two lines of work. The presence of frequently-changing binary
+files also increases the burden. So make lots of small commits, sync
+regularly with GitHub, and only track binary files with good reason.
 
-*once I understand Nick's point about binary files causing merge conflicts <https://github.com/dsscollection/git-github-for-stats/issues/6> I may need to reword the above*
+\emph{once I understand Nick's point about binary files causing merge
+conflicts
+\url{https://github.com/dsscollection/git-github-for-stats/issues/6} I
+may need to reword the above}
 
-*something I would like to write about, but have no space for here: different models of collaboration, i.e.
-everyone working on one repo & one branch vs.
-making pull requests from branches in the main repo or from forks*
+\emph{something I would like to write about, but have no space for here:
+different models of collaboration, i.e.~everyone working on one repo \&
+one branch vs.~making pull requests from branches in the main repo or
+from forks}
 
-*inspiration for possible figures on collaboration/merging and the role of a remote, from the barlett talk*
+\emph{inspiration for possible figures on collaboration/merging and the
+role of a remote, from the barlett talk}
 
-![](bartlett-merge-commit.png)
+\begin{figure}
+\centering
+\includegraphics{bartlett-merge-commit.png}
+\caption{}
+\end{figure}
 
-![](bartlett-pull.png)
+\begin{figure}
+\centering
+\includegraphics{bartlett-pull.png}
+\caption{}
+\end{figure}
 
-## GitHub as course management system
+\subsection{GitHub as course management
+system}\label{github-as-course-management-system}
 
-[STAT 545](http://stat545.com) is a data wrangling and analysis course at the University of British Columbia.
-I was the instructor in charge for several years, which coincided with my own adoption of Git/GitHub.
-GitHub is used to manage the development of course material, to serve the course website, to create a discussion forum, and to host all student-submitted work.
+\href{http://stat545.com}{STAT 545} is a data wrangling and analysis
+course at the University of British Columbia. I was the instructor in
+charge for several years, which coincided with my own adoption of
+Git/GitHub. GitHub is used to manage the development of course material,
+to serve the course website, to create a discussion forum, and to host
+all student-submitted work.
 
-Given that students must submit their work and provide peer review of others' work via GitHub, the use of hosted version control is an explicit, though modest, part of the course.
-The website [Happy Git and GitHub for the useR](http://happygitwithr.com) holds our battle-tested instructions for setup and early usage.
-The students achieve basic competence quite quickly and find it gratifying to see their formatted, figure-rich R Markdown reports up on the internet.
-Since it’s easy to expose their work within the class, we conduct peer review, which helps expertise to spread quickly through the group.
+Given that students must submit their work and provide peer review of
+others' work via GitHub, the use of hosted version control is an
+explicit, though modest, part of the course. The website
+\href{http://happygitwithr.com}{Happy Git and GitHub for the useR} holds
+our battle-tested instructions for setup and early usage. The students
+achieve basic competence quite quickly and find it gratifying to see
+their formatted, figure-rich R Markdown reports up on the internet.
+Since it's easy to expose their work within the class, we conduct peer
+review, which helps expertise to spread quickly through the group.
 
-### Use a GitHub Organization
+\subsubsection{Use a GitHub
+Organization}\label{use-a-github-organization}
 
-[GitHub Organizations](https://help.github.com/articles/differences-between-user-and-organization-accounts/) are "shared accounts where groups of people can collaborate across many projects at once".
-This is the most appropriate structure for stewarding course resources, since I can grant TAs and students different levels of access to various repositories.
-Access can be controlled at the individual user level or, more conveniently, for entire [Teams](https://help.github.com/articles/setting-up-teams/).
-The TA Team shares write access with me on a private repository for internal matters.
-I provide each student with their own private repository for coursework and grant other members of the Students Team read access, in order to facilitate peer review.
-There is a public repository that underpins the course website (see below).
-We have one other public repository that exists solely so the [Issues](https://github.com/STAT545-UBC/Discussion/issues) can be used as a discussion forum.
+\href{https://help.github.com/articles/differences-between-user-and-organization-accounts/}{GitHub
+Organizations} are ``shared accounts where groups of people can
+collaborate across many projects at once''. This is the most appropriate
+structure for stewarding course resources, since I can grant TAs and
+students different levels of access to various repositories. Access can
+be controlled at the individual user level or, more conveniently, for
+entire \href{https://help.github.com/articles/setting-up-teams/}{Teams}.
+The TA Team shares write access with me on a private repository for
+internal matters. I provide each student with their own private
+repository for coursework and grant other members of the Students Team
+read access, in order to facilitate peer review. There is a public
+repository that underpins the course website (see below). We have one
+other public repository that exists solely so the
+\href{https://github.com/STAT545-UBC/Discussion/issues}{Issues} can be
+used as a discussion forum.
 
-GitHub actively encourages the use of its platform in teaching.
-As an instructor you can request a [free Organization account](https://help.github.com/articles/discounted-organization-accounts/) that provides features normally available only on paid plans, such as private repositories.
-In fact, GitHub provides tooling for specific teaching workflows via [GitHub Classroom](https://classroom.github.com/), although I do not use it.
-That is not an intentional knock on their tools.
-I started teaching with GitHub several years before this existed and developed a different way of using the platform.
-I also find the [GitHub Education](https://education.github.com) resources to be geared more towards computer science than data science.
+GitHub actively encourages the use of its platform in teaching. As an
+instructor you can request a
+\href{https://help.github.com/articles/discounted-organization-accounts/}{free
+Organization account} that provides features normally available only on
+paid plans, such as private repositories. In fact, GitHub provides
+tooling for specific teaching workflows via
+\href{https://classroom.github.com/}{GitHub Classroom}, although I do
+not use it. That is not an intentional knock on their tools. I started
+teaching with GitHub several years before this existed and developed a
+different way of using the platform. I also find the
+\href{https://education.github.com}{GitHub Education} resources to be
+geared more towards computer science than data science.
 
-### GitHub Pages for course website
+\subsubsection{GitHub Pages for course
+website}\label{github-pages-for-course-website}
 
-All course content is provided on the [STAT 545 website](http://stat545.com).
-Each page is generated from an R Markdown document that is rendered to HTML locally using the rmarkdown package, retaining the intermediate Markdown.
-These pages are a mix of prose and rendered R code, reflecting the live coding done in class.
-All of these files and their history can be explored in the [source repository](https://github.com/STAT545-UBC/STAT545-UBC.github.io).
-The TA team has permission to write to this repo, meaning they can (and do!) help me maintain the website.
-I rejoice that I am no longer the webmaster.
-We also get typo corrections and other input from the world at large, since this is entirely public.
+All course content is provided on the \href{http://stat545.com}{STAT 545
+website}. Each page is generated from an R Markdown document that is
+rendered to HTML locally using the rmarkdown package, retaining the
+intermediate Markdown. These pages are a mix of prose and rendered R
+code, reflecting the live coding done in class. All of these files and
+their history can be explored in the
+\href{https://github.com/STAT545-UBC/STAT545-UBC.github.io}{source
+repository}. The TA team has permission to write to this repo, meaning
+they can (and do!) help me maintain the website. I rejoice that I am no
+longer the webmaster. We also get typo corrections and other input from
+the world at large, since this is entirely public.
 
-If I were starting from scratch today, I would continue to use R Markdown, RStudio, and GitHub Pages (see below), but would upgrade to a more modern, automated approach to rendering the pages.
-I now recommend [R Markdown websites](http://rmarkdown.rstudio.com/rmarkdown_websites.html), [bookdown](https://bookdown.org), or [blogdown](https://bookdown.org/yihui/blogdown/) to manage the process of creating a static website from a large and inter-related set of `.Rmd` files.
+If I were starting from scratch today, I would continue to use R
+Markdown, RStudio, and GitHub Pages (see below), but would upgrade to a
+more modern, automated approach to rendering the pages. I now recommend
+\href{http://rmarkdown.rstudio.com/rmarkdown_websites.html}{R Markdown
+websites}, \href{https://bookdown.org}{bookdown}, or
+\href{https://bookdown.org/yihui/blogdown/}{blogdown} to manage the
+process of creating a static website from a large and inter-related set
+of \texttt{.Rmd} files.
 
-GitHub offers several ways to host a website directly from a repository, collectively known as [GitHub Pages](https://help.github.com/categories/github-pages-basics/).
-The STAT 545 website is a very simple [Organization Page](https://help.github.com/articles/user-organization-and-project-pages/) that uses a [custom domain](https://help.github.com/articles/custom-domain-redirects-for-github-pages-sites/), `stat545.com`, instead of the default `orgname.github.io`.
+GitHub offers several ways to host a website directly from a repository,
+collectively known as
+\href{https://help.github.com/categories/github-pages-basics/}{GitHub
+Pages}. The STAT 545 website is a very simple
+\href{https://help.github.com/articles/user-organization-and-project-pages/}{Organization
+Page} that uses a
+\href{https://help.github.com/articles/custom-domain-redirects-for-github-pages-sites/}{custom
+domain}, \texttt{stat545.com}, instead of the default
+\texttt{orgname.github.io}.
 
-This system for managing course content is a great example of integrating the doing of work and the sharing of it.
-We analyze data live in class, using R, based on the scripts on the website.
-I re-render the associated `.R` or `.Rmd`, commit the changed files, push, and see it reflected right away on http://stat545.com.
-There is no separation between having an idea, implementing it, and posting on the website.
+This system for managing course content is a great example of
+integrating the doing of work and the sharing of it. We analyze data
+live in class, using R, based on the scripts on the website. I re-render
+the associated \texttt{.R} or \texttt{.Rmd}, commit the changed files,
+push, and see it reflected right away on \url{http://stat545.com}. There
+is no separation between having an idea, implementing it, and posting on
+the website.
 
-### Student-specific private repos
+\subsubsection{Student-specific private
+repos}\label{student-specific-private-repos}
 
-Early in the course I elicit GitHub usernames for registered students, via a [Shiny app](https://shiny.rstudio.com), and invite them to join the course Organization.
-I then create one private repository per student, in the STAT 545 Organization.
-The targeted student has write access and the other students have read access.
-This is somewhat controversial, due to the possibility of cheating, but I have seen more pros than cons for this setup, in the STAT 545 context.
-In other settings, I have also used one repo per student *per homework assignment*, which allows you to keep the repos completely private until homework submission, then increase their visibility during marking and peer review.
-Some courses will work better with one model or the other.
+Early in the course I elicit GitHub usernames for registered students,
+via a \href{https://shiny.rstudio.com}{Shiny app}, and invite them to
+join the course Organization. I then create one private repository per
+student, in the STAT 545 Organization. The targeted student has write
+access and the other students have read access. This is somewhat
+controversial, due to the possibility of cheating, but I have seen more
+pros than cons for this setup, in the STAT 545 context. In other
+settings, I have also used one repo per student \emph{per homework
+assignment}, which allows you to keep the repos completely private until
+homework submission, then increase their visibility during marking and
+peer review. Some courses will work better with one model or the other.
 
-Each student does their work in this repo, submitting a major assignment approximately once a week.
-The first assignment is simply to claim the repository and create a README, which proves they have all the relevant software setup and they can write a little Markdown.
-Each week we tackle some new data analysis or wrangling task, with increasing latitude for independence.
-Homework is implemented in R Markdown documents, rendered to Markdown, and pushed to GitHub.
-Students submit their work by opening an issue in their repo, naming the assignment in the title, providing the SHA of the associated final commit, and linking to the main `.md` file.
-We leave feedback as comments in the issue thread or, occasionally, propose changes to code via "pull requests".
-Two peers are selected at random to review each assignment, a process that we also implement via GitHub Issues.
+Each student does their work in this repo, submitting a major assignment
+approximately once a week. The first assignment is simply to claim the
+repository and create a README, which proves they have all the relevant
+software setup and they can write a little Markdown. Each week we tackle
+some new data analysis or wrangling task, with increasing latitude for
+independence. Homework is implemented in R Markdown documents, rendered
+to Markdown, and pushed to GitHub. Students submit their work by opening
+an issue in their repo, naming the assignment in the title, providing
+the SHA of the associated final commit, and linking to the main
+\texttt{.md} file. We leave feedback as comments in the issue thread or,
+occasionally, propose changes to code via ``pull requests''. Two peers
+are selected at random to review each assignment, a process that we also
+implement via GitHub Issues.
 
-At the end of term, the student (and their instructor!) can visit the repo to find an organized, navigable sequence of ~10 assignments.
-Each student leaves with self-written documentation of everything they've done, ready to consult in future projects.
-The last assignments require writing an R package or Shiny app, which they generally do in public repositories under their own accounts.
-They finish STAT 545 with several months of Git/GitHub experience and the start of a data science portfolio.
+At the end of term, the student (and their instructor!) can visit the
+repo to find an organized, navigable sequence of \textasciitilde{}10
+assignments. Each student leaves with self-written documentation of
+everything they've done, ready to consult in future projects. The last
+assignments require writing an R package or Shiny app, which they
+generally do in public repositories under their own accounts. They
+finish STAT 545 with several months of Git/GitHub experience and the
+start of a data science portfolio.
 
-## GitHub as web presence
+\subsection{GitHub as web presence}\label{github-as-web-presence}
 
-Simply having a project on GitHub gives it a web presence!
-Non-users of Git/GitHub can visit the project in the browser and interact with it like a webpage.
-They can grab a snapshot of all the files as a ZIP archive by simply clicking a button.
-People with GitHub accounts have even more options.
-They can clone or fork the repository to get their own copy, which also makes it easy for them to stay current on future changes.
+Simply having a project on GitHub gives it a web presence! Non-users of
+Git/GitHub can visit the project in the browser and interact with it
+like a webpage. They can grab a snapshot of all the files as a ZIP
+archive by simply clicking a button. People with GitHub accounts have
+even more options. They can clone or fork the repository to get their
+own copy, which also makes it easy for them to stay current on future
+changes.
 
-As described above, GitHub Pages offer various ways to serve proper websites, even quite sophisticated ones, directly out of a GitHub repo.
-But before you even worry about that, certain practices can make a GitHub repository much more browsable.
-For many projects, this is more than sufficient for granting people access to your work.
+As described above, GitHub Pages offer various ways to serve proper
+websites, even quite sophisticated ones, directly out of a GitHub repo.
+But before you even worry about that, certain practices can make a
+GitHub repository much more browsable. For many projects, this is more
+than sufficient for granting people access to your work.
 
-**Be savvy about file formats.** Keep files in the plainest, web-friendliest form that is compatible with your main goals.
-As explained above, Markdown is the ideal format for prose, because it is just plain text with some markup, but will be displayed like HTML on GitHub.
-Files named `README.md` are extra special, acting as the index or landing page for their host directory.
-[CSV and TSV](https://help.github.com/articles/rendering-csv-and-tsv-data/) files also get special treatment, including an attractive grid layout and search.
-GitHub has excellent support for displaying and diffing common image formats.
+\textbf{Be savvy about file formats.} Keep files in the plainest,
+web-friendliest form that is compatible with your main goals. As
+explained above, Markdown is the ideal format for prose, because it is
+just plain text with some markup, but will be displayed like HTML on
+GitHub. Files named \texttt{README.md} are extra special, acting as the
+index or landing page for their host directory.
+\href{https://help.github.com/articles/rendering-csv-and-tsv-data/}{CSV
+and TSV} files also get special treatment, including an attractive grid
+layout and search. GitHub has excellent support for displaying and
+diffing common image formats.
 
-**Use conventional file extensions.** GitHub is very code-focused and will apply proper syntax high-lighting for almost any language you can think of, if you use one of the standard file extensions.
-This also has advantages for people searching GitHub and trying to filter by language.
+\textbf{Use conventional file extensions.} GitHub is very code-focused
+and will apply proper syntax high-lighting for almost any language you
+can think of, if you use one of the standard file extensions. This also
+has advantages for people searching GitHub and trying to filter by
+language.
 
-**Use internal links.** `README.md` is a great place to explain how your project fits together.
-Any Markdown file can include relative links to other files in the repo.
-Embedded images are also displayed.
-For figures produced by R code, these links are part of what rmarkdown takes care of for you, but there's no reason you can't do the same yourself for any Markdown file.
+\textbf{Use internal links.} \texttt{README.md} is a great place to
+explain how your project fits together. Any Markdown file can include
+relative links to other files in the repo. Embedded images are also
+displayed. For figures produced by R code, these links are part of what
+rmarkdown takes care of for you, but there's no reason you can't do the
+same yourself for any Markdown file.
 
-*link to http://happygitwithr.com/repo-browsability.html*
+*link to \url{http://happygitwithr.com/repo-browsability.html*}
 
-## More resources
+\subsection{More resources}\label{more-resources}
 
-I've tried to convey the main points about the use of Git and GitHub in statistical and data analytical settings, but I've had to leave many things out.
-There are more advanced topics that will come up once your use of Git becomes more sophisticated and there are topics that are only relevant to certain types of reader.
+I've tried to convey the main points about the use of Git and GitHub in
+statistical and data analytical settings, but I've had to leave many
+things out. There are more advanced topics that will come up once your
+use of Git becomes more sophisticated and there are topics that are only
+relevant to certain types of reader.
 
-I targeted [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity.
-However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms.
-I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today.
-I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.
-For example, we host our own instance of GitHub Enterprise at UBC to support our Master of Data Science program
+I targeted \href{https://github.com}{GitHub} -- not
+\href{https://bitbucket.org}{Bitbucket} or
+\href{https://about.gitlab.com}{GitLab} -- for the sake of specificity.
+However, all the big-picture principles and even some mechanics will
+carry over to these alternative hosting platforms. I am advocating for
+the use of hosted version control as a general concept, with GitHub
+being the best and most common provider today. I note that many
+companies and even universities are starting to make GitHub Enterprise
+or GitLab available internally. For example, we host our own instance of
+GitHub Enterprise at UBC to support our Master of Data Science program
 
-Don't fret too much about public versus private repositories at this point.
-All the major hosting providers offer private repositories with flexible control over who can read or write to the repo.
-There are many ways to get private repositories for low or no cost, especially for academics.
-If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem.
-You can either pay for a higher level of service, self-host one of these platforms, or advocate for organization-wide solutions.
+Don't fret too much about public versus private repositories at this
+point. All the major hosting providers offer private repositories with
+flexible control over who can read or write to the repo. There are many
+ways to get private repositories for low or no cost, especially for
+academics. If you outgrow this initial arrangement, you can throw some
+combination of technical savvy and money at the problem. You can either
+pay for a higher level of service, self-host one of these platforms, or
+advocate for organization-wide solutions.
 
-Branches and pull requests are an extremely powerful feature of Git/GitHub and should be your first foray beyond the basics of commit, push, and pull.
-A branch is an independent line of development within a repo, with the notion it will eventually be merged back into the main branch, a.k.a.
-"master".
-In a course website, you might work in a branch to update a series of lessons, while leaving the current version intact in the meantime.
-A pull request is a specific and formal GitHub process for merging one branch into another.
-You can make a pull request between two branches in the same repo or between two copies of a repo.
-This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages *link to a search?*.
+Branches and pull requests are an extremely powerful feature of
+Git/GitHub and should be your first foray beyond the basics of commit,
+push, and pull. A branch is an independent line of development within a
+repo, with the notion it will eventually be merged back into the main
+branch, a.k.a. ``master''. In a course website, you might work in a
+branch to update a series of lessons, while leaving the current version
+intact in the meantime. A pull request is a specific and formal GitHub
+process for merging one branch into another. You can make a pull request
+between two branches in the same repo or between two copies of a repo.
+This is the mechanism for making and accepting contributions to open
+source software projects on GitHub, including many popular R packages
+\emph{link to a search?}.
 
-## Conclusion
+\subsection{Conclusion}\label{conclusion}
 
-*NEEDS ONE ... BUT SHORT ... SOME FODDER*
+\emph{NEEDS ONE \ldots{} BUT SHORT \ldots{} SOME FODDER}
 
-Transparency about process and product is increasingly important in science.
-The SOMETHING for reproducibility is well accepted.
-A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.
+Transparency about \citep{greenwade93} process and product is
+increasingly important in science. The SOMETHING for reproducibility is
+well accepted. A more underappreciated benefit is democratization of our
+field, as this affords a much broader audience a clear view of how
+scientists and programmers work.
 
-## Random things lying around
+\subsection{Random things lying
+around}\label{random-things-lying-around}
 
-Does creation and presence in GitHub repo make the project more readable and navigable? If so DO IT.
+Does creation and presence in GitHub repo make the project more readable
+and navigable? If so DO IT.
 
@@
 possible github explainer
-https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/
+\url{https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/}
 
-http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst
+\url{http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst}
 
-https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem
+\url{https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem}
 
-https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3
+\url{https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3}
 
@@
 
-  * If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub.
-You can watch the repository to get notified of major activity.
-You can fork it to keep your own copy.
-You can modify your fork to add features or fix bugs and send them back to the owner as a proposed change.
+\begin{itemize}
+\tightlist
+\item
+  If you care deeply about someone else's project, such as an R package
+  you use heavily, you can track its development on GitHub. You can
+  watch the repository to get notified of major activity. You can fork
+  it to keep your own copy. You can modify your fork to add features or
+  fix bugs and send them back to the owner as a proposed change.
+\end{itemize}
 
-The read-only mirrors of R source and all of CRAN. Coupled with GitHub search features, you can answer alot of your own questions this way.
+The read-only mirrors of R source and all of CRAN. Coupled with GitHub
+search features, you can answer alot of your own questions this way.
 
-[RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).
+\href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio IDE}.
+
+\bibliographystyle{agsm}
+\bibliography{git-github-for-stats.bib}
+
+\end{document}

### 2017-05-19 README.Rmd
Revisions, citations, links

@@
 
-Why would a statistician use version control, such as Git? And what is the point of hosting your work online, e.g., on GitHub? Could the gains possibly justify the inevitable pain?
+Why would a statistician use version control, such as Git? And what is the point
+of hosting your work online, e.g., on GitHub? Could the gains possibly justify 
+the inevitable pain?
 
@@
 
-There are many benefits of using hosted version control in your statistical practice:
+There are many benefits of using hosted version control in your statistical 
+practice:
 
@@
 
-[Git](http://git-scm.com) is a __version control system__.
-Its original purpose was to help groups of developers work collaboratively on big software projects.
-Git manages the evolution of a set of files -- called a __repository__ or __repo__ -- in a sane, highly structured way.
-It is a bit like the "Track Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up to multiple files.
+[Git](http://git-scm.com) is a __version control system__. Its original purpose 
+was to help groups of developers work collaboratively on big software projects. 
+Git manages the evolution of a set of files -- called a __repository__ or 
+__repo__ -- in a sane, highly structured way. It is a bit like the "Track 
+Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up
+to multiple files.
 
-Git has been re-purposed by the data science community.
-In addition to using it for source code, we use it to manage the motley collection of files that make up typical data analytical projects, which often consist of data, figures, reports, and, yes, source code.
-Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.
+Git has been re-purposed by the data science community [@Ram2013;
+@git-for-humans]. In addition to using it for source code, we use it to manage 
+the motley collection of files that make up typical data analytical projects, 
+which often consist of data, figures, reports, and, yes, source code. Even those
+who identify more as statistician than data scientist generally have a similar 
+mix of files that are the artifacts of a project.
 
-A lone ranger, working on a single computer, can benefit from adopting version control.
-But not nearly enough to justify the pain of installation and workflow upheaval.
-There are much easier ways to get versioned back ups of files, if that's all you're worried about.
+A lone ranger, working on a single computer, can benefit from adopting version 
+control. But not nearly enough to justify the pain of installation and workflow 
+upheaval. There are much easier ways to get versioned back ups of files, if 
+that's all you're worried about.
 
-In my opinion, **for new users**, the pros of Git only outweigh the cons when you factor in the overhead of communicating and collaborating with other people, including your future self.
-And who among us does not need to do that?
-Your life is much easier if this is baked into your workflow, as opposed to being a separate process that you dread and avoid.
-Communication and collaboration are the killer apps of version control.
-Git enforces a rigorous model of file management, but it is critical for the distribution of files across different people, computers, and time.
+In my opinion, __for new users__, the pros of Git only outweigh the cons when 
+you factor in the overhead of communicating and collaborating with other people,
+including your future self. And who among us does not need to do that? Your life
+is much easier if this is baked into your workflow, as opposed to being a 
+separate process that you dread and avoid. Communication and collaboration are 
+the killer apps of version control. Git enforces a rigorous model of file 
+management, but it is critical for the distribution of files across different 
+people, computers, and time.
 
-This has an implication for selecting your first Git projects: in order to see substantial gain for your pain, you need to pick projects that require you to share rapidly evolving files with others.
-It is tempting to pick a quiet little project that you work on privately, but you risk missing the main point of formal version control.
+This has an implication for selecting your first Git projects: in order to see 
+substantial gain for your pain, you need to pick projects that require you to 
+share rapidly evolving files with others. It is tempting to pick a quiet little 
+project that you work on privately, but you risk missing the main point of 
+formal version control.
 
 Many people who don't use Git unwittingly re-invent a poor man's version of it. 
-They take an important file and distribute it via email.
-Various parties make changes, decorating the file name with initials, dates, and other descriptors.
-Before you know it, the original file is the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve.
+They take an important file and distribute it via email. Various parties make 
+changes, decorating the file name with initials, dates, and other descriptors. 
+Before you know it, the original file is the root of a complicated phylogeny 
+that no amount of "Track changes" and good intentions can resolve.
 
-*If you're looking at the PDF draft, you're missing out on some figures around here in the GitHub Markdown version that hint at the figure I plan to embed.*
+*If you're looking at the PDF draft, you're missing out on some figures around 
+here in the GitHub Markdown version that hint at the figure I plan to embed.*
 
@@
 
-cat("*make such a phylogeny with a set of crazy filenames like this*\n")
+cat("\n\n*make such a phylogeny with a set of crazy filenames like this*\n") 
 cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")
 
-cat("*or riff on the famous PhD Comic*")
+cat("\n\n*or riff on the famous PhD Comic*") 
 cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
@@
 
-The Git way is to track the evolution of that file, through a series of commits, each equipped with an explanatory message and a nickname.
-All collaborators sync regularly to a common version, acknowledging that the difficulty of merging goes up faster than the size of the difference (cite good enough).
-Especially important versions get a human-readable tag, to signal a meaningful milestone.
-Yes, there is some pain in adopting the formalism of Git, but it is worth it.
+The Git way is to track the evolution of that file, through a series of commits,
+each equipped with an explanatory message and a nickname. All collaborators sync
+regularly to a common version, acknowledging that the difficulty of merging goes
+up faster than the size of the difference [@good-enough]. Especially important 
+versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
+is some pain in adopting the formalism of Git, but it is worth it.
 
-*another good image to emulate, from the Bartlett Git for Designers talk*
-![](bartlett-copies-vs-commits.png)
+*make a figure related to the above (reference those file names?) but 
+juxtaposing different workflows, inspiration from the Bartlett Git for Designers
+talk*
+
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("bartlett-copies-vs-commits.png")
+```
 
@@
 
-The target reader is anyone who does statistical research, analysis, or instruction. Those whose work is some combination of these three may find the work style described here especially rewarding.
+The target reader is anyone who does statistical research, analysis, or 
+instruction. Those whose work is some combination of these three may find the 
+work style described here especially rewarding.
 
-This article does not provide step-by-step instructions on how to use Git and GitHub.
-This format would not be effective, but we do provide annotated links to such resources in ?the appendix or online supplement?.
-Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context.
-The goal is to help the Git-curious generate the activation energy needed to get started.
+This article does not provide step-by-step instructions on how to use Git and 
+GitHub. This format would not be effective, but we do provide annotated links to
+such resources in ?the appendix or online supplement?. Instead, I'll convey what
+the workflow feels like and what the payoffs are, with special attention to the 
+statistics and R context. The goal is to help the Git-curious generate the 
+activation energy needed to get started.
 
@@
 [GitHub](https://github.com) is currently the most popular Git hosting service. 
-Others include [GitLab](https://about.gitlab.com) and [Bitbucket](https://bitbucket.org).
-These services provide a home for Git-based projects on the internet.
-It is a bit like DropBox or Google Drive, but more structured, powerful, and programmatic.
+Others include [GitLab](https://about.gitlab.com) and 
+[Bitbucket](https://bitbucket.org). These services provide a home for Git-based 
+projects on the internet. It is a bit like DropBox or Google Drive, but more 
+structured, powerful, and programmatic.
 
-The remote host acts as the clearinghouse for a Git-managed project.
-This allows other people to see the project files, sync up with the current version, and perhaps even make or propose changes.
-These hosting providers offer well-designed web interfaces that are a dramatic improvement over traditional Unix Git servers.
-Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files.
-These hosts also offer granular control over who can see, edit, and administer a project.
+The remote host acts as the clearinghouse for a Git-managed project. This allows
+other people to see the project files, sync up with the current version, and 
+perhaps even make or propose changes. These hosting providers offer 
+well-designed web interfaces that are a dramatic improvement over traditional 
+Unix Git servers. Many operations can be done entirely in the browser, such as 
+viewing current or past versions of a file, commenting on a recent change, and 
+editing or adding files. These hosts also offer granular control over who can 
+see, edit, and administer a project.
 
-Even for private solo projects, there are two advantages to keeping a synced copy on GitHub:
+*possible images include a modified version of this atlassian image (no arrows 
+between people, yes arrow from people to github), somehow connected to a 
+screenshot of a GitHub repo*
 
-  1.
-When you are new with Git (or, frankly, even when you're not), it's common to goof up the Git infrastructure for a project.
-Note that your files can absolutely be intact and safe, even while the Git tracking is a bit confused.
-Of course there are official Git remedies, but sometimes the easiest fix is to clone a fresh copy from GitHub, patch things up with the changes that only exist locally, and move on with your life.
-This workaround obviously requires the existence of a recent copy on GitHub.
-  2.
-The highly functional web interfaces mentioned above are often the most pleasant and natural way to navigate and search your files, even though all the same information exists locally.
-It is a pleasure to browse through your own work, across multiple projects or files and across time, as if it's a well-designed website.
-You must push your work to GitHub to enjoy this.
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("atlassian-git-tutorials-syncing.png")
+```
 
-GitHub __issues__ are another powerful feature of the platform.
-Recall that we are repurposing Git, a tool designed to facilitate software development.
-The issues for a project are its bug tracker.
-For projects that are not pure software development, we co-opt this machinery to organize our to do list more generally.
-The basic unit is an issue and you can interact with them in two ways.
+Even for private solo projects, there are two advantages to keeping a synced 
+copy on GitHub:
 
-First, issues are integrated into the project's web interface on GitHub, with a rich set of options for linking to project files and incremental changes.
-Second, issues and their associated comment threads appear in your email, just like regular messages (this can, of course, be configured).
-The result is that all correspondence about a project comes through your normal channels, but is also tracked inside the project itself, with excellent navigability and search capabilities.
-For software, issues are used to track bugs and feature requests.
-In a data analysis project, you might open an issue to flesh out a specific sub-analysis or to develop a complicated figure.
-In a course, we use them to manage homework submission, marking, and peer review.
+1. When you are new with Git (or, frankly, even when you're not), it's common to
+goof up the Git infrastructure for a project. Note that your files can be intact
+and safe, even while the Git tracking is a bit confused. Of course there are 
+official Git remedies, but sometimes the easiest fix is to clone a fresh copy 
+from GitHub, patch things up with the changes that only exist locally, and move 
+on with your life. This workaround obviously requires the existence of a recent 
+copy on GitHub.
 
-Issues can be assigned to specific people and they can be labelled, e.g.
-"bug", "simulation-study", or "final-exam".
-Coupled with the ability to cross-link issues and the project files or file changes, you have extraordinary power to document why things have happened in the past and to organize what needs to happen in the future.
+2. The highly functional web interfaces mentioned above are often the most 
+pleasant and natural way to navigate and search your files, even though all the 
+same information exists locally. It is a pleasure to browse through your own 
+work, across multiple projects or files and across time, as if it's a 
+well-designed website. You must push your work to GitHub to enjoy this.
+
+__GitHub issues__ are another powerful feature of the platform. Recall that we 
+are repurposing Git, a tool designed to facilitate software development. The 
+issues for a project are its bug tracker. For projects that are not pure 
+software development, we co-opt this machinery to organize our to do list more 
+generally. The basic unit is an issue and you can interact with one in two ways.
+
+First, issues are integrated into the project's web interface on GitHub, with a 
+rich set of options for linking to project files and incremental changes. 
+Second, issues and their associated comment threads appear in your email, just 
+like regular messages (this can, of course, be configured). The result is that 
+all correspondence about a project comes through your normal channels, but is 
+also tracked inside the project itself, with excellent navigability and search 
+capabilities. For software, issues are used to track bugs and feature requests. 
+In a data analysis project, you might open an issue to flesh out a specific 
+sub-analysis or to develop a complicated figure. In a course, we use them to 
+manage homework submission, marking, and peer review.
+
+Issues can be assigned to specific people and they can be labelled, e.g. "bug", 
+"simulation-study", or "final-exam". Coupled with the ability to cross-link 
+issues and the project files or file changes, you have extraordinary power to 
+document why things have happened in the past and to organize what needs to 
+happen in the future.
 
@@
 
-This is one-time or once-per-computer setup.
+Here is the initial setup process, which you do once or, for some steps, once per computer.
 
@@
 
-  * Install Git.
-Depending on your OS, Git might already be installed.
-But many of us will need to install it or might choose to update to the most recent version.
+* Install Git. Depending on your OS, Git might already be installed. But many of
+us may need to install it or will choose to update to a more recent version.
 Some basic configuration is critical, such as setting your username and email.
 
-  * Install a local Git client.
-Optional but highly recommended.
-A Git client is software that provides a graphical user interface for Git, which is otherwise command-line only.
-If you are an R user, you will find that RStudio provides a great deal of this functionality.
-There are some notable gaps, however, so you might still choose to install a dedicated and comprehensive Git client such as SourceTree or GitKraken.
-Git is a file-based system, so you can do some operations from RStudio, others from SourceTree, and others from the shell.
+* Install a local Git client *(optional but highly recommended*). A Git client 
+is software that provides a graphical user interface for Git, which is otherwise
+command-line only. If you are an R user, you will find that 
+[RStudio](https://www.rstudio.com/products/rstudio-desktop/) provides a great 
+deal of this functionality. There are some notable gaps, however, so you might 
+still choose to install a dedicated and comprehensive Git client such as 
+[SourceTree](https://www.sourcetreeapp.com) or
+[GitKraken](https://www.gitkraken.com). Git is a file-based system, so you can
+do some operations from RStudio, others from SourceTree, and others from the
+shell.
 
-  * Confirm, with a practice repository, that local Git can send and receive the current version of the repository on GitHub, known as *pushing* and *pulling*, respectively.
+* Confirm, with a practice repository, that local Git can send and receive the 
+current version of the repository on GitHub, known as __pushing__ and __pulling__, respectively.
 This will require authenticating yourself with GitHub from a local shell or Git client.
@@
 
-Once this setup is done, you are ready to start using Git and GitHub with your projects.
+Once this setup is done, you are ready to start using Git and GitHub with your 
+projects.
 
@@
 
-Git was built neither for the exact usage described here, nor for broad usability.
-You will undoubtedly notice this, so it can be helpful to know this in advance.
-Happily there are many helpful tools that mediate your interactions with Git.
-GitHub itself is a fine example.
-In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.
+Git was built neither for the exact usage described here, nor for broad 
+usability. You will undoubtedly notice this, so it can be helpful to know this 
+in advance. Happily there are many helpful tools that mediate your interactions 
+with Git. GitHub itself is a fine example, as is RStudio. In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.
 
@@
 
-  * I repeat: consider using a graphical front-end for Git, a.k.a. a "Git client", versus restricting yourself to the command line interface.
+  * I repeat: consider using a graphical front-end for Git, a.k.a. a Git client, versus restricting yourself to the command line interface.
   * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
@@
 
-  * Dedicate a local directory (a.k.a "folder") to it.
-  * Make it an RStudio Project. Optional but recommended; obviously only applies to projects involving R and users of RStudio.
+  * Dedicate a local directory or folder to it.
+  * Make it an RStudio Project. *Optional but recommended; obviously only applies to projects involving R and users of RStudio.*
   * Make it a Git repository.
 
-This is once-per-project setup and can happen at project inception or at any later point.
+This setup happens once per project and can happen at project inception or at any later point.
 Chances are your project already lives in a dedicated directory.
@@
 
-  * Go about your usual business, e.g.
-writing R scripts or authoring reports in LaTeX or R Markdown.
-But instead of only *saving* individual files, periodically you make a **commit**, which takes a snapshot of all the files in the entire project.
-    - Have you ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531)?
-That is effectively a **commit**, albeit only for a single file: it is a version that is significant to you and that you might want to inspect or revert to later.
+  * Go about your usual business, writing R scripts or authoring reports in LaTeX or R Markdown.
+But instead of only *saving* individual files, periodically you make a __commit__, which takes a snapshot of all the files in the entire project.
+    - If you have ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531), you have effectively made a commit, albeit only for a single file.
+It is a version that is significant to you and that you might want to inspect or revert to later.
   * Push commits to GitHub periodically.
@@
 In [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub, starting in week one.
-Most have never seen Git before and do not identify as "programmers".
+Most have never seen Git before and do not identify as programmers.
 It is a major topic in class and office hours for the first two weeks.
 Then we practically never discuss it again.
-
-*I am not saying anything about pulling and collaboration here, which seems a bit extreme, even if I can't say much*
 
@@
 
-Recall that a repository -- or "repo" -- is just a directory of files that Git is going to manage holistically.
+Recall that a repository or repo is just a directory of files that Git will manage holistically.
 A commit functions like a snapshot of all the files in the repo, at a specific moment.
@@
 
-![](bartlett-commit-history.png)
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("bartlett-commit-history.png")
+```
 
@@
 
-Every time you make a commit you must also write a short *commit message*.
-Ideally, this conveys the *motivation* for the change.
+Every time you make a commit you must also write a short __commit message__.
+Ideally, this conveys the motivation for the change.
 Remember, the diff will show the content.
-When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the *history*, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.
+When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the __history__, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.
 
 Every commit needs some sort of nickname, so you can identify it.
-Git does this automatically, assigning each commit what is called a *SHA*, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a SHA-1 checksum hash of the commit).
+Git does this automatically, assigning each commit what is called a SHA, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a SHA-1 checksum hash of the commit).
 Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice.
-You can also designate certain snapshots as special with a *tag*, which is a name of your choosing.
+You can also designate certain snapshots as special with a __tag__, which is a name of your choosing.
 In a software project, it is typical to tag a release with its version, e.g., "v1.0.3".
@@
 
-This may seem unrelated to Git, GitHub, and R, but we need to talk about Markdown.
+This may seem unrelated to Git, GitHub, and R, but we need to talk about [Markdown](https://daringfireball.net/projects/markdown/syntax).
 Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible.
 The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features.
-Markdown is in wide use on sites like Wikipedia, WordPress, StackOverflow, and, yes, GitHub.
+Markdown is in wide use on sites like [WordPress](https://en.support.wordpress.com/markdown/), [StackOverflow](https://stackoverflow.com/editing-help), and, yes, [GitHub](https://help.github.com/categories/writing-on-github/).
 These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting.
@@
 
-Markdown is extra special for R users because of R Markdown, which is just Markdown that includes chunks of R code.
+Markdown is extra special for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code.
 Again, do not regard R Markdown as something you must clear your schedule to learn.
 If you can write email and a bit of R code, you can write R Markdown.
-The rmarkdown package converts R Markdown (`.Rmd` files) to Markdown (`.md` files), running the code and inserting the results, including figures, into the document.
-This is powered by another package, knitr, under the hood.
+The [rmarkdown package](https://CRAN.R-project.org/package=rmarkdown) [@rmd-pkg] converts R Markdown (`.Rmd` files) to Markdown (`.md` files), running the code and inserting the results, including figures, into the document.
+This is powered by another package, [knitr](https://CRAN.R-project.org/package=knitr) [@knitr-pkg; @knitr-book], under the hood.
 This process is made especially easy in RStudio, but is by no means limited to users of that application.
@@
 These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described above.
-Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, **and view the results of running that code**, including figures.
+Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, __and view the results of running that code__, including figures.
 It is the best of all worlds, because the code is revealed and, by definition, is the code that produced the results.
@@
 You do not, in fact, even need to work in R Markdown to exploit this.
-It works with plain R scripts as well.
+It [works with plain R scripts](http://rmarkdown.rstudio.com/articles_report_from_r_script.html) as well.
 You can use exactly the same machinery to prepare a rendered version of an R script, i.e.
@@
 I have emphasized the production of Markdown here because it is extremely useful in GitHub-hosted projects and that seems to be underappreciated.
-But the rmarkdown package can convert `.Rmd` to a wide array of output formats, including HTML, PDF, and Word (`.docx`).
+But the rmarkdown package can convert `.Rmd` to a wide array of [output formats](http://rmarkdown.rstudio.com/lesson-9.html), including HTML, PDF, and Word (`.docx`).
 The Markdown must be created regardless.
@@
   * R script `.R` --> results as `.csv` or `.rds` and figures `.png`
-  * LaTeX `.tex` --> WHAT GOES HERE NOW? --> `.pdf`
+  * LaTeX `.tex` --> `.aux`, `.bbl`, etc. --> `.pdf`
 
@@
 In the case of an R script, this is literally true, but it's morally true for R markdown and LaTeX files too.
-These are files that you directly create and edit "by hand".
+These are files that you directly create and edit by hand.
 
-The files at the far right are clearly derived and are often described as "targets".
+The files at the far right are clearly derived and are often described as __targets__.
 These files are programmatically generated from source files (and possibly other inputs).
@@
 This reduces clutter in your project: Git will not pester you to commit changes to these files and they will not appear in the associated GitHub repository.
-A file that Git does not ignore is said to be *tracked*.
+A file that Git does not ignore is said to be __tracked__.
 
@@
 What are the main considerations when deciding whether to track a derived file or file type?
+
+*I suspect this is a good place to be less wordy? I am so frustrated with people doing stupid things here, though!*
 
@@
 
-*Will it make collaboration harder?* *talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>*
+*Will it make collaboration harder?* Prose to be written: Talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>.
+I have a solution!
+Stop using PDF as your default output format!
+Make Markdown the output format during development.
+Your Git/GitHub problems just went away.
+You're welcome.
 
@@
 Collaboration is probably the most compelling reason to manage a project with Git and GitHub.
-I have a broad definition of collaboration here, that includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
+I have a broad definition of collaboration.
+It includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
 
@@
 With Git/GitHub, collaborators can work offline and there can be independent lines of development.
-The real power comes from regular and structured reconciliation of all the files in the project via Git/GitHub.
+The real power comes from regular and structured reconciliation of all versions of all the files in the project via Git/GitHub.
 It is definitely more complicated than collaborating on a Google Doc, but also more powerful.
@@
 This is another reason for all parties to commit, pull, and push often.
-Small changes, being integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you.
+Small changes, integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you.
 The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work.
@@
 
-![](bartlett-merge-commit.png)
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("bartlett-merge-commit.png")
+```
 
-![](bartlett-pull.png)
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("bartlett-pull.png")
+```
 
@@
 
-Transparency about [@greenwade93] process and product is increasingly important in science.
+Transparency about process and product is increasingly important in science.
 The SOMETHING for reproducibility is well accepted.
@@
 
+make statistical thought and implementation available
+
 ## Random things lying around
 
-Does creation and presence in GitHub repo make the project more readable and navigable? If so DO IT.
-
-need to add a ton of links, citations
-
-make some bespoke examples or diagrams?
-
-Doing = documenting = sharing
-
-make statistical thought and implementation available
+Consider links to GitHub repos that exemplify certain points and are very unlikely to disappear any time soon.n
 
@@
 
-[RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).
-
 ```{r eval = is_github_document, echo = FALSE, results = 'asis'}

### 2017-05-19 README.tex
Revisions, citations, links

@@
 
-Git has been re-purposed by the data science community. In addition to
-using it for source code, we use it to manage the motley collection of
-files that make up typical data analytical projects, which often consist
-of data, figures, reports, and, yes, source code. Even those who
-identify more as statistician than data scientist generally have a
-similar mix of files that are the artifacts of a project.
+Git has been re-purposed by the data science community
+\citep{Ram2013, git-for-humans}. In addition to using it for source
+code, we use it to manage the motley collection of files that make up
+typical data analytical projects, which often consist of data, figures,
+reports, and, yes, source code. Even those who identify more as
+statistician than data scientist generally have a similar mix of files
+that are the artifacts of a project.
 
@@
 difficulty of merging goes up faster than the size of the difference
-(cite good enough). Especially important versions get a human-readable
+\citep{good-enough}. Especially important versions get a human-readable
 tag, to signal a meaningful milestone. Yes, there is some pain in
@@
 
-\emph{another good image to emulate, from the Bartlett Git for Designers
-talk} \includegraphics{bartlett-copies-vs-commits.png}
+\emph{make a figure related to the above (reference those file names?)
+but juxtaposing different workflows, inspiration from the Bartlett Git
+for Designers talk}
+
+\includegraphics[width=1\linewidth]{bartlett-copies-vs-commits}
 
@@
 
+\emph{possible images include a modified version of this atlassian image
+(no arrows between people, yes arrow from people to github), somehow
+connected to a screenshot of a GitHub repo}
+
+\includegraphics[width=1\linewidth]{atlassian-git-tutorials-syncing}
+
 Even for private solo projects, there are two advantages to keeping a
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
@@
   common to goof up the Git infrastructure for a project. Note that your
-  files can absolutely be intact and safe, even while the Git tracking
-  is a bit confused. Of course there are official Git remedies, but
-  sometimes the easiest fix is to clone a fresh copy from GitHub, patch
-  things up with the changes that only exist locally, and move on with
-  your life. This workaround obviously requires the existence of a
-  recent copy on GitHub.
+  files can be intact and safe, even while the Git tracking is a bit
+  confused. Of course there are official Git remedies, but sometimes the
+  easiest fix is to clone a fresh copy from GitHub, patch things up with
+  the changes that only exist locally, and move on with your life. This
+  workaround obviously requires the existence of a recent copy on
+  GitHub.
 \item
@@
 
-GitHub \textbf{issues} are another powerful feature of the platform.
+\textbf{GitHub issues} are another powerful feature of the platform.
 Recall that we are repurposing Git, a tool designed to facilitate
@@
 machinery to organize our to do list more generally. The basic unit is
-an issue and you can interact with them in two ways.
+an issue and you can interact with one in two ways.
 
@@
 
-This is one-time or once-per-computer setup.
+Here is the initial setup process, which you do once or, for some steps,
+once per computer.
 
@@
   Install Git. Depending on your OS, Git might already be installed. But
-  many of us will need to install it or might choose to update to the
-  most recent version. Some basic configuration is critical, such as
-  setting your username and email.
-\item
-  Install a local Git client. Optional but highly recommended. A Git
-  client is software that provides a graphical user interface for Git,
-  which is otherwise command-line only. If you are an R user, you will
-  find that RStudio provides a great deal of this functionality. There
-  are some notable gaps, however, so you might still choose to install a
-  dedicated and comprehensive Git client such as SourceTree or
-  GitKraken. Git is a file-based system, so you can do some operations
-  from RStudio, others from SourceTree, and others from the shell.
+  many of us may need to install it or will choose to update to a more
+  recent version. Some basic configuration is critical, such as setting
+  your username and email.
+\item
+  Install a local Git client \emph{(optional but highly recommended}). A
+  Git client is software that provides a graphical user interface for
+  Git, which is otherwise command-line only. If you are an R user, you
+  will find that
+  \href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio}
+  provides a great deal of this functionality. There are some notable
+  gaps, however, so you might still choose to install a dedicated and
+  comprehensive Git client such as
+  \href{https://www.sourcetreeapp.com}{SourceTree} or
+  \href{https://www.gitkraken.com}{GitKraken}. Git is a file-based
+  system, so you can do some operations from RStudio, others from
+  SourceTree, and others from the shell.
 \item
@@
   receive the current version of the repository on GitHub, known as
-  \emph{pushing} and \emph{pulling}, respectively. This will require
+  \textbf{pushing} and \textbf{pulling}, respectively. This will require
   authenticating yourself with GitHub from a local shell or Git client.
@@
 know this in advance. Happily there are many helpful tools that mediate
-your interactions with Git. GitHub itself is a fine example. In addition
-to pointing out tools that soften Git's sharpest edges, I recommend
-specific habits and attitudes that reduce frustration.
+your interactions with Git. GitHub itself is a fine example, as is
+RStudio. In addition to pointing out tools that soften Git's sharpest
+edges, I recommend specific habits and attitudes that reduce
+frustration.
 
@@
 \item
-  I repeat: consider using a graphical front-end for Git, a.k.a. a ``Git
-  client'', versus restricting yourself to the command line interface.
+  I repeat: consider using a graphical front-end for Git, a.k.a. a Git
+  client, versus restricting yourself to the command line interface.
 \item
@@
 \item
-  Dedicate a local directory (a.k.a ``folder'') to it.
-\item
-  Make it an RStudio Project. Optional but recommended; obviously only
-  applies to projects involving R and users of RStudio.
+  Dedicate a local directory or folder to it.
+\item
+  Make it an RStudio Project. \emph{Optional but recommended; obviously
+  only applies to projects involving R and users of RStudio.}
 \item
@@
 
-This is once-per-project setup and can happen at project inception or at
-any later point. Chances are your project already lives in a dedicated
-directory. Making this directory an RStudio Project and Git repository
-boils down to allowing those applications to leave notes for themselves
-in hidden files or directories. The project is still a regular directory
-on your computer, that you can locate, name, move, and generally
-interact with as you wish. You don't have to handle it with special
-gloves!
+This setup happens once per project and can happen at project inception
+or at any later point. Chances are your project already lives in a
+dedicated directory. Making this directory an RStudio Project and Git
+repository boils down to allowing those applications to leave notes for
+themselves in hidden files or directories. The project is still a
+regular directory on your computer, that you can locate, name, move, and
+generally interact with as you wish. You don't have to handle it with
+special gloves!
 
@@
 \item
-  Go about your usual business, e.g.~writing R scripts or authoring
-  reports in LaTeX or R Markdown. But instead of only \emph{saving}
-  individual files, periodically you make a \textbf{commit}, which takes
-  a snapshot of all the files in the entire project.
+  Go about your usual business, writing R scripts or authoring reports
+  in LaTeX or R Markdown. But instead of only \emph{saving} individual
+  files, periodically you make a \textbf{commit}, which takes a snapshot
+  of all the files in the entire project.
 
@@
   \item
-    Have you ever versioned a file
+    If you have ever versioned a file
     \href{http://www.phdcomics.com/comics/archive.php?comicid=1531}{by
-    adding your initials or the date}? That is effectively a
-    \textbf{commit}, albeit only for a single file: it is a version that
-    is significant to you and that you might want to inspect or revert
-    to later.
+    adding your initials or the date}, you have effectively made a
+    commit, albeit only for a single file. It is a version that is
+    significant to you and that you might want to inspect or revert to
+    later.
   \end{itemize}
@@
 coursework via GitHub, starting in week one. Most have never seen Git
-before and do not identify as ``programmers''. It is a major topic in
-class and office hours for the first two weeks. Then we practically
-never discuss it again.
-
-\emph{I am not saying anything about pulling and collaboration here,
-which seems a bit extreme, even if I can't say much}
+before and do not identify as programmers. It is a major topic in class
+and office hours for the first two weeks. Then we practically never
+discuss it again.
 
@@
 
-Recall that a repository -- or ``repo'' -- is just a directory of files
-that Git is going to manage holistically. A commit functions like a
-snapshot of all the files in the repo, at a specific moment. Under the
-hood, that is not exactly how Git implements things. Mental models don't
-have to be accurate in order to be useful, but in this case there's some
-value in aligning the two.
+Recall that a repository or repo is just a directory of files that Git
+will manage holistically. A commit functions like a snapshot of all the
+files in the repo, at a specific moment. Under the hood, that is not
+exactly how Git implements things. Mental models don't have to be
+accurate in order to be useful, but in this case there's some value in
+aligning the two.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{bartlett-commit-history.png}
-\caption{}
-\end{figure}
+\includegraphics[width=1\linewidth]{bartlett-commit-history}
 
@@
 
-Every time you make a commit you must also write a short \emph{commit
-message}. Ideally, this conveys the \emph{motivation} for the change.
-Remember, the diff will show the content. When you revisit a project
-after a break or need to digest recent changes made by a colleague,
-looking at the \emph{history}, by reading commit messages and skimming
-through diffs, is an extremely efficient way to get up to speed.
+Every time you make a commit you must also write a short \textbf{commit
+message}. Ideally, this conveys the motivation for the change. Remember,
+the diff will show the content. When you revisit a project after a break
+or need to digest recent changes made by a colleague, looking at the
+\textbf{history}, by reading commit messages and skimming through diffs,
+is an extremely efficient way to get up to speed.
 
 Every commit needs some sort of nickname, so you can identify it. Git
-does this automatically, assigning each commit what is called a
-\emph{SHA}, a seemingly random string of 40 letters and numbers (it is
-not, in fact, random but is a SHA-1 checksum hash of the commit). Though
-you will be exposed to these, you don't have to handle them directly
-very often and, when you do, usually the first 7 characters suffice. You
-can also designate certain snapshots as special with a \emph{tag}, which
-is a name of your choosing. In a software project, it is typical to tag
-a release with its version, e.g., ``v1.0.3''. For a manuscript or
+does this automatically, assigning each commit what is called a SHA, a
+seemingly random string of 40 letters and numbers (it is not, in fact,
+random but is a SHA-1 checksum hash of the commit). Though you will be
+exposed to these, you don't have to handle them directly very often and,
+when you do, usually the first 7 characters suffice. You can also
+designate certain snapshots as special with a \textbf{tag}, which is a
+name of your choosing. In a software project, it is typical to tag a
+release with its version, e.g., ``v1.0.3''. For a manuscript or
 analytical project, you might tag the version submitted to a journal or
@@
 This may seem unrelated to Git, GitHub, and R, but we need to talk about
-Markdown. Markdown is a markup language, like HTML and LaTeX, but
-designed to be as lightweight as possible. The goal is still to separate
-form and content, but also to prioritize human-readability, even at the
-cost of fancy features. Markdown is in wide use on sites like Wikipedia,
-WordPress, StackOverflow, and, yes, GitHub. These sites use Markdown
-because it allows a diverse population of site users to author
-decent-looking web content, with hyperlinks and some formatting. Do not
-build this up into some heroic, LaTeX-level learning task, for it is
-not. If you can write an email, you can write Markdown.
+\href{https://daringfireball.net/projects/markdown/syntax}{Markdown}.
+Markdown is a markup language, like HTML and LaTeX, but designed to be
+as lightweight as possible. The goal is still to separate form and
+content, but also to prioritize human-readability, even at the cost of
+fancy features. Markdown is in wide use on sites like
+\href{https://en.support.wordpress.com/markdown/}{WordPress},
+\href{https://stackoverflow.com/editing-help}{StackOverflow}, and, yes,
+\href{https://help.github.com/categories/writing-on-github/}{GitHub}.
+These sites use Markdown because it allows a diverse population of site
+users to author decent-looking web content, with hyperlinks and some
+formatting. Do not build this up into some heroic, LaTeX-level learning
+task, for it is not. If you can write an email, you can write Markdown.
 
@@
 
-Markdown is extra special for R users because of R Markdown, which is
-just Markdown that includes chunks of R code. Again, do not regard R
-Markdown as something you must clear your schedule to learn. If you can
-write email and a bit of R code, you can write R Markdown. The rmarkdown
-package converts R Markdown (\texttt{.Rmd} files) to Markdown
+Markdown is extra special for R users because of
+\href{http://rmarkdown.rstudio.com}{R Markdown}, which is just Markdown
+that includes chunks of R code. Again, do not regard R Markdown as
+something you must clear your schedule to learn. If you can write email
+and a bit of R code, you can write R Markdown. The
+\href{https://CRAN.R-project.org/package=rmarkdown}{rmarkdown package}
+\citep{rmd-pkg} converts R Markdown (\texttt{.Rmd} files) to Markdown
 (\texttt{.md} files), running the code and inserting the results,
 including figures, into the document. This is powered by another
-package, knitr, under the hood. This process is made especially easy in
-RStudio, but is by no means limited to users of that application.
+package, \href{https://CRAN.R-project.org/package=knitr}{knitr}
+\citep{knitr-pkg, knitr-book}, under the hood. This process is made
+especially easy in RStudio, but is by no means limited to users of that
+application.
 
@@
 You do not, in fact, even need to work in R Markdown to exploit this. It
-works with plain R scripts as well. You can use exactly the same
-machinery to prepare a rendered version of an R script, i.e.~to go from
-\texttt{.R} to \texttt{.md}. Again, RStudio makes this especially easy,
-but this is not limited to RStudio. Once the Markdown file is pushed to
-GitHub, it is as if the reader has run your code or is able to look over
-your shoulder at your R session. This provides an lightweight system for
+\href{http://rmarkdown.rstudio.com/articles_report_from_r_script.html}{works
+with plain R scripts} as well. You can use exactly the same machinery to
+prepare a rendered version of an R script, i.e.~to go from \texttt{.R}
+to \texttt{.md}. Again, RStudio makes this especially easy, but this is
+not limited to RStudio. Once the Markdown file is pushed to GitHub, it
+is as if the reader has run your code or is able to look over your
+shoulder at your R session. This provides an lightweight system for
 exposing work-in-progress to collaborators, without slowing down to
@@
 GitHub-hosted projects and that seems to be underappreciated. But the
-rmarkdown package can convert \texttt{.Rmd} to a wide array of output
-formats, including HTML, PDF, and Word (\texttt{.docx}). The Markdown
-must be created regardless. Sometimes it is all you need, such as on
-GitHub. In other contexts, it is just a necessary intermediate and may
-even be discarded.
+rmarkdown package can convert \texttt{.Rmd} to a wide array of
+\href{http://rmarkdown.rstudio.com/lesson-9.html}{output formats},
+including HTML, PDF, and Word (\texttt{.docx}). The Markdown must be
+created regardless. Sometimes it is all you need, such as on GitHub. In
+other contexts, it is just a necessary intermediate and may even be
+discarded.
 
@@
 \item
-  LaTeX \texttt{.tex} --\textgreater{} WHAT GOES HERE NOW?
-  --\textgreater{} \texttt{.pdf}
+  LaTeX \texttt{.tex} --\textgreater{} \texttt{.aux}, \texttt{.bbl},
+  etc. --\textgreater{} \texttt{.pdf}
 \end{itemize}
@@
 script, this is literally true, but it's morally true for R markdown and
-LaTeX files too. These are files that you directly create and edit ``by
-hand''.
+LaTeX files too. These are files that you directly create and edit by
+hand.
 
 The files at the far right are clearly derived and are often described
-as ``targets''. These files are programmatically generated from source
-files (and possibly other inputs). These files are the product and they
-have external value, often for communicating ideas and results.
+as \textbf{targets}. These files are programmatically generated from
+source files (and possibly other inputs). These files are the product
+and they have external value, often for communicating ideas and results.
 
@@
 repository. A file that Git does not ignore is said to be
-\emph{tracked}.
+\textbf{tracked}.
 
@@
 derived file or file type?
+
+\emph{I suspect this is a good place to be less wordy? I am so
+frustrated with people doing stupid things here, though!}
 
@@
 
-\emph{Will it make collaboration harder?} \emph{talk about the scenario
-Nick brought up re: binary files like PDF being a common source of merge
-conflict
-\url{https://github.com/dsscollection/git-github-for-stats/issues/6}}
+\emph{Will it make collaboration harder?} Prose to be written: Talk
+about the scenario Nick brought up re: binary files like PDF being a
+common source of merge conflict
+\url{https://github.com/dsscollection/git-github-for-stats/issues/6}. I
+have a solution! Stop using PDF as your default output format! Make
+Markdown the output format during development. Your Git/GitHub problems
+just went away. You're welcome.
 
@@
 Collaboration is probably the most compelling reason to manage a project
-with Git and GitHub. I have a broad definition of collaboration here,
-that includes hands-on participation by multiple people as well as an
+with Git and GitHub. I have a broad definition of collaboration. It
+includes hands-on participation by multiple people as well as an
 asymmetric model, in which some people are active makers and others only
@@
 there can be independent lines of development. The real power comes from
-regular and structured reconciliation of all the files in the project
-via Git/GitHub. It is definitely more complicated than collaborating on
-a Google Doc, but also more powerful.
+regular and structured reconciliation of all versions of all the files
+in the project via Git/GitHub. It is definitely more complicated than
+collaborating on a Google Doc, but also more powerful.
 
@@
 This is another reason for all parties to commit, pull, and push often.
-Small changes, being integrated frequently, in non-binary files, are the
+Small changes, integrated frequently, in non-binary files, are the
 easiest for Git to automatically merge for you. The difficulty of
@@
 
-\begin{figure}
-\centering
-\includegraphics{bartlett-merge-commit.png}
-\caption{}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{bartlett-pull.png}
-\caption{}
-\end{figure}
+\includegraphics[width=1\linewidth]{bartlett-merge-commit}
+
+\includegraphics[width=1\linewidth]{bartlett-pull}
 
@@
 
-Transparency about \citep{greenwade93} process and product is
-increasingly important in science. The SOMETHING for reproducibility is
-well accepted. A more underappreciated benefit is democratization of our
-field, as this affords a much broader audience a clear view of how
-scientists and programmers work.
+Transparency about process and product is increasingly important in
+science. The SOMETHING for reproducibility is well accepted. A more
+underappreciated benefit is democratization of our field, as this
+affords a much broader audience a clear view of how scientists and
+programmers work.
+
+make statistical thought and implementation available
 
@@
 
-Does creation and presence in GitHub repo make the project more readable
-and navigable? If so DO IT.
-
-need to add a ton of links, citations
-
-make some bespoke examples or diagrams?
-
-Doing = documenting = sharing
-
-make statistical thought and implementation available
+Consider links to GitHub repos that exemplify certain points and are
+very unlikely to disappear any time soon.n
 
@@
 
-\href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio IDE}.
-
 \bibliographystyle{agsm}

### 2017-05-29 README.Rmd
Trim section about what to commit

@@
     The author gratefully acknowledges the constructive feedback from reviewers
-    Nicholas Horton and Colin Rundel.
+    Nicholas Horton, Colin Rundel, and Hadley Wickham.
 blinded: 0
@@
 
-The files in a project play different roles and arise in different ways. Let's have a few examples in mind for this discussion:
+The files in a project arise in different ways and play different roles.
+A critical issue for workflow happiness is to figure out how you want to handle different file types with respect to Git.
+You can direct Git to ignore specific files or file types, such as autosaves created by your editor.
+This reduces noise and clutter: Git will not pester you to commit changes to these files and they will not appear in your GitHub repository.
+A file that Git does not ignore is said to be __tracked__.
+How should you think about this?
 
-  * R markdown `.Rmd` --> markdown `.md`
-  * R markdown `.Rmd` --> markdown `.md` --> `.html` or `.pdf` or `.docx`
-  * R script `.R` --> results as `.csv` or `.rds` and figures `.png`
-  * LaTeX `.tex` --> `.aux`, `.bbl`, etc. --> `.pdf`
+__Source files:__ These files are created and edited by hand, such as R scripts and R Markdown or LaTeX files. This could also include the raw data for an analysis.
 
-The files at the far left are clearly source files.
-In the case of an R script, this is literally true, but it's morally true for R markdown and LaTeX files too.
-These are files that you directly create and edit by hand.
+__Configuration files:__ These files modify the behavior of a tool, for example `.gitignore` identifies files Git should not track and `some-project.Rproj` records RStudio project settings.
 
-The files at the far right are clearly derived and are often described as __targets__.
-These files are programmatically generated from source files (and possibly other inputs).
-These files are the product and they have external value, often for communicating ideas and results.
+__Derived products:__ These files are programmatically generated from source files and have external value.
+By executing `.R` or rendering `.Rmd` files, you obtain artifacts such as intermediate data (e.g., `.csv` or `.rds`), figures (e.g., `.png` or `.pdf`), and reports (e.g., `.md`, `.pdf`, `.docx`, or `.html`).
 
-The files in the middle are intermediates.
-Like targets, they are programmatically generated, but, unlike targets, no one necessarily cares about them.
-However, note that intermediate Markdown `.md` is an exception, since it is extremely useful on GitHub -- much more so than `.html`, `.pdf`, or `.docx` -- and is more like an additional target.
+**Intermediates:** These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTex workflows).
 
-A critical issue for workflow happiness is figuring out how to manage the production and storage of source, intermediates, and targets with respect to Git.
-You can direct Git to ignore specific files or certain types of files, such as autosaves created by your editor.
-This reduces clutter in your project: Git will not pester you to commit changes to these files and they will not appear in the associated GitHub repository.
-A file that Git does not ignore is said to be __tracked__.
+There is clear consensus that source files should be tracked.
+It is also common to track project-specific configuration files and to ignore intermediates.
+However, reasonable people can disagree about how to handle derived products and whether a specific file is an intermediate or derived product.
+Therefore, the main takeaway is to pick a policy that works for you and adapt as your needs change.
+There is no right answer.
+I would err on the side of committing more rather than less at first.
+What else should you consider when choosing files to track with Git and share on GitHub?
 
-The only point on which there is consensus is that source files should absolutely be tracked. The best treatment of intermediates and targets with respect to Git is much less clear cut.
-
-Therefore, the main message for intermediates and targets is that you can pick a policy that works for you and adapt as your needs change.
-There is no right answer.
-I suggest erring on the side of committing everything at first.
-
-What are the main considerations when deciding whether to track a derived file or file type?
-
-*I suspect this is a good place to be less wordy? I am so frustrated with people doing stupid things here, though!*
-
-*Is it immediately useful to someone?* If so, that is a reason to track it and push it to GitHub.
-There is a taboo against committing derived products, inherited from Git's software development roots.
-The reasoning is that compiled programs are platform-specific and, therefore, people are better served by getting current source from Git and compiling themselves.
-I think data analytic targets, like figures and rendered reports, are very different beasts and it's misguided to reflexively exclude them from version control.
-They are immediately useful, especially to consumers of a project (versus the makers).
-To the extent that a GitHub repo is meant for dissemination, there is no reason to burden every consumer with unnecessary friction.
-Most simply will not bother to clone the repo, install all the necessary dependencies, and remake the products.
+**Is it useful to someone?** If so, track and share!
+There is a taboo against committing derived products, inherited from Git's software development roots, because their typical product is a platform-specific executable.
+This rationale, however, does not apply to many data science products.
+Rendered reports, figures, and cleaned data are often extremely valuable to others.
 Make them readily available.
 
-*Is it available elsewhere?* If so, then perhaps you don't need to track it and push it to GitHub.
-Many people who have a policy of not tracking derived products also have a system that makes these available elsewhere, such as on a separate website.
-There are ways to automate this via GitHub, but that is beyond the scope of this article and not recommended for your early days with Git and GitHub.
-Beware those who recommend the exclusion of derived products without offering any practical solution for making them available some other way.
+**Will it play especially well with Git/GitHub?** This boils down to whether Git diffs will be informative and whether GitHub has nice handling for the file type.
+Small-to-medium plain text files with hard line breaks are a perfect fit, but there are a few more pleasant surprises.
 
-*Is it huge or changing often?
-Is it a format that is of little use on GitHub?* These are all good reasons to not track a file with Git.
-They can make your repository bloated and slow down pushes and pulls.
-If a file is binary, such as a Word document or Excel spreadsheet, Git and Git clients will not be able to provide human-readable diffs.
-Neither will GitHub be able to directly display this file in the browser.
-Be aware, however, GitHub-friendliness does not just boil down to "is it plain text?".
-GitHub has excellent support for non-code files, such as image formats (PNG, JPG, GIF, and SVG) and PDFs.
-It even provides visual diffs, which are extremely useful for understanding changes in figures.
-Finally, even though HTML is plain text, it is of little direct use on GitHub, because, unlike Markdown, it is not rendered.
+  * Some derived files are too miserable to read casually, such as `.csv` files of processed results or `.html` derived from `.Rmd`, but are still worth tracking with Git.
+When you re-run an analysis with updated input data or after updating R packages, *the diffs are often quite modest* and help you pinpoint unexpected changes.
+  * GitHub has excellent [support for a variety of non-code files](https://help.github.com/categories/working-with-non-code-files/), such as CSV and TSV.
+It also [displays and provides visual diffs](https://help.github.com/articles/rendering-and-diffing-images/) for the most common image formats, which is extremely useful for spotting unexpected changes in figures.
+  * Never, ever forget that Markdown is highly privileged on GitHub and is, therefore, a crucial output format for rendered `.R` and `.Rmd`.
 
-*Will diffs be useful to you?* Some derived files are simply too big or miserable to read casually, such as `.csv` files of processed results or `.html` derived from `.Rmd`.
-But they may still be worth tracking with Git, because the diffs are often modest and quite interesting.
-I have caught unexpected changes in analytical results and student-facing webpages this way.
-When you re-run an analysis with updated input data or after updating R packages, the diffs presented by Git help you quickly pinpoint the downstream consequences.
-
-*Will it make collaboration harder?* Prose to be written: Talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>.
-I have a solution!
-Stop using PDF as your default output format!
-Make Markdown the output format during development.
-Your Git/GitHub problems just went away.
-You're welcome.
-
-In summary, I recommend you default to including a file in your Git repository, unless there's a specific reason not to. But good reasons absolutely do exist.
+**Will it cause problems with Git/GitHub?** This boils down to whether Git diffs are informative, whether GitHub can render the file usefully, and the file's likely effect on Git operations.
+A file that is large and changing quite often can make your repository bloated and slow down pushes and pulls.
+If a file is binary, such as a Word document or Excel spreadsheet, you will not get human-readable diffs, nor can GitHub display it in the browser.
+Html is not rendered on GitHub, but is, instead, displayed in raw form.
+Binary files are also a prime source of merge conflicts, because they are beyond the reach of Git's sophisticated automatic merging logic.
+For many of us, adoption of Git/GitHub should prompt a pivot away from `.docx` and `.pdf` as primary output formats and towards `.md`, at least during periods of rapid development.
 

### 2017-05-29 README.tex
Trim section about what to commit

@@
         Jennifer Bryan \thanks{The author gratefully acknowledges the constructive feedback from
-reviewers Nicholas Horton and Colin Rundel.} \\
+reviewers Nicholas Horton, Colin Rundel, and Hadley Wickham.} \\
     RStudio and the Department of Statistics, University of British Columbia\\
@@
 
-The files in a project play different roles and arise in different ways.
-Let's have a few examples in mind for this discussion:
+The files in a project arise in different ways and play different roles.
+A critical issue for workflow happiness is to figure out how you want to
+handle different file types with respect to Git. You can direct Git to
+ignore specific files or file types, such as autosaves created by your
+editor. This reduces noise and clutter: Git will not pester you to
+commit changes to these files and they will not appear in your GitHub
+repository. A file that Git does not ignore is said to be
+\textbf{tracked}. How should you think about this?
+
+\textbf{Source files:} These files are created and edited by hand, such
+as R scripts and R Markdown or LaTeX files. This could also include the
+raw data for an analysis.
+
+\textbf{Configuration files:} These files modify the behavior of a tool,
+for example \texttt{.gitignore} identifies files Git should not track
+and \texttt{some-project.Rproj} records RStudio project settings.
+
+\textbf{Derived products:} These files are programmatically generated
+from source files and have external value. By executing \texttt{.R} or
+rendering \texttt{.Rmd} files, you obtain artefacts such as intermediate
+data (e.g., \texttt{.csv} or \texttt{.rds}), figures (e.g.,
+\texttt{.png} or \texttt{.pdf}), and reports (e.g., \texttt{.md},
+\texttt{.pdf}, \texttt{.docx}, or \texttt{.html}).
+
+\textbf{Intermediates:} These files are programmatically generated and
+serve a temporary purpose, but are not intrinsically valuable (e.g.,
+\texttt{.aux} and \texttt{.log} in LaTex workflows).
+
+There is clear consensus that source files should be tracked. It is also
+common to track project-specific configuration files and to ignore
+intermediates. However, reasonable people can disagree about how to
+handle derived products and whether a specific file is an intermediate
+or derived product. Therefore, the main takeaway is to pick a policy
+that works for you and adapt as your needs change. There is no right
+answer. I would err on the side of committing more rather than less at
+first. What else should you consider when choosing files to track with
+Git and share on GitHub?
+
+\textbf{Is it useful to someone?} If so, track and share! There is a
+taboo against committing derived products, inherited from Git's software
+development roots, because their typical product is a platform-specific
+executable. This rationale, however, does not apply to many data science
+products. Rendered reports, figures, and cleaned data are often
+extremely valuable to others. Make them readily available.
+
+\textbf{Will it play especially well with Git/GitHub?} This boils down
+to whether Git diffs will be informative and whether GitHub has nice
+handling for the file type. Small-to-medium plain text files with hard
+line breaks are a perfect fit, but there are a few more pleasant
+surprises.
 
@@
 \item
-  R markdown \texttt{.Rmd} --\textgreater{} markdown \texttt{.md}
-\item
-  R markdown \texttt{.Rmd} --\textgreater{} markdown \texttt{.md}
-  --\textgreater{} \texttt{.html} or \texttt{.pdf} or \texttt{.docx}
-\item
-  R script \texttt{.R} --\textgreater{} results as \texttt{.csv} or
-  \texttt{.rds} and figures \texttt{.png}
-\item
-  LaTeX \texttt{.tex} --\textgreater{} \texttt{.aux}, \texttt{.bbl},
-  etc. --\textgreater{} \texttt{.pdf}
+  Some derived files are too miserable to read casually, such as
+  \texttt{.csv} files of processed results or \texttt{.html} derived
+  from \texttt{.Rmd}, but are still worth tracking with Git. When you
+  re-run an analysis with updated input data or after updating R
+  packages, \emph{the diffs are often quite modest} and help you
+  pinpoint unexpected changes.
+\item
+  GitHub has excellent
+  \href{https://help.github.com/categories/working-with-non-code-files/}{support
+  for a variety of non-code files}, such as CSV and TSV. It also
+  \href{https://help.github.com/articles/rendering-and-diffing-images/}{displays
+  and provides visual diffs} for the most common image formats, which is
+  extremely useful for spotting unexpected changes in figures.
+\item
+  Never, ever forget that Markdown is highly privileged on GitHub and
+  is, therefore, a crucial output format for rendered \texttt{.R} and
+  \texttt{.Rmd}.
 \end{itemize}
 
-The files at the far left are clearly source files. In the case of an R
-script, this is literally true, but it's morally true for R markdown and
-LaTeX files too. These are files that you directly create and edit by
-hand.
-
-The files at the far right are clearly derived and are often described
-as \textbf{targets}. These files are programmatically generated from
-source files (and possibly other inputs). These files are the product
-and they have external value, often for communicating ideas and results.
-
-The files in the middle are intermediates. Like targets, they are
-programmatically generated, but, unlike targets, no one necessarily
-cares about them. However, note that intermediate Markdown \texttt{.md}
-is an exception, since it is extremely useful on GitHub -- much more so
-than \texttt{.html}, \texttt{.pdf}, or \texttt{.docx} -- and is more
-like an additional target.
-
-A critical issue for workflow happiness is figuring out how to manage
-the production and storage of source, intermediates, and targets with
-respect to Git. You can direct Git to ignore specific files or certain
-types of files, such as autosaves created by your editor. This reduces
-clutter in your project: Git will not pester you to commit changes to
-these files and they will not appear in the associated GitHub
-repository. A file that Git does not ignore is said to be
-\textbf{tracked}.
-
-The only point on which there is consensus is that source files should
-absolutely be tracked. The best treatment of intermediates and targets
-with respect to Git is much less clear cut.
-
-Therefore, the main message for intermediates and targets is that you
-can pick a policy that works for you and adapt as your needs change.
-There is no right answer. I suggest erring on the side of committing
-everything at first.
-
-What are the main considerations when deciding whether to track a
-derived file or file type?
-
-\emph{I suspect this is a good place to be less wordy? I am so
-frustrated with people doing stupid things here, though!}
-
-\emph{Is it immediately useful to someone?} If so, that is a reason to
-track it and push it to GitHub. There is a taboo against committing
-derived products, inherited from Git's software development roots. The
-reasoning is that compiled programs are platform-specific and,
-therefore, people are better served by getting current source from Git
-and compiling themselves. I think data analytic targets, like figures
-and rendered reports, are very different beasts and it's misguided to
-reflexively exclude them from version control. They are immediately
-useful, especially to consumers of a project (versus the makers). To the
-extent that a GitHub repo is meant for dissemination, there is no reason
-to burden every consumer with unnecessary friction. Most simply will not
-bother to clone the repo, install all the necessary dependencies, and
-remake the products. Make them readily available.
-
-\emph{Is it available elsewhere?} If so, then perhaps you don't need to
-track it and push it to GitHub. Many people who have a policy of not
-tracking derived products also have a system that makes these available
-elsewhere, such as on a separate website. There are ways to automate
-this via GitHub, but that is beyond the scope of this article and not
-recommended for your early days with Git and GitHub. Beware those who
-recommend the exclusion of derived products without offering any
-practical solution for making them available some other way.
-
-\emph{Is it huge or changing often? Is it a format that is of little use
-on GitHub?} These are all good reasons to not track a file with Git.
-They can make your repository bloated and slow down pushes and pulls. If
-a file is binary, such as a Word document or Excel spreadsheet, Git and
-Git clients will not be able to provide human-readable diffs. Neither
-will GitHub be able to directly display this file in the browser. Be
-aware, however, GitHub-friendliness does not just boil down to ``is it
-plain text?''. GitHub has excellent support for non-code files, such as
-image formats (PNG, JPG, GIF, and SVG) and PDFs. It even provides visual
-diffs, which are extremely useful for understanding changes in figures.
-Finally, even though HTML is plain text, it is of little direct use on
-GitHub, because, unlike Markdown, it is not rendered.
-
-\emph{Will diffs be useful to you?} Some derived files are simply too
-big or miserable to read casually, such as \texttt{.csv} files of
-processed results or \texttt{.html} derived from \texttt{.Rmd}. But they
-may still be worth tracking with Git, because the diffs are often modest
-and quite interesting. I have caught unexpected changes in analytical
-results and student-facing webpages this way. When you re-run an
-analysis with updated input data or after updating R packages, the diffs
-presented by Git help you quickly pinpoint the downstream consequences.
-
-\emph{Will it make collaboration harder?} Prose to be written: Talk
-about the scenario Nick brought up re: binary files like PDF being a
-common source of merge conflict
-\url{https://github.com/dsscollection/git-github-for-stats/issues/6}. I
-have a solution! Stop using PDF as your default output format! Make
-Markdown the output format during development. Your Git/GitHub problems
-just went away. You're welcome.
-
-In summary, I recommend you default to including a file in your Git
-repository, unless there's a specific reason not to. But good reasons
-absolutely do exist.
+\textbf{Will it cause problems with Git/GitHub?} This boils down to
+whether Git diffs are informative, whether GitHub can render the file
+usefully, and the file's likely effect on Git operations. A file that is
+large and changing quite often can make your repository bloated and slow
+down pushes and pulls. If a file is binary, such as a Word document or
+Excel spreadsheet, you will not get human-readable diffs, nor can GitHub
+display it in the browser. Html is not rendered on GitHub, but is,
+instead, displayed in raw form. Binary files are also a prime source of
+merge conflicts, because they are beyond the reach of Git's
+sophisticated automatic merging logic. For many of us, adoption of
+Git/GitHub should prompt a pivot away from \texttt{.docx} and
+\texttt{.pdf} as primary output formats and towards \texttt{.md}, at
+least during periods of rapid development.
 

### 2017-05-29 README.Rmd
Cite Ten Simple Rules for Taking Advantage of Git and GitHub

@@
 Git has been re-purposed by the data science community [@Ram2013;
-@git-for-humans]. In addition to using it for source code, we use it to manage 
-the motley collection of files that make up typical data analytical projects, 
-which often consist of data, figures, reports, and, yes, source code. Even those
-who identify more as statistician than data scientist generally have a similar 
-mix of files that are the artifacts of a project.
+@git-for-humans; @ten-simple-rules-git].
+We use it to manage the motley collection of files that make up typical data analytical projects, which often consist of data, figures, reports, and, yes, source code.
+Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.
 
@@
 
+A Quick Introduction to Version Control with Git and GitHub
+http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004668
+
 The active R package development community on GitHub.

### 2017-05-30 README.Rmd
Remove course stuff; rescue paragraph on GitHub Pages

@@
 
-## GitHub as course management system
-
-[STAT 545](http://stat545.com) is a data wrangling and analysis course at the University of British Columbia.
-I was the instructor in charge for several years, which coincided with my own adoption of Git/GitHub.
-GitHub is used to manage the development of course material, to serve the course website, to create a discussion forum, and to host all student-submitted work.
-
-Given that students must submit their work and provide peer review of others' work via GitHub, the use of hosted version control is an explicit, though modest, part of the course.
-The website [Happy Git and GitHub for the useR](http://happygitwithr.com) holds our battle-tested instructions for setup and early usage.
-The students achieve basic competence quite quickly and find it gratifying to see their formatted, figure-rich R Markdown reports up on the internet.
-Since it’s easy to expose their work within the class, we conduct peer review, which helps expertise to spread quickly through the group.
-
-### Use a GitHub Organization
-
-[GitHub Organizations](https://help.github.com/articles/differences-between-user-and-organization-accounts/) are "shared accounts where groups of people can collaborate across many projects at once".
-This is the most appropriate structure for stewarding course resources, since I can grant TAs and students different levels of access to various repositories.
-Access can be controlled at the individual user level or, more conveniently, for entire [Teams](https://help.github.com/articles/setting-up-teams/).
-The TA Team shares write access with me on a private repository for internal matters.
-I provide each student with their own private repository for coursework and grant other members of the Students Team read access, in order to facilitate peer review.
-There is a public repository that underpins the course website (see below).
-We have one other public repository that exists solely so the [Issues](https://github.com/STAT545-UBC/Discussion/issues) can be used as a discussion forum.
-
-GitHub actively encourages the use of its platform in teaching.
-As an instructor you can request a [free Organization account](https://help.github.com/articles/discounted-organization-accounts/) that provides features normally available only on paid plans, such as private repositories.
-In fact, GitHub provides tooling for specific teaching workflows via [GitHub Classroom](https://classroom.github.com/), although I do not use it.
-That is not an intentional knock on their tools.
-I started teaching with GitHub several years before this existed and developed a different way of using the platform.
-I also find the [GitHub Education](https://education.github.com) resources to be geared more towards computer science than data science.
-
-### GitHub Pages for course website
-
-All course content is provided on the [STAT 545 website](http://stat545.com).
-Each page is generated from an R Markdown document that is rendered to HTML locally using the rmarkdown package, retaining the intermediate Markdown.
-These pages are a mix of prose and rendered R code, reflecting the live coding done in class.
-All of these files and their history can be explored in the [source repository](https://github.com/STAT545-UBC/STAT545-UBC.github.io).
-The TA team has permission to write to this repo, meaning they can (and do!) help me maintain the website.
-I rejoice that I am no longer the webmaster.
-We also get typo corrections and other input from the world at large, since this is entirely public.
-
-If I were starting from scratch today, I would continue to use R Markdown, RStudio, and GitHub Pages (see below), but would upgrade to a more modern, automated approach to rendering the pages.
-I now recommend [R Markdown websites](http://rmarkdown.rstudio.com/rmarkdown_websites.html), [bookdown](https://bookdown.org), or [blogdown](https://bookdown.org/yihui/blogdown/) to manage the process of creating a static website from a large and inter-related set of `.Rmd` files.
-
-GitHub offers several ways to host a website directly from a repository, collectively known as [GitHub Pages](https://help.github.com/categories/github-pages-basics/).
-The STAT 545 website is a very simple [Organization Page](https://help.github.com/articles/user-organization-and-project-pages/) that uses a [custom domain](https://help.github.com/articles/custom-domain-redirects-for-github-pages-sites/), `stat545.com`, instead of the default `orgname.github.io`.
-
-This system for managing course content is a great example of integrating the doing of work and the sharing of it.
-We analyze data live in class, using R, based on the scripts on the website.
-I re-render the associated `.R` or `.Rmd`, commit the changed files, push, and see it reflected right away on http://stat545.com.
-There is no separation between having an idea, implementing it, and posting on the website.
-
-### Student-specific private repos
-
-Early in the course I elicit GitHub usernames for registered students, via a [Shiny app](https://shiny.rstudio.com), and invite them to join the course Organization.
-I then create one private repository per student, in the STAT 545 Organization.
-The targeted student has write access and the other students have read access.
-This is somewhat controversial, due to the possibility of cheating, but I have seen more pros than cons for this setup, in the STAT 545 context.
-In other settings, I have also used one repo per student *per homework assignment*, which allows you to keep the repos completely private until homework submission, then increase their visibility during marking and peer review.
-Some courses will work better with one model or the other.
-
-Each student does their work in this repo, submitting a major assignment approximately once a week.
-The first assignment is simply to claim the repository and create a README, which proves they have all the relevant software setup and they can write a little Markdown.
-Each week we tackle some new data analysis or wrangling task, with increasing latitude for independence.
-Homework is implemented in R Markdown documents, rendered to Markdown, and pushed to GitHub.
-Students submit their work by opening an issue in their repo, naming the assignment in the title, providing the SHA of the associated final commit, and linking to the main `.md` file.
-We leave feedback as comments in the issue thread or, occasionally, propose changes to code via "pull requests".
-Two peers are selected at random to review each assignment, a process that we also implement via GitHub Issues.
-
-At the end of term, the student (and their instructor!) can visit the repo to find an organized, navigable sequence of ~10 assignments.
-Each student leaves with self-written documentation of everything they've done, ready to consult in future projects.
-The last assignments require writing an R package or Shiny app, which they generally do in public repositories under their own accounts.
-They finish STAT 545 with several months of Git/GitHub experience and the start of a data science portfolio.
-
 ## GitHub as web presence
@@
 They can grab a snapshot of all the files as a ZIP archive by simply clicking a button.
-People with GitHub accounts have even more options.
-They can clone or fork the repository to get their own copy, which also makes it easy for them to stay current on future changes.
+People with GitHub accounts can use Git-specific methods to make their own copy, i.e.
+a clone or a fork, which make it easy to keep current with future changes.
 
-As described above, GitHub Pages offer various ways to serve proper websites, even quite sophisticated ones, directly out of a GitHub repo.
-But before you even worry about that, certain practices can make a GitHub repository much more browsable.
-For many projects, this is more than sufficient for granting people access to your work.
+GitHub offers several ways to host a proper website directly from a repository, collectively known as [GitHub Pages](https://help.github.com/categories/github-pages-basics/).
+At one extreme, as long as you've got one Markdown file, [GitHub Pages can create a simple website](https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3) automatically.
+At the other extreme, sophisticated users can take full advantage of the [Jekyll static site generator](https://jekyllrb.com).
+The STAT 545 website ([stat545.com](http://stat545.com), [GitHub repo](https://github.com/STAT545-UBC/STAT545-UBC.github.io)) falls on the more primitive side.
+[R Markdown websites](http://rmarkdown.rstudio.com/rmarkdown_websites.html), [bookdown](https://bookdown.org) [@bookdown-pkg; @bookdown-book], and [blogdown](https://bookdown.org/yihui/blogdown/) provide several R-focused options for rendering the pages *en masse*.
+
+But even before you make an actual website, certain practices can [make your GitHub repository much more browsable](http://happygitwithr.com/repo-browsability.html).
+For many projects, this is more than sufficient for helping people connect with your work.
 
@@
 For figures produced by R code, these links are part of what rmarkdown takes care of for you, but there's no reason you can't do the same yourself for any Markdown file.
-
-*link to http://happygitwithr.com/repo-browsability.html*
 

### 2017-05-30 README.tex
Re-render everything

@@
 Git has been re-purposed by the data science community
-\citep{Ram2013, git-for-humans}. In addition to using it for source
-code, we use it to manage the motley collection of files that make up
-typical data analytical projects, which often consist of data, figures,
-reports, and, yes, source code. Even those who identify more as
-statistician than data scientist generally have a similar mix of files
-that are the artifacts of a project.
+\citep{Ram2013, git-for-humans, ten-simple-rules-git}. We use it to
+manage the motley collection of files that make up typical data
+analytical projects, which often consist of data, figures, reports, and,
+yes, source code. Even those who identify more as statistician than data
+scientist generally have a similar mix of files that are the artifacts
+of a project.
 
@@
 from source files and have external value. By executing \texttt{.R} or
-rendering \texttt{.Rmd} files, you obtain artefacts such as intermediate
+rendering \texttt{.Rmd} files, you obtain artifacts such as intermediate
 data (e.g., \texttt{.csv} or \texttt{.rds}), figures (e.g.,
@@
 
-\subsection{GitHub as course management
-system}\label{github-as-course-management-system}
-
-\href{http://stat545.com}{STAT 545} is a data wrangling and analysis
-course at the University of British Columbia. I was the instructor in
-charge for several years, which coincided with my own adoption of
-Git/GitHub. GitHub is used to manage the development of course material,
-to serve the course website, to create a discussion forum, and to host
-all student-submitted work.
-
-Given that students must submit their work and provide peer review of
-others' work via GitHub, the use of hosted version control is an
-explicit, though modest, part of the course. The website
-\href{http://happygitwithr.com}{Happy Git and GitHub for the useR} holds
-our battle-tested instructions for setup and early usage. The students
-achieve basic competence quite quickly and find it gratifying to see
-their formatted, figure-rich R Markdown reports up on the internet.
-Since it's easy to expose their work within the class, we conduct peer
-review, which helps expertise to spread quickly through the group.
-
-\subsubsection{Use a GitHub
-Organization}\label{use-a-github-organization}
-
-\href{https://help.github.com/articles/differences-between-user-and-organization-accounts/}{GitHub
-Organizations} are ``shared accounts where groups of people can
-collaborate across many projects at once''. This is the most appropriate
-structure for stewarding course resources, since I can grant TAs and
-students different levels of access to various repositories. Access can
-be controlled at the individual user level or, more conveniently, for
-entire \href{https://help.github.com/articles/setting-up-teams/}{Teams}.
-The TA Team shares write access with me on a private repository for
-internal matters. I provide each student with their own private
-repository for coursework and grant other members of the Students Team
-read access, in order to facilitate peer review. There is a public
-repository that underpins the course website (see below). We have one
-other public repository that exists solely so the
-\href{https://github.com/STAT545-UBC/Discussion/issues}{Issues} can be
-used as a discussion forum.
-
-GitHub actively encourages the use of its platform in teaching. As an
-instructor you can request a
-\href{https://help.github.com/articles/discounted-organization-accounts/}{free
-Organization account} that provides features normally available only on
-paid plans, such as private repositories. In fact, GitHub provides
-tooling for specific teaching workflows via
-\href{https://classroom.github.com/}{GitHub Classroom}, although I do
-not use it. That is not an intentional knock on their tools. I started
-teaching with GitHub several years before this existed and developed a
-different way of using the platform. I also find the
-\href{https://education.github.com}{GitHub Education} resources to be
-geared more towards computer science than data science.
-
-\subsubsection{GitHub Pages for course
-website}\label{github-pages-for-course-website}
-
-All course content is provided on the \href{http://stat545.com}{STAT 545
-website}. Each page is generated from an R Markdown document that is
-rendered to HTML locally using the rmarkdown package, retaining the
-intermediate Markdown. These pages are a mix of prose and rendered R
-code, reflecting the live coding done in class. All of these files and
-their history can be explored in the
-\href{https://github.com/STAT545-UBC/STAT545-UBC.github.io}{source
-repository}. The TA team has permission to write to this repo, meaning
-they can (and do!) help me maintain the website. I rejoice that I am no
-longer the webmaster. We also get typo corrections and other input from
-the world at large, since this is entirely public.
-
-If I were starting from scratch today, I would continue to use R
-Markdown, RStudio, and GitHub Pages (see below), but would upgrade to a
-more modern, automated approach to rendering the pages. I now recommend
-\href{http://rmarkdown.rstudio.com/rmarkdown_websites.html}{R Markdown
-websites}, \href{https://bookdown.org}{bookdown}, or
-\href{https://bookdown.org/yihui/blogdown/}{blogdown} to manage the
-process of creating a static website from a large and inter-related set
-of \texttt{.Rmd} files.
-
-GitHub offers several ways to host a website directly from a repository,
-collectively known as
-\href{https://help.github.com/categories/github-pages-basics/}{GitHub
-Pages}. The STAT 545 website is a very simple
-\href{https://help.github.com/articles/user-organization-and-project-pages/}{Organization
-Page} that uses a
-\href{https://help.github.com/articles/custom-domain-redirects-for-github-pages-sites/}{custom
-domain}, \texttt{stat545.com}, instead of the default
-\texttt{orgname.github.io}.
-
-This system for managing course content is a great example of
-integrating the doing of work and the sharing of it. We analyze data
-live in class, using R, based on the scripts on the website. I re-render
-the associated \texttt{.R} or \texttt{.Rmd}, commit the changed files,
-push, and see it reflected right away on \url{http://stat545.com}. There
-is no separation between having an idea, implementing it, and posting on
-the website.
-
-\subsubsection{Student-specific private
-repos}\label{student-specific-private-repos}
-
-Early in the course I elicit GitHub usernames for registered students,
-via a \href{https://shiny.rstudio.com}{Shiny app}, and invite them to
-join the course Organization. I then create one private repository per
-student, in the STAT 545 Organization. The targeted student has write
-access and the other students have read access. This is somewhat
-controversial, due to the possibility of cheating, but I have seen more
-pros than cons for this setup, in the STAT 545 context. In other
-settings, I have also used one repo per student \emph{per homework
-assignment}, which allows you to keep the repos completely private until
-homework submission, then increase their visibility during marking and
-peer review. Some courses will work better with one model or the other.
-
-Each student does their work in this repo, submitting a major assignment
-approximately once a week. The first assignment is simply to claim the
-repository and create a README, which proves they have all the relevant
-software setup and they can write a little Markdown. Each week we tackle
-some new data analysis or wrangling task, with increasing latitude for
-independence. Homework is implemented in R Markdown documents, rendered
-to Markdown, and pushed to GitHub. Students submit their work by opening
-an issue in their repo, naming the assignment in the title, providing
-the SHA of the associated final commit, and linking to the main
-\texttt{.md} file. We leave feedback as comments in the issue thread or,
-occasionally, propose changes to code via ``pull requests''. Two peers
-are selected at random to review each assignment, a process that we also
-implement via GitHub Issues.
-
-At the end of term, the student (and their instructor!) can visit the
-repo to find an organized, navigable sequence of \textasciitilde{}10
-assignments. Each student leaves with self-written documentation of
-everything they've done, ready to consult in future projects. The last
-assignments require writing an R package or Shiny app, which they
-generally do in public repositories under their own accounts. They
-finish STAT 545 with several months of Git/GitHub experience and the
-start of a data science portfolio.
-
 \subsection{GitHub as web presence}\label{github-as-web-presence}
@@
 like a webpage. They can grab a snapshot of all the files as a ZIP
-archive by simply clicking a button. People with GitHub accounts have
-even more options. They can clone or fork the repository to get their
-own copy, which also makes it easy for them to stay current on future
-changes.
+archive by simply clicking a button. People with GitHub accounts can use
+Git-specific methods to make their own copy, i.e.~a clone or a fork,
+which make it easy to keep current with future changes.
 
-As described above, GitHub Pages offer various ways to serve proper
-websites, even quite sophisticated ones, directly out of a GitHub repo.
-But before you even worry about that, certain practices can make a
-GitHub repository much more browsable. For many projects, this is more
-than sufficient for granting people access to your work.
+GitHub offers several ways to host a proper website directly from a
+repository, collectively known as
+\href{https://help.github.com/categories/github-pages-basics/}{GitHub
+Pages}. At one extreme, as long as you've got one Markdown file,
+\href{https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3}{GitHub
+Pages can create a simple website} automatically. At the other extreme,
+sophisticated users can take full advantage of the
+\href{https://jekyllrb.com}{Jekyll static site generator}. The STAT 545
+website (\href{http://stat545.com}{stat545.com},
+\href{https://github.com/STAT545-UBC/STAT545-UBC.github.io}{GitHub
+repo}) falls on the more primitive side.
+\href{http://rmarkdown.rstudio.com/rmarkdown_websites.html}{R Markdown
+websites}, \href{https://bookdown.org}{bookdown}
+\citep{bookdown-pkg, bookdown-book}, and
+\href{https://bookdown.org/yihui/blogdown/}{blogdown} provide several
+R-focused options for rendering the pages \emph{en masse}.
+
+But even before you make an actual website, certain practices can
+\href{http://happygitwithr.com/repo-browsability.html}{make your GitHub
+repository much more browsable}. For many projects, this is more than
+sufficient for helping people connect with your work.
 
@@
 same yourself for any Markdown file.
-
-*link to \url{http://happygitwithr.com/repo-browsability.html*}
 
@@
 
+A Quick Introduction to Version Control with Git and GitHub
+\url{http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004668}
+
 The active R package development community on GitHub.

### 2017-05-30 README.Rmd
Act on many items from Hadley's review + more

@@
 
-Why would a statistician use version control, such as Git? And what is the point
-of hosting your work online, e.g., on GitHub? Could the gains possibly justify 
-the inevitable pain?
+Why would a statistician use version control, such as [Git](http://git-scm.com)?
+And what is the point of hosting your work online, e.g., on [GitHub](https://github.com)?
+Could the gains possibly justify the inevitable pain?
 
@@
   * The marginal effort required to create a web presence for a project is negligible.
-  * GitHub makes a fantastic course management system for courses that use R. You and your students can exchange actual working code and explore the associated results.
+  * GitHub makes a fantastic course management system for courses that use R. You can exchange actual working code with your students and explore the associated results.
   * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.
+
+Now the bad news: Git was built neither for the exact usage described here, nor for broad usability.
+You will undoubtedly notice this, so it's best to know in advance.
+Happily, there are many helpful tools that mediate your interactions with Git.
+GitHub itself is a fine example, as is [RStudio](https://www.rstudio.com/products/rstudio/).
+In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.
 
@@
 In my opinion, __for new users__, the pros of Git only outweigh the cons when 
-you factor in the overhead of communicating and collaborating with other people,
-including your future self. And who among us does not need to do that? Your life
-is much easier if this is baked into your workflow, as opposed to being a 
-separate process that you dread and avoid. Communication and collaboration are 
-the killer apps of version control. Git enforces a rigorous model of file 
-management, but it is critical for the distribution of files across different 
-people, computers, and time.
+you consider the overhead of working with other people, including your future self.
+And who among us does not need to do that?
+In a Git-based workflow, you document and, optionally, expose your work as you go.
+Communication and collaboration are the killer apps of version control.
+Git's model of file management can feel uncomfortably rigid, but it enables the distribution of files across different people, computers, and time.
 
-This has an implication for selecting your first Git projects: in order to see 
-substantial gain for your pain, you need to pick projects that require you to 
-share rapidly evolving files with others. It is tempting to pick a quiet little 
-project that you work on privately, but you risk missing the main point of 
-formal version control.
+This has an implication for selecting your first Git projects: you will enjoy the most gain for your pain if you pick a project that involves sharing rapidly evolving files with others.
+It is tempting to pick a quiet, private project, but you risk missing out on the main benefits of formal version control.
 
@@
 each equipped with an explanatory message and a nickname. All collaborators sync
-regularly to a common version, acknowledging that the difficulty of merging goes
-up faster than the size of the difference [@good-enough]. Especially important 
+regularly to a common version, acknowledging that the difficulty of merging only goes up as more changes accumulate [@good-enough]. Especially important 
 versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
@@
 This article does not provide step-by-step instructions on how to use Git and 
-GitHub. This format would not be effective, but we do provide annotated links to
-such resources in ?the appendix or online supplement?. Instead, I'll convey what
-the workflow feels like and what the payoffs are, with special attention to the 
-statistics and R context. The goal is to help the Git-curious generate the 
+GitHub. This format would not be effective, but annotated links to
+such resources are given in the supplementary materials.
+Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context.
+The goal is to help the Git-curious generate the
 activation energy needed to get started.
@@
 
-[GitHub](https://github.com) is currently the most popular Git hosting service. 
-Others include [GitLab](https://about.gitlab.com) and 
-[Bitbucket](https://bitbucket.org). These services provide a home for Git-based 
-projects on the internet. It is a bit like DropBox or Google Drive, but more 
-structured, powerful, and programmatic.
+Git offers a powerful structure for file management and [GitHub](https://github.com) complements it with a slick user interface and distribution mechanism for Git repositories.
+GitHub is currently the most popular Git hosting service, providing a home for Git-based projects on the internet.
+It is a bit like DropBox or Google Drive, but more structured, powerful, and programmatic.
 
 The remote host acts as the clearinghouse for a Git-managed project. This allows
-other people to see the project files, sync up with the current version, and 
-perhaps even make or propose changes. These hosting providers offer 
-well-designed web interfaces that are a dramatic improvement over traditional 
-Unix Git servers. Many operations can be done entirely in the browser, such as 
-viewing current or past versions of a file, commenting on a recent change, and 
-editing or adding files. These hosts also offer granular control over who can 
-see, edit, and administer a project.
+others to see the project files, sync up with the current version, and perhaps even propose or make changes.
+GitHub has a well-designed web interface that is a dramatic improvement over traditional Unix Git servers.
+Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files.
+GitHub also offers granular control over who can see, edit, and administer a project.
 
@@
 
-Here is the initial setup process, which you do once or, for some steps, once per computer.
+Here is an overview of the initial setup process, which happens once or, for some steps, once per computer.
+This is excerpted from [Happy Git and GitHub for the useR](http://happygitwithr.com), which holds battle-tested instructions honed over several years in [STAT 545](http://stat545.com) at the University of British Columbia.
 
@@
 
-* Install Git. Depending on your OS, Git might already be installed. But many of
-us may need to install it or will choose to update to a more recent version.
+* Install Git. Depending on your OS, Git might already be installed. But many may need to install it or will choose to update to a more recent version.
 Some basic configuration is critical, such as setting your username and email.
@@
 current version of the repository on GitHub, known as __pushing__ and __pulling__, respectively.
-This will require authenticating yourself with GitHub from a local shell or Git client.
-At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.
 
 Once this setup is done, you are ready to start using Git and GitHub with your 
-projects.
-
-## Is this going to hurt?
-
-Yes.
-
-Git was built neither for the exact usage described here, nor for broad 
-usability. You will undoubtedly notice this, so it can be helpful to know this 
-in advance. Happily there are many helpful tools that mediate your interactions 
-with Git. GitHub itself is a fine example, as is RStudio. In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.
-
-General recommendations for agony reduction:
+projects. Some general recommendations for agony reduction:
 
@@
 
-Recall that a repository or repo is just a directory of files that Git will manage holistically.
+Recall that a repository or repo is just a directory of files that Git manages holistically.
 A commit functions like a snapshot of all the files in the repo, at a specific moment.
@@
 
-**Will it play especially well with Git/GitHub?** This boils down to whether Git diffs will be informative and whether GitHub has nice handling for the file type.
-Small-to-medium plain text files with hard line breaks are a perfect fit, but there are a few more pleasant surprises.
+**Will it play nicely with Git/GitHub?** This boils down to whether Git diffs will be informative and whether GitHub has nice handling for the file type.
+Small-to-medium plain text files with hard line breaks are ideal, but there are a few more pleasant surprises.
 
@@
 **Will it cause problems with Git/GitHub?** This boils down to whether Git diffs are informative, whether GitHub can render the file usefully, and the file's likely effect on Git operations.
-A file that is large and changing quite often can make your repository bloated and slow down pushes and pulls.
+A file that is large and changing often can make your repository bloated and slow down pushes and pulls.
 If a file is binary, such as a Word document or Excel spreadsheet, you will not get human-readable diffs, nor can GitHub display it in the browser.
 Html is not rendered on GitHub, but is, instead, displayed in raw form.
-Binary files are also a prime source of merge conflicts, because they are beyond the reach of Git's sophisticated automatic merging logic.
-For many of us, adoption of Git/GitHub should prompt a pivot away from `.docx` and `.pdf` as primary output formats and towards `.md`, at least during periods of rapid development.
+Binary files are also a prime source of merge conflicts (see below), because they are beyond the reach of Git's sophisticated automatic merging logic.
+This all implies that adoption of Git/GitHub should prompt a pivot away from `.docx` and `.pdf` as primary output formats and towards `.md`, at least during periods of rapid development.
 
@@
 
-Collaboration is probably the most compelling reason to manage a project with Git and GitHub.
-I have a broad definition of collaboration.
-It includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
+Collaboration is the most compelling reason to manage a project with Git and GitHub.
+My definition of collaboration includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
 
@@
 *once I understand Nick's point about binary files causing merge conflicts <https://github.com/dsscollection/git-github-for-stats/issues/6> I may need to reword the above*
-
-*something I would like to write about, but have no space for here: different models of collaboration, i.e.
-everyone working on one repo & one branch vs.
-making pull requests from branches in the main repo or from forks*
 
@@
 
+instructions re: supplementary materials
+http://amstat.tfjournals.com/supplementary-materials/
+
+TAS author instructions
+http://www.tandfonline.com/action/authorSubmission?journalCode=utas20&page=instructions
+
 ```{r eval = is_github_document, echo = FALSE, results = 'asis'}

### 2017-05-30 README.tex
Re-render

@@
 
-Why would a statistician use version control, such as Git? And what is
-the point of hosting your work online, e.g., on GitHub? Could the gains
+Why would a statistician use version control, such as
+\href{http://git-scm.com}{Git}? And what is the point of hosting your
+work online, e.g., on \href{https://github.com}{GitHub}? Could the gains
 possibly justify the inevitable pain?
@@
   GitHub makes a fantastic course management system for courses that use
-  R. You and your students can exchange actual working code and explore
+  R. You can exchange actual working code with your students and explore
   the associated results.
@@
 \end{itemize}
+
+Now the bad news: Git was built neither for the exact usage described
+here, nor for broad usability. You will undoubtedly notice this, so it's
+best to know in advance. Happily, there are many helpful tools that
+mediate your interactions with Git. GitHub itself is a fine example, as
+is \href{https://www.rstudio.com/products/rstudio/}{RStudio}. In
+addition to pointing out tools that soften Git's sharpest edges, I
+recommend specific habits and attitudes that reduce frustration.
 
@@
 In my opinion, \textbf{for new users}, the pros of Git only outweigh the
-cons when you factor in the overhead of communicating and collaborating
-with other people, including your future self. And who among us does not
-need to do that? Your life is much easier if this is baked into your
-workflow, as opposed to being a separate process that you dread and
-avoid. Communication and collaboration are the killer apps of version
-control. Git enforces a rigorous model of file management, but it is
-critical for the distribution of files across different people,
+cons when you consider the overhead of working with other people,
+including your future self. And who among us does not need to do that?
+In a Git-based workflow, you document and, optionally, expose your work
+as you go. Communication and collaboration are the killer apps of
+version control. Git's model of file management can feel uncomfortably
+rigid, but it enables the distribution of files across different people,
 computers, and time.
 
-This has an implication for selecting your first Git projects: in order
-to see substantial gain for your pain, you need to pick projects that
-require you to share rapidly evolving files with others. It is tempting
-to pick a quiet little project that you work on privately, but you risk
-missing the main point of formal version control.
+This has an implication for selecting your first Git projects: you will
+enjoy the most gain for your pain if you pick a project that involves
+sharing rapidly evolving files with others. It is tempting to pick a
+quiet, private project, but you risk missing out on the main benefits of
+formal version control.
 
@@
 collaborators sync regularly to a common version, acknowledging that the
-difficulty of merging goes up faster than the size of the difference
+difficulty of merging only goes up as more changes accumulate
 \citep{good-enough}. Especially important versions get a human-readable
@@
 This article does not provide step-by-step instructions on how to use
-Git and GitHub. This format would not be effective, but we do provide
-annotated links to such resources in ?the appendix or online
-supplement?. Instead, I'll convey what the workflow feels like and what
-the payoffs are, with special attention to the statistics and R context.
-The goal is to help the Git-curious generate the activation energy
-needed to get started.
+Git and GitHub. This format would not be effective, but annotated links
+to such resources are given in the supplementary materials. Instead,
+I'll convey what the workflow feels like and what the payoffs are, with
+special attention to the statistics and R context. The goal is to help
+the Git-curious generate the activation energy needed to get started.
 
@@
 
-\href{https://github.com}{GitHub} is currently the most popular Git
-hosting service. Others include \href{https://about.gitlab.com}{GitLab}
-and \href{https://bitbucket.org}{Bitbucket}. These services provide a
-home for Git-based projects on the internet. It is a bit like DropBox or
-Google Drive, but more structured, powerful, and programmatic.
+Git offers a powerful structure for file management and
+\href{https://github.com}{GitHub} complements it with a slick user
+interface and distribution mechanism for Git repositories. GitHub is
+currently the most popular Git hosting service, providing a home for
+Git-based projects on the internet. It is a bit like DropBox or Google
+Drive, but more structured, powerful, and programmatic.
 
 The remote host acts as the clearinghouse for a Git-managed project.
-This allows other people to see the project files, sync up with the
-current version, and perhaps even make or propose changes. These hosting
-providers offer well-designed web interfaces that are a dramatic
-improvement over traditional Unix Git servers. Many operations can be
-done entirely in the browser, such as viewing current or past versions
-of a file, commenting on a recent change, and editing or adding files.
-These hosts also offer granular control over who can see, edit, and
-administer a project.
+This allows others to see the project files, sync up with the current
+version, and perhaps even propose or make changes. GitHub has a
+well-designed web interface that is a dramatic improvement over
+traditional Unix Git servers. Many operations can be done entirely in
+the browser, such as viewing current or past versions of a file,
+commenting on a recent change, and editing or adding files. GitHub also
+offers granular control over who can see, edit, and administer a
+project.
 
@@
 
-Here is the initial setup process, which you do once or, for some steps,
-once per computer.
+Here is an overview of the initial setup process, which happens once or,
+for some steps, once per computer. This is excerpted from
+\href{http://happygitwithr.com}{Happy Git and GitHub for the useR},
+which holds battle-tested instructions honed over several years in
+\href{http://stat545.com}{STAT 545} at the University of British
+Columbia.
 
@@
   Install Git. Depending on your OS, Git might already be installed. But
-  many of us may need to install it or will choose to update to a more
-  recent version. Some basic configuration is critical, such as setting
-  your username and email.
+  many may need to install it or will choose to update to a more recent
+  version. Some basic configuration is critical, such as setting your
+  username and email.
 \item
@@
   receive the current version of the repository on GitHub, known as
-  \textbf{pushing} and \textbf{pulling}, respectively. This will require
-  authenticating yourself with GitHub from a local shell or Git client.
-  At this point, most people elect to do a bit of extra setup to ensure
-  that they are not repeatedly challenged for their GitHub credentials
-  going forward.
+  \textbf{pushing} and \textbf{pulling}, respectively.
 \end{itemize}
@@
 Once this setup is done, you are ready to start using Git and GitHub
-with your projects.
-
-\subsection{Is this going to hurt?}\label{is-this-going-to-hurt}
-
-Yes.
-
-Git was built neither for the exact usage described here, nor for broad
-usability. You will undoubtedly notice this, so it can be helpful to
-know this in advance. Happily there are many helpful tools that mediate
-your interactions with Git. GitHub itself is a fine example, as is
-RStudio. In addition to pointing out tools that soften Git's sharpest
-edges, I recommend specific habits and attitudes that reduce
-frustration.
-
-General recommendations for agony reduction:
+with your projects. Some general recommendations for agony reduction:
 
@@
 Recall that a repository or repo is just a directory of files that Git
-will manage holistically. A commit functions like a snapshot of all the
+manages holistically. A commit functions like a snapshot of all the
 files in the repo, at a specific moment. Under the hood, that is not
@@
 
-\textbf{Will it play especially well with Git/GitHub?} This boils down
-to whether Git diffs will be informative and whether GitHub has nice
-handling for the file type. Small-to-medium plain text files with hard
-line breaks are a perfect fit, but there are a few more pleasant
-surprises.
+\textbf{Will it play nicely with Git/GitHub?} This boils down to whether
+Git diffs will be informative and whether GitHub has nice handling for
+the file type. Small-to-medium plain text files with hard line breaks
+are ideal, but there are a few more pleasant surprises.
 
@@
 usefully, and the file's likely effect on Git operations. A file that is
-large and changing quite often can make your repository bloated and slow
-down pushes and pulls. If a file is binary, such as a Word document or
-Excel spreadsheet, you will not get human-readable diffs, nor can GitHub
+large and changing often can make your repository bloated and slow down
+pushes and pulls. If a file is binary, such as a Word document or Excel
+spreadsheet, you will not get human-readable diffs, nor can GitHub
 display it in the browser. Html is not rendered on GitHub, but is,
 instead, displayed in raw form. Binary files are also a prime source of
-merge conflicts, because they are beyond the reach of Git's
-sophisticated automatic merging logic. For many of us, adoption of
+merge conflicts (see below), because they are beyond the reach of Git's
+sophisticated automatic merging logic. This all implies that adoption of
 Git/GitHub should prompt a pivot away from \texttt{.docx} and
@@
 
-Collaboration is probably the most compelling reason to manage a project
-with Git and GitHub. I have a broad definition of collaboration. It
-includes hands-on participation by multiple people as well as an
-asymmetric model, in which some people are active makers and others only
-read or review.
+Collaboration is the most compelling reason to manage a project with Git
+and GitHub. My definition of collaboration includes hands-on
+participation by multiple people as well as an asymmetric model, in
+which some people are active makers and others only read or review.
 
@@
 may need to reword the above}
-
-\emph{something I would like to write about, but have no space for here:
-different models of collaboration, i.e.~everyone working on one repo \&
-one branch vs.~making pull requests from branches in the main repo or
-from forks}
 
@@
 
+instructions re: supplementary materials
+\url{http://amstat.tfjournals.com/supplementary-materials/}
+
+TAS author instructions
+\url{http://www.tandfonline.com/action/authorSubmission?journalCode=utas20\&page=instructions}
+
 \bibliographystyle{agsm}

### 2017-06-29 README.Rmd
Various edits + first 3 figures

@@
 
-Why would a statistician use version control, such as [Git](http://git-scm.com)?
-And what is the point of hosting your work online, e.g., on [GitHub](https://github.com)?
+Why would a statistician use a version control system, such as [Git](http://git-scm.com) [@git]?
+And what is the point of hosting your work online, e.g., on [GitHub](https://github.com) [@github]?
 Could the gains possibly justify the inevitable pain?
@@
 Git manages the evolution of a set of files -- called a __repository__ or 
-__repo__ -- in a sane, highly structured way. It is a bit like the "Track 
+__repo__ -- in a sane, highly structured way. It is like the "Track 
 Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up
@@
 @git-for-humans; @ten-simple-rules-git].
-We use it to manage the motley collection of files that make up typical data analytical projects, which often consist of data, figures, reports, and, yes, source code.
+We use it to manage the motley collection of files that make up typical data analytical projects, which consist of data, figures, reports, and, yes, source code.
 Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.
@@
 Many people who don't use Git unwittingly re-invent a poor man's version of it.
-They take an important file and distribute it via email. Various parties make 
-changes, decorating the file name with initials, dates, and other descriptors. 
-Before you know it, the original file is the root of a complicated phylogeny 
-that no amount of "Track changes" and good intentions can resolve.
+Figure X depicts a hypothetical analysis of the iris data, captured in a single R source file.
+With informal version control, contributors create derivative copies of `iris.R`, decorating the file name with initials, dates, and other descriptors.
+Even when working alone, this leads to multiple versions of `iris.R` of indeterminate relatedness (Fig Xa).
+In collaborative settings based on email distribution, the original file swiftly becomes the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve (Fig Xb).
 
-*If you're looking at the PDF draft, you're missing out on some figures around 
-here in the GitHub Markdown version that hint at the figure I plan to embed.*
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("diy-vs-git-solo-vs-collab.png")
+```
 
-```{r eval = is_github_document, echo = FALSE, results = 'asis'}
+```{r eval = FALSE, echo = FALSE, results = 'asis'}
 cat("*here's one way I see phylogeny in my mind*\n")
@@
 
-The Git way is to track the evolution of that file, through a series of commits,
-each equipped with an explanatory message and a nickname. All collaborators sync
-regularly to a common version, acknowledging that the difficulty of merging only goes up as more changes accumulate [@good-enough]. Especially important 
+The Git way is to track the evolution of `iris.R`, through a series of commits,
+each equipped with an explanatory message and a nickname.
+Figure Xc depicts this sort of linear *in situ* development.
+Figure Xd shows the same history for a common collaborative Git workflow, where contributors work independently but sync regularly to a common version.
+Especially important
 versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
@@
 
-*make a figure related to the above (reference those file names?) but 
-juxtaposing different workflows, inspiration from the Bartlett Git for Designers
-talk*
-
-```{r echo = FALSE, out.width="100%"}
+```{r include = FALSE, echo = FALSE, out.width="100%"}
 knitr::include_graphics("bartlett-copies-vs-commits.png")
@@
 such resources are given in the supplementary materials.
-Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context.
+Instead, I convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context.
 The goal is to help the Git-curious generate the
@@
 
-Git offers a powerful structure for file management and [GitHub](https://github.com) complements it with a slick user interface and distribution mechanism for Git repositories.
-GitHub is currently the most popular Git hosting service, providing a home for Git-based projects on the internet.
-It is a bit like DropBox or Google Drive, but more structured, powerful, and programmatic.
+We've explored Git's powerful structure for file management, so where does GitHub fit in?
+[GitHub](https://github.com) complements Git by providing a slick user interface and distribution mechanism for Git repositories.
+Git is the software you will use locally to record changes to a set of files.
+GitHub is a hosting service that provides a Git-aware home for such projects on the internet.
+It is like DropBox or Google Drive, but more structured, powerful, and programmatic.
 
 The remote host acts as the clearinghouse for a Git-managed project. This allows
-others to see the project files, sync up with the current version, and perhaps even propose or make changes.
-GitHub has a well-designed web interface that is a dramatic improvement over traditional Unix Git servers.
-Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files.
+others to browse project files, explore their history, sync up with the current version, and perhaps even propose or make changes.
+GitHub's well-designed web interface is a dramatic improvement over traditional Unix Git servers.
+Many operations can be done entirely in the browser, including editing or adding files.
+It is easy to create a hyperlink to a specific file or location in a file, at a specific version, which can make meta-conversations about project code or reports much more productive.
 GitHub also offers granular control over who can see, edit, and administer a project.
 
-*possible images include a modified version of this atlassian image (no arrows 
-between people, yes arrow from people to github), somehow connected to a 
-screenshot of a GitHub repo*
-
 ```{r echo = FALSE, out.width="100%"}
-knitr::include_graphics("atlassian-git-tutorials-syncing.png")
+knitr::include_graphics("your-repo-their-repo-central-remote-repo.png")
 ```
@@
 __GitHub issues__ are another powerful feature of the platform. Recall that we 
-are repurposing Git, a tool designed to facilitate software development. The 
-issues for a project are its bug tracker. For projects that are not pure 
-software development, we co-opt this machinery to organize our to do list more 
+are repurposing Git, a tool that facilitates software development. Think of the 
+issues for a project as its bug tracker. For projects that are not pure 
+software development, we co-opt this machinery to organize our to-do list more 
 generally. The basic unit is an issue and you can interact with one in two ways.
@@
 
-Here is an overview of the initial setup process, which happens once or, for some steps, once per computer.
+If I've convinced you to experiment with Git and GitHub, you need to do some initial setup.
+These first steps happen one or, for some steps, once per computer.
 This is excerpted from [Happy Git and GitHub for the useR](http://happygitwithr.com), which holds battle-tested instructions honed over several years in [STAT 545](http://stat545.com) at the University of British Columbia.
@@
 
-* Install Git. Depending on your OS, Git might already be installed. But many may need to install it or will choose to update to a more recent version.
+* Install Git. Depending on your OS, Git might already be installed. But many will need to install it or will choose to update to a more recent version.
 Some basic configuration is critical, such as setting your username and email.
 
-* Install a local Git client *(optional but highly recommended*). A Git client 
-is software that provides a graphical user interface for Git, which is otherwise
+* Install a local Git client, *optional but highly recommended*. A Git client 
+provides a graphical user interface for Git, which is otherwise
 command-line only. If you are an R user, you will find that 
@@
 [SourceTree](https://www.sourcetreeapp.com) or
-[GitKraken](https://www.gitkraken.com). Git is a file-based system, so you can
+[GitKraken](https://www.gitkraken.com). Git just operates on files, so you can
 do some operations from RStudio, others from SourceTree, and others from the
@@
 
-We now explore the fundamental concepts of Git and connect them to the data science workflow:
+We now connect the fundamental concepts of Git to the data science workflow:
 
@@
 
-*this is really crying out for an example and/or diagrams, something that shows commits unfolding and let's you illustrate a diff, sort of like this from the barlett talk*
-
 ```{r echo = FALSE, out.width="100%"}
-knitr::include_graphics("bartlett-commit-history.png")
+knitr::include_graphics("commit-diff-sha-tag.png")
 ```
 
-Consider version A of a file and a modified version, version B.
-Assume that version A was part of a Git commit and version B was part of the next commit.
+Figure X is another look at our fictional analysis of the iris data, focusing on the evolution of its script `iris.R`.
+Consider version A of this file and a modified version, version B.
+Assume that version A was part of one Git commit and version B was part of the next commit.
 The set of differences between A and B is called a "diff" and Git users contemplate diffs a lot.
@@
 
-In fact, Git's notion of version B of your file is as an accumulation of diffs.
-At some earlier point, the file was created in the first place.
-That version of the file was part of a commit and, therefore, a diff.
-Git stores Version A of the file as the initial version, plus all the intervening diffs in the history that affect the file.
-And version B simply includes one more.
-We'll set these internal details aside now, but understanding the importance of these deltas will eventually make Git's operations less baffling.
+In fact, Git's notion of any specific version of `iris.R` is as an accumulation of diffs.
+If you go back far enough, you find the commit where the file was created in the first place.
+Every later version is stored by Git as that initial version, plus all the intervening diffs in the history that affect the file.
+We'll set these internal details aside now, but understanding the importance of these deltas will make Git's operations less baffling in the long run.
 
@@
 When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the __history__, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.
+Figure X shows the messages associated with the last three commits.
 
@@
 Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice.
+The commit messages in Figure X are prefixed by such abbreviated SHAs.
 You can also designate certain snapshots as special with a __tag__, which is a name of your choosing.
@@
 For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.
+Figure X shows a tag, "draft-01", associated with the last commit.
 
@@
 
-This may seem unrelated to Git, GitHub, and R, but we need to talk about [Markdown](https://daringfireball.net/projects/markdown/syntax).
+This may seem unrelated to Git, GitHub, and R, but it is now necessary to talk about [Markdown](https://daringfireball.net/projects/markdown/syntax).
 Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible.
@@
 
-Markdown is extra special for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code.
+Markdown is especially holy for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code.
 Again, do not regard R Markdown as something you must clear your schedule to learn.
@@
 Once the Markdown file is pushed to GitHub, it is as if the reader has run your code or is able to look over your shoulder at your R session.
-This provides an lightweight system for exposing work-in-progress to collaborators, without slowing down to create separate reports.
+This provides a lightweight system for exposing work-in-progress to collaborators, without slowing down to create separate reports.
 Comment lines that begin with `#'` are elevated to top-level prose, providing a way to make the document more welcoming for a reader.
@@
 
-Let's zoom back out again and consider R Markdown and the rmarkdown package more generally.
-I want to point out that R Markdown can be rendered to many more formats than Markdown!
-I have emphasized the production of Markdown here because it is extremely useful in GitHub-hosted projects and that seems to be underappreciated.
-But the rmarkdown package can convert `.Rmd` to a wide array of [output formats](http://rmarkdown.rstudio.com/lesson-9.html), including HTML, PDF, and Word (`.docx`).
-The Markdown must be created regardless.
-Sometimes it is all you need, such as on GitHub.
-In other contexts, it is just a necessary intermediate and may even be discarded.
+Before we move on, I want to zoom out and revisit R Markdown and the rmarkdown package more generally.
+Note that R Markdown can be rendered to many more formats than Markdown, including HTML, PDF, and Microsoft Word, and can incorporate code chunks in many languages other than R.
+Markdown is emphasized here, because it is extraordinarily useful in GitHub-hosted projects.
+Even when targetting another output format, the Markdown must be created as a necessary intermediate.
+Sometimes it is all you neeed, such as on GitHub, and in other contexts it may even be discarded.
 
@@
 A file that Git does not ignore is said to be __tracked__.
-How should you think about this?
+Here's a useful framework for thinking about what to track:
 
@@
 **Is it useful to someone?** If so, track and share!
-There is a taboo against committing derived products, inherited from Git's software development roots, because their typical product is a platform-specific executable.
+There is a taboo against committing derived products, inherited from Git's software development roots, because the typical product in that context is a platform-specific executable.
 This rationale, however, does not apply to many data science products.
@@
 
-  * Some derived files are too miserable to read casually, such as `.csv` files of processed results or `.html` derived from `.Rmd`, but are still worth tracking with Git.
+  * Some derived files are too miserable to read casually, such as `.csv` files of processed results or `.html` derived from `.Rmd`, yet they are still worth tracking with Git.
 When you re-run an analysis with updated input data or after updating R packages, *the diffs are often quite modest* and help you pinpoint unexpected changes.
@@
 It also [displays and provides visual diffs](https://help.github.com/articles/rendering-and-diffing-images/) for the most common image formats, which is extremely useful for spotting unexpected changes in figures.
-  * Never, ever forget that Markdown is highly privileged on GitHub and is, therefore, a crucial output format for rendered `.R` and `.Rmd`.
 
-**Will it cause problems with Git/GitHub?** This boils down to whether Git diffs are informative, whether GitHub can render the file usefully, and the file's likely effect on Git operations.
+**Will it actively cause problems with Git/GitHub?** This boils down to the file's likely effect on Git operations.
 A file that is large and changing often can make your repository bloated and slow down pushes and pulls.
-If a file is binary, such as a Word document or Excel spreadsheet, you will not get human-readable diffs, nor can GitHub display it in the browser.
-Html is not rendered on GitHub, but is, instead, displayed in raw form.
-Binary files are also a prime source of merge conflicts (see below), because they are beyond the reach of Git's sophisticated automatic merging logic.
-This all implies that adoption of Git/GitHub should prompt a pivot away from `.docx` and `.pdf` as primary output formats and towards `.md`, at least during periods of rapid development.
+If a file is binary, such as a Word document or Excel spreadsheet, you will not get human-readable diffs anyway, nor can GitHub display the content in the browser.
+Binary files are also a reliable source of merge conflicts (see below), because they are beyond the reach of Git's sophisticated automatic merging logic.
+A large binary file that changes often is, therefore, the worst of all worlds.
+This implies that adoption of Git/GitHub rewards a pivot away from `.docx`, `.xlsx`, and `.pdf` as primary file formats and towards `.Rmd`, `.md`, and `.csv`, at least during periods of rapid development.
 
@@
 Collaboration is the most compelling reason to manage a project with Git and GitHub.
-My definition of collaboration includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.
+My definition of collaboration includes hands-on participation by multiple people, including your past and future self, as well as an asymmetric model, in which some people are active makers and others only read or review.
 
-Consider two different models of collaboration on a document:
+Consider two different ways to collaborate on a document:
 
-  * **Edit, save, attach.** In this workflow, everyone has one (or more!) copies of the document and they circulate via email attachment, accumulating initials and dates in the filename.
+  * **Edit, save, attach.** In this workflow, everyone has one (or more!) copies of the document, which circulate as email attachments, accumulating initials and dates in the filename.
 Which one is "master"?
-Does this question even make sense anymore?
-How do different versions of the document relate to each other?
-If you want to see a version combining the best versions of each section, how would you reconcile the different copies into one?
+Does this question even make sense anymore?!?
+If you want a version combining the edits made by different authors to different sections, how do you reconcile the copies into one?
 All of this usually gets sorted out by social contract, a fairly manual process, and at least one miserable person.
@@
 Managing a project via Git/GitHub is much more like the Google Doc scenario, but also offers some of the attractive features of "edit, save, attach".
-With Git/GitHub, collaborators can work offline and there can be independent lines of development.
-The real power comes from regular and structured reconciliation of all versions of all the files in the project via Git/GitHub.
+With Git/GitHub, collaborators can work offline and there can be independent lines of development, i.e.
+branches.
+The killer feature is that Git/GitHub enable regular and structured reconciliation of all versions of all the files in the project.
 It is definitely more complicated than collaborating on a Google Doc, but also more powerful.
@@
 Git is a decentralized version control system, meaning each collaborator has their own complete copy of the repo and its history.
-Everyone can work offline and/or simultaneously, but with regular syncing to GitHub.
+Everyone can work offline and/or simultaneously.
 GitHub plays the role of another collaborator, but a very special one.
-By convention, everyone agrees that GitHub keeps the master copy of the project.
-GitHub is the clearinghouse.
+By convention, everyone agrees that GitHub is the clearinghouse, i.e.
+it holds the master copy of the project.
 The joke is that GitHub puts the "central" in decentralized version control.
@@
 
-*These two points were originally written as comparisons with Google Docs, but now there are intervening paragraphs. Is it too disconnected? Is it worth weaving in?*
+What if two people have made changes to the repository?
+Imagine that your collaborator makes a change to a file, commits it locally, and pushes to GitHub.
+Meanwhile you also make a different change to the same file and also commit locally.
+When you try to push your commit to GitHub, you will fail because there are commits on GitHub that you do not have.
+You must pull from GitHub.
+The good news is that quite often, this will "just work", i.e.
+the GitHub version and your version will merge cleanly.
+Git is quite clever at reconciliation and changes to different files or even distinct parts of the same file will merge.
+This derives from the "diff" based model of Git described earlier.
+After a successful merge, you can push your changes and the cycle goes on.
 
-*Manage multiple files*.
-A Git repository is inherently multi-file and therefore well suited to projects comprised of many files, evolving in a coordinated fashion.
-Examples include a data analysis, a course website, a blog, an R package, or a book.
-If there is any way to proactively check or enforce their joint functionality, this is something you could verify manually prior to a commit or at certain milestones.
-In the case of a website, you might choose to rebuild the site prior to a commit.
-In the case of an R package, `R CMD check` is one of the easiest things to automate on GitHub.
-
-*Diffs and time travel*.
-Google Docs are fantastic for simple collaborative work, when you don't need detailed access to the history.
-But the version control offered by Google Drive is very limited compared to Git.
-You can't compare versions at arbitrary points in time, temporarily checkout previous versions, or maintain two lines of development.
-
-*May need a segue ... depends on the fate of the two paragraphs above.*
-
+But sometimes it's not clear how to reconcile your changes with the new ones from GitHub and you get a *merge conflict*.
 Merge conflicts are the most frustrating thing about using Git and GitHub.
@@
 What is a merge conflict?
-Here is a typical first encounter: your collaborator makes a change to a file, commits it locally, and pushes to GitHub.
-Meanwhile you also make a different change to the same file and also commit locally.
-When you try to push your commit to GitHub, it will fail because there are commits on GitHub that you do not have.
-You must pull from GitHub.
-The good news is that quite often, this will "just work", i.e.
-the GitHub version and your version will merge cleanly.
-Git is quite clever at reconciliation and changes to different files or even distinct parts of the same file will merge.
-If this pull and merge goes smoothly, you'll able to push your changes and the cycle goes on.
+It happens when Git can't be certain how to jointly apply the diffs from two different commits to their common parent.
+At each location of conflict, you must pick one version or the other -- or create a hybrid -- and mark it as resolved.
+Many Git clients have special tooling for this specific task, which can be very convenient.
+Once you've resolved all conflicts, you will be able to finalize the merge and push a version integrating your recent changes to GitHub.
 
-But sometimes it's not clear how to reconcile your changes with the new ones from GitHub and you get a *merge conflict*.
-You must inspect the locations of conflict, which Git marks for you.
-You will pick one version or the other -- or create a hybrid -- at each location of conflict and mark it as resolved.
-Once you've resolved all conflicts, you will be able to push a version integrating your recent changes to GitHub.
-
-The best way to deal with merge conflicts is to avoid them altogether.
+The best way to deal with merge conflicts is to prevent them.
 This is another reason for all parties to commit, pull, and push often.
@@
 So make lots of small commits, sync regularly with GitHub, and only track binary files with good reason.
-
-*once I understand Nick's point about binary files causing merge conflicts <https://github.com/dsscollection/git-github-for-stats/issues/6> I may need to reword the above*
 
@@
 
-GitHub offers several ways to host a proper website directly from a repository, collectively known as [GitHub Pages](https://help.github.com/categories/github-pages-basics/).
+GitHub also offers several ways to host a proper website directly from a repository, collectively known as [GitHub Pages](https://help.github.com/categories/github-pages-basics/).
 At one extreme, as long as you've got one Markdown file, [GitHub Pages can create a simple website](https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3) automatically.
@@
 
-**Use conventional file extensions.** GitHub is very code-focused and will apply proper syntax high-lighting for almost any language you can think of, if you use one of the standard file extensions.
+**Use conventional file extensions.** GitHub is very code-aware and will apply proper syntax highlighting for almost any language you can think of, if you use one of the standard file extensions.
 This also has advantages for people searching GitHub and trying to filter by language.
@@
 
+http://mynameismichelle.com/git-frost/
+
+*These two points were originally written as comparisons with Google Docs, but now there are intervening paragraphs. I think I will cut them.*
+
+*Manage multiple files*.
+A Git repository is inherently multi-file and therefore well suited to projects comprised of many files, evolving in a coordinated fashion.
+Examples include a data analysis, a course website, a blog, an R package, or a book.
+If there is any way to proactively check or enforce their joint functionality, this is something you could verify manually prior to a commit or at certain milestones.
+In the case of a website, you might choose to rebuild the site prior to a commit.
+In the case of an R package, `R CMD check` is one of the easiest things to automate on GitHub.
+
+*Diffs and time travel*.
+Google Docs are fantastic for simple collaborative work, when you don't need detailed access to the history.
+But the version control offered by Google Drive is very limited compared to Git.
+You can't compare versions at arbitrary points in time, temporarily checkout previous versions, or maintain two lines of development.
+
+*May need a segue ... depends on the fate of the two paragraphs above.*
+
 ```{r eval = is_github_document, echo = FALSE, results = 'asis'}

### 2017-06-29 README.tex
Various edits + first 3 figures

@@
 
-Why would a statistician use version control, such as
-\href{http://git-scm.com}{Git}? And what is the point of hosting your
-work online, e.g., on \href{https://github.com}{GitHub}? Could the gains
-possibly justify the inevitable pain?
+Why would a statistician use a version control system, such as
+\href{http://git-scm.com}{Git} \citep{git}? And what is the point of
+hosting your work online, e.g., on \href{https://github.com}{GitHub}
+\citep{github}? Could the gains possibly justify the inevitable pain?
 
@@
 called a \textbf{repository} or \textbf{repo} -- in a sane, highly
-structured way. It is a bit like the ``Track Changes'' feature from
-Microsoft Word, but more rigorous, powerful, and scaled up to multiple
-files.
+structured way. It is like the ``Track Changes'' feature from Microsoft
+Word, but more rigorous, powerful, and scaled up to multiple files.
 
@@
 manage the motley collection of files that make up typical data
-analytical projects, which often consist of data, figures, reports, and,
-yes, source code. Even those who identify more as statistician than data
+analytical projects, which consist of data, figures, reports, and, yes,
+source code. Even those who identify more as statistician than data
 scientist generally have a similar mix of files that are the artifacts
@@
 Many people who don't use Git unwittingly re-invent a poor man's version
-of it. They take an important file and distribute it via email. Various
-parties make changes, decorating the file name with initials, dates, and
-other descriptors. Before you know it, the original file is the root of
-a complicated phylogeny that no amount of ``Track changes'' and good
-intentions can resolve.
+of it. Figure X depicts a hypothetical analysis of the iris data,
+captured in a single R source file. With informal version control,
+contributors create derivative copies of \texttt{iris.R}, decorating the
+file name with initials, dates, and other descriptors. Even when working
+alone, this leads to multiple versions of \texttt{iris.R} of
+indeterminate relatedness (Fig Xa). In collaborative settings based on
+email distribution, the original file swiftly becomes the root of a
+complicated phylogeny that no amount of ``Track changes'' and good
+intentions can resolve (Fig Xb).
 
-\emph{If you're looking at the PDF draft, you're missing out on some
-figures around here in the GitHub Markdown version that hint at the
-figure I plan to embed.}
+\includegraphics[width=1\linewidth]{diy-vs-git-solo-vs-collab}
 
-The Git way is to track the evolution of that file, through a series of
-commits, each equipped with an explanatory message and a nickname. All
-collaborators sync regularly to a common version, acknowledging that the
-difficulty of merging only goes up as more changes accumulate
-\citep{good-enough}. Especially important versions get a human-readable
+The Git way is to track the evolution of \texttt{iris.R}, through a
+series of commits, each equipped with an explanatory message and a
+nickname. Figure Xc depicts this sort of linear \emph{in situ}
+development. Figure Xd shows the same history for a common collaborative
+Git workflow, where contributors work independently but sync regularly
+to a common version. Especially important versions get a human-readable
 tag, to signal a meaningful milestone. Yes, there is some pain in
 adopting the formalism of Git, but it is worth it.
-
-\emph{make a figure related to the above (reference those file names?)
-but juxtaposing different workflows, inspiration from the Bartlett Git
-for Designers talk}
-
-\includegraphics[width=1\linewidth]{bartlett-copies-vs-commits}
 
@@
 Git and GitHub. This format would not be effective, but annotated links
-to such resources are given in the supplementary materials. Instead,
-I'll convey what the workflow feels like and what the payoffs are, with
+to such resources are given in the supplementary materials. Instead, I
+convey what the workflow feels like and what the payoffs are, with
 special attention to the statistics and R context. The goal is to help
@@
 
-Git offers a powerful structure for file management and
-\href{https://github.com}{GitHub} complements it with a slick user
-interface and distribution mechanism for Git repositories. GitHub is
-currently the most popular Git hosting service, providing a home for
-Git-based projects on the internet. It is a bit like DropBox or Google
+We've explored Git's powerful structure for file management, so where
+does GitHub fit in? \href{https://github.com}{GitHub} complements Git by
+providing a slick user interface and distribution mechanism for Git
+repositories. Git is the software you will use locally to record changes
+to a set of files. GitHub is a hosting service that provides a Git-aware
+home for such projects on the internet. It is like DropBox or Google
 Drive, but more structured, powerful, and programmatic.
@@
 The remote host acts as the clearinghouse for a Git-managed project.
-This allows others to see the project files, sync up with the current
-version, and perhaps even propose or make changes. GitHub has a
-well-designed web interface that is a dramatic improvement over
+This allows others to browse project files, explore their history, sync
+up with the current version, and perhaps even propose or make changes.
+GitHub's well-designed web interface is a dramatic improvement over
 traditional Unix Git servers. Many operations can be done entirely in
-the browser, such as viewing current or past versions of a file,
-commenting on a recent change, and editing or adding files. GitHub also
-offers granular control over who can see, edit, and administer a
-project.
+the browser, including editing or adding files. It is easy to create a
+hyperlink to a specific file or location in a file, at a specific
+version, which can make meta-conversations about project code or reports
+much more productive. GitHub also offers granular control over who can
+see, edit, and administer a project.
 
-\emph{possible images include a modified version of this atlassian image
-(no arrows between people, yes arrow from people to github), somehow
-connected to a screenshot of a GitHub repo}
-
-\includegraphics[width=1\linewidth]{atlassian-git-tutorials-syncing}
+\includegraphics[width=1\linewidth]{your-repo-their-repo-central-remote-repo}
 
@@
 \textbf{GitHub issues} are another powerful feature of the platform.
-Recall that we are repurposing Git, a tool designed to facilitate
-software development. The issues for a project are its bug tracker. For
+Recall that we are repurposing Git, a tool that facilitates software
+development. Think of the issues for a project as its bug tracker. For
 projects that are not pure software development, we co-opt this
-machinery to organize our to do list more generally. The basic unit is
+machinery to organize our to-do list more generally. The basic unit is
 an issue and you can interact with one in two ways.
@@
 
-Here is an overview of the initial setup process, which happens once or,
-for some steps, once per computer. This is excerpted from
+If I've convinced you to experiment with Git and GitHub, you need to do
+some initial setup. These first steps happen one or, for some steps,
+once per computer. This is excerpted from
 \href{http://happygitwithr.com}{Happy Git and GitHub for the useR},
@@
   Install Git. Depending on your OS, Git might already be installed. But
-  many may need to install it or will choose to update to a more recent
+  many will need to install it or will choose to update to a more recent
   version. Some basic configuration is critical, such as setting your
@@
 \item
-  Install a local Git client \emph{(optional but highly recommended}). A
-  Git client is software that provides a graphical user interface for
-  Git, which is otherwise command-line only. If you are an R user, you
-  will find that
+  Install a local Git client, \emph{optional but highly recommended}. A
+  Git client provides a graphical user interface for Git, which is
+  otherwise command-line only. If you are an R user, you will find that
   \href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio}
@@
   \href{https://www.sourcetreeapp.com}{SourceTree} or
-  \href{https://www.gitkraken.com}{GitKraken}. Git is a file-based
-  system, so you can do some operations from RStudio, others from
+  \href{https://www.gitkraken.com}{GitKraken}. Git just operates on
+  files, so you can do some operations from RStudio, others from
   SourceTree, and others from the shell.
@@
 
-We now explore the fundamental concepts of Git and connect them to the
-data science workflow:
+We now connect the fundamental concepts of Git to the data science
+workflow:
 
@@
 
-\emph{this is really crying out for an example and/or diagrams,
-something that shows commits unfolding and let's you illustrate a diff,
-sort of like this from the barlett talk}
+\includegraphics[width=1\linewidth]{commit-diff-sha-tag}
 
-\includegraphics[width=1\linewidth]{bartlett-commit-history}
+Figure X is another look at our fictional analysis of the iris data,
+focusing on the evolution of its script \texttt{iris.R}. Consider
+version A of this file and a modified version, version B. Assume that
+version A was part of one Git commit and version B was part of the next
+commit. The set of differences between A and B is called a ``diff'' and
+Git users contemplate diffs a lot. Diff inspection is how you re-explain
+to yourself how version A differs from version B. Diff inspection is not
+limited to adjacent commits. You can inspect the diffs between any two
+commits.
 
-Consider version A of a file and a modified version, version B. Assume
-that version A was part of a Git commit and version B was part of the
-next commit. The set of differences between A and B is called a ``diff''
-and Git users contemplate diffs a lot. Diff inspection is how you
-re-explain to yourself how version A differs from version B. Diff
-inspection is not limited to adjacent commits. You can inspect the diffs
-between any two commits.
-
-In fact, Git's notion of version B of your file is as an accumulation of
-diffs. At some earlier point, the file was created in the first place.
-That version of the file was part of a commit and, therefore, a diff.
-Git stores Version A of the file as the initial version, plus all the
-intervening diffs in the history that affect the file. And version B
-simply includes one more. We'll set these internal details aside now,
-but understanding the importance of these deltas will eventually make
-Git's operations less baffling.
+In fact, Git's notion of any specific version of \texttt{iris.R} is as
+an accumulation of diffs. If you go back far enough, you find the commit
+where the file was created in the first place. Every later version is
+stored by Git as that initial version, plus all the intervening diffs in
+the history that affect the file. We'll set these internal details aside
+now, but understanding the importance of these deltas will make Git's
+operations less baffling in the long run.
 
@@
 \textbf{history}, by reading commit messages and skimming through diffs,
-is an extremely efficient way to get up to speed.
+is an extremely efficient way to get up to speed. Figure X shows the
+messages associated with the last three commits.
 
@@
 exposed to these, you don't have to handle them directly very often and,
-when you do, usually the first 7 characters suffice. You can also
+when you do, usually the first 7 characters suffice. The commit messages
+in Figure X are prefixed by such abbreviated SHAs. You can also
 designate certain snapshots as special with a \textbf{tag}, which is a
@@
 analytical project, you might tag the version submitted to a journal or
-transmitted to external collaborators.
+transmitted to external collaborators. Figure X shows a tag,
+``draft-01'', associated with the last commit.
 
@@
 
-This may seem unrelated to Git, GitHub, and R, but we need to talk about
+This may seem unrelated to Git, GitHub, and R, but it is now necessary
+to talk about
 \href{https://daringfireball.net/projects/markdown/syntax}{Markdown}.
@@
 
-Markdown is extra special for R users because of
+Markdown is especially holy for R users because of
 \href{http://rmarkdown.rstudio.com}{R Markdown}, which is just Markdown
@@
 is as if the reader has run your code or is able to look over your
-shoulder at your R session. This provides an lightweight system for
+shoulder at your R session. This provides a lightweight system for
 exposing work-in-progress to collaborators, without slowing down to
@@
 
-Let's zoom back out again and consider R Markdown and the rmarkdown
-package more generally. I want to point out that R Markdown can be
-rendered to many more formats than Markdown! I have emphasized the
-production of Markdown here because it is extremely useful in
-GitHub-hosted projects and that seems to be underappreciated. But the
-rmarkdown package can convert \texttt{.Rmd} to a wide array of
-\href{http://rmarkdown.rstudio.com/lesson-9.html}{output formats},
-including HTML, PDF, and Word (\texttt{.docx}). The Markdown must be
-created regardless. Sometimes it is all you need, such as on GitHub. In
-other contexts, it is just a necessary intermediate and may even be
+Before we move on, I want to zoom out and revisit R Markdown and the
+rmarkdown package more generally. Note that R Markdown can be rendered
+to many more formats than Markdown, including HTML, PDF, and Microsoft
+Word, and can incorporate code chunks in many languages other than R.
+Markdown is emphasized here, because it is extraordinarily useful in
+GitHub-hosted projects. Even when targetting another output format, the
+Markdown must be created as a necessary intermediate. Sometimes it is
+all you neeed, such as on GitHub, and in other contexts it may even be
 discarded.
@@
 repository. A file that Git does not ignore is said to be
-\textbf{tracked}. How should you think about this?
+\textbf{tracked}. Here's a useful framework for thinking about what to
+track:
 
@@
 taboo against committing derived products, inherited from Git's software
-development roots, because their typical product is a platform-specific
-executable. This rationale, however, does not apply to many data science
-products. Rendered reports, figures, and cleaned data are often
-extremely valuable to others. Make them readily available.
+development roots, because the typical product in that context is a
+platform-specific executable. This rationale, however, does not apply to
+many data science products. Rendered reports, figures, and cleaned data
+are often extremely valuable to others. Make them readily available.
 
@@
   \texttt{.csv} files of processed results or \texttt{.html} derived
-  from \texttt{.Rmd}, but are still worth tracking with Git. When you
-  re-run an analysis with updated input data or after updating R
+  from \texttt{.Rmd}, yet they are still worth tracking with Git. When
+  you re-run an analysis with updated input data or after updating R
   packages, \emph{the diffs are often quite modest} and help you
@@
   extremely useful for spotting unexpected changes in figures.
-\item
-  Never, ever forget that Markdown is highly privileged on GitHub and
-  is, therefore, a crucial output format for rendered \texttt{.R} and
-  \texttt{.Rmd}.
 \end{itemize}
 
-\textbf{Will it cause problems with Git/GitHub?} This boils down to
-whether Git diffs are informative, whether GitHub can render the file
-usefully, and the file's likely effect on Git operations. A file that is
-large and changing often can make your repository bloated and slow down
-pushes and pulls. If a file is binary, such as a Word document or Excel
-spreadsheet, you will not get human-readable diffs, nor can GitHub
-display it in the browser. Html is not rendered on GitHub, but is,
-instead, displayed in raw form. Binary files are also a prime source of
-merge conflicts (see below), because they are beyond the reach of Git's
-sophisticated automatic merging logic. This all implies that adoption of
-Git/GitHub should prompt a pivot away from \texttt{.docx} and
-\texttt{.pdf} as primary output formats and towards \texttt{.md}, at
+\textbf{Will it actively cause problems with Git/GitHub?} This boils
+down to the file's likely effect on Git operations. A file that is large
+and changing often can make your repository bloated and slow down pushes
+and pulls. If a file is binary, such as a Word document or Excel
+spreadsheet, you will not get human-readable diffs anyway, nor can
+GitHub display the content in the browser. Binary files are also a
+reliable source of merge conflicts (see below), because they are beyond
+the reach of Git's sophisticated automatic merging logic. A large binary
+file that changes often is, therefore, the worst of all worlds. This
+implies that adoption of Git/GitHub rewards a pivot away from
+\texttt{.docx}, \texttt{.xlsx}, and \texttt{.pdf} as primary file
+formats and towards \texttt{.Rmd}, \texttt{.md}, and \texttt{.csv}, at
 least during periods of rapid development.
@@
 and GitHub. My definition of collaboration includes hands-on
-participation by multiple people as well as an asymmetric model, in
-which some people are active makers and others only read or review.
+participation by multiple people, including your past and future self,
+as well as an asymmetric model, in which some people are active makers
+and others only read or review.
 
-Consider two different models of collaboration on a document:
+Consider two different ways to collaborate on a document:
 
@@
   \textbf{Edit, save, attach.} In this workflow, everyone has one (or
-  more!) copies of the document and they circulate via email attachment,
+  more!) copies of the document, which circulate as email attachments,
   accumulating initials and dates in the filename. Which one is
-  ``master''? Does this question even make sense anymore? How do
-  different versions of the document relate to each other? If you want
-  to see a version combining the best versions of each section, how
-  would you reconcile the different copies into one? All of this usually
-  gets sorted out by social contract, a fairly manual process, and at
-  least one miserable person.
+  ``master''? Does this question even make sense anymore?!? If you want
+  a version combining the edits made by different authors to different
+  sections, how do you reconcile the copies into one? All of this
+  usually gets sorted out by social contract, a fairly manual process,
+  and at least one miserable person.
 \item
@@
 save, attach''. With Git/GitHub, collaborators can work offline and
-there can be independent lines of development. The real power comes from
-regular and structured reconciliation of all versions of all the files
-in the project via Git/GitHub. It is definitely more complicated than
-collaborating on a Google Doc, but also more powerful.
+there can be independent lines of development, i.e.~branches. The killer
+feature is that Git/GitHub enable regular and structured reconciliation
+of all versions of all the files in the project. It is definitely more
+complicated than collaborating on a Google Doc, but also more powerful.
 
@@
 has their own complete copy of the repo and its history. Everyone can
-work offline and/or simultaneously, but with regular syncing to GitHub.
-GitHub plays the role of another collaborator, but a very special one.
-By convention, everyone agrees that GitHub keeps the master copy of the
-project. GitHub is the clearinghouse. The joke is that GitHub puts the
-``central'' in decentralized version control. You pull regularly from
-GitHub, to receive and integrate changes made by your collaborators. You
-also push regularly to GitHub, to return the favor, and to maintain its
-status as the comprehensive, authoritative version of the project.
+work offline and/or simultaneously. GitHub plays the role of another
+collaborator, but a very special one. By convention, everyone agrees
+that GitHub is the clearinghouse, i.e.~it holds the master copy of the
+project. The joke is that GitHub puts the ``central'' in decentralized
+version control. You pull regularly from GitHub, to receive and
+integrate changes made by your collaborators. You also push regularly to
+GitHub, to return the favor, and to maintain its status as the
+comprehensive, authoritative version of the project.
 
-\emph{These two points were originally written as comparisons with
-Google Docs, but now there are intervening paragraphs. Is it too
-disconnected? Is it worth weaving in?}
+What if two people have made changes to the repository? Imagine that
+your collaborator makes a change to a file, commits it locally, and
+pushes to GitHub. Meanwhile you also make a different change to the same
+file and also commit locally. When you try to push your commit to
+GitHub, you will fail because there are commits on GitHub that you do
+not have. You must pull from GitHub. The good news is that quite often,
+this will ``just work'', i.e.~the GitHub version and your version will
+merge cleanly. Git is quite clever at reconciliation and changes to
+different files or even distinct parts of the same file will merge. This
+derives from the ``diff'' based model of Git described earlier. After a
+successful merge, you can push your changes and the cycle goes on.
 
-\emph{Manage multiple files}. A Git repository is inherently multi-file
-and therefore well suited to projects comprised of many files, evolving
-in a coordinated fashion. Examples include a data analysis, a course
-website, a blog, an R package, or a book. If there is any way to
-proactively check or enforce their joint functionality, this is
-something you could verify manually prior to a commit or at certain
-milestones. In the case of a website, you might choose to rebuild the
-site prior to a commit. In the case of an R package,
-\texttt{R\ CMD\ check} is one of the easiest things to automate on
+But sometimes it's not clear how to reconcile your changes with the new
+ones from GitHub and you get a \emph{merge conflict}. Merge conflicts
+are the most frustrating thing about using Git and GitHub. You can avoid
+them if you only work alone, on one computer, but I've also said that
+collaboration is the best reason to use GitHub! So this problem must be
+confronted.
+
+What is a merge conflict? It happens when Git can't be certain how to
+jointly apply the diffs from two different commits to their common
+parent. At each location of conflict, you must pick one version or the
+other -- or create a hybrid -- and mark it as resolved. Many Git clients
+have special tooling for this specific task, which can be very
+convenient. Once you've resolved all conflicts, you will be able to
+finalize the merge and push a version integrating your recent changes to
 GitHub.
 
-\emph{Diffs and time travel}. Google Docs are fantastic for simple
-collaborative work, when you don't need detailed access to the history.
-But the version control offered by Google Drive is very limited compared
-to Git. You can't compare versions at arbitrary points in time,
-temporarily checkout previous versions, or maintain two lines of
-development.
-
-\emph{May need a segue \ldots{} depends on the fate of the two
-paragraphs above.}
-
-Merge conflicts are the most frustrating thing about using Git and
-GitHub. You can avoid them if you only work alone, on one computer, but
-I've also said that collaboration is the best reason to use GitHub! So
-this problem must be confronted.
-
-What is a merge conflict? Here is a typical first encounter: your
-collaborator makes a change to a file, commits it locally, and pushes to
-GitHub. Meanwhile you also make a different change to the same file and
-also commit locally. When you try to push your commit to GitHub, it will
-fail because there are commits on GitHub that you do not have. You must
-pull from GitHub. The good news is that quite often, this will ``just
-work'', i.e.~the GitHub version and your version will merge cleanly. Git
-is quite clever at reconciliation and changes to different files or even
-distinct parts of the same file will merge. If this pull and merge goes
-smoothly, you'll able to push your changes and the cycle goes on.
-
-But sometimes it's not clear how to reconcile your changes with the new
-ones from GitHub and you get a \emph{merge conflict}. You must inspect
-the locations of conflict, which Git marks for you. You will pick one
-version or the other -- or create a hybrid -- at each location of
-conflict and mark it as resolved. Once you've resolved all conflicts,
-you will be able to push a version integrating your recent changes to
-GitHub.
-
-The best way to deal with merge conflicts is to avoid them altogether.
-This is another reason for all parties to commit, pull, and push often.
-Small changes, integrated frequently, in non-binary files, are the
-easiest for Git to automatically merge for you. The difficulty of
-merging (by Git or by you) is proportional to the evolutionary distance
-between two lines of work. The presence of frequently-changing binary
-files also increases the burden. So make lots of small commits, sync
-regularly with GitHub, and only track binary files with good reason.
-
-\emph{once I understand Nick's point about binary files causing merge
-conflicts
-\url{https://github.com/dsscollection/git-github-for-stats/issues/6} I
-may need to reword the above}
+The best way to deal with merge conflicts is to prevent them. This is
+another reason for all parties to commit, pull, and push often. Small
+changes, integrated frequently, in non-binary files, are the easiest for
+Git to automatically merge for you. The difficulty of merging (by Git or
+by you) is proportional to the evolutionary distance between two lines
+of work. The presence of frequently-changing binary files also increases
+the burden. So make lots of small commits, sync regularly with GitHub,
+and only track binary files with good reason.
 
@@
 
-GitHub offers several ways to host a proper website directly from a
+GitHub also offers several ways to host a proper website directly from a
 repository, collectively known as
@@
 
-\textbf{Use conventional file extensions.} GitHub is very code-focused
-and will apply proper syntax high-lighting for almost any language you
-can think of, if you use one of the standard file extensions. This also
-has advantages for people searching GitHub and trying to filter by
-language.
+\textbf{Use conventional file extensions.} GitHub is very code-aware and
+will apply proper syntax highlighting for almost any language you can
+think of, if you use one of the standard file extensions. This also has
+advantages for people searching GitHub and trying to filter by language.
 
@@
 
+\url{http://mynameismichelle.com/git-frost/}
+
+\emph{These two points were originally written as comparisons with
+Google Docs, but now there are intervening paragraphs. I think I will
+cut them.}
+
+\emph{Manage multiple files}. A Git repository is inherently multi-file
+and therefore well suited to projects comprised of many files, evolving
+in a coordinated fashion. Examples include a data analysis, a course
+website, a blog, an R package, or a book. If there is any way to
+proactively check or enforce their joint functionality, this is
+something you could verify manually prior to a commit or at certain
+milestones. In the case of a website, you might choose to rebuild the
+site prior to a commit. In the case of an R package,
+\texttt{R\ CMD\ check} is one of the easiest things to automate on
+GitHub.
+
+\emph{Diffs and time travel}. Google Docs are fantastic for simple
+collaborative work, when you don't need detailed access to the history.
+But the version control offered by Google Drive is very limited compared
+to Git. You can't compare versions at arbitrary points in time,
+temporarily checkout previous versions, or maintain two lines of
+development.
+
+\emph{May need a segue \ldots{} depends on the fate of the two
+paragraphs above.}
+
 \bibliographystyle{agsm}

### 2017-06-29 README.Rmd
Figure infrastructure: captions, references, etc.

@@
 ---
+
+```{r knitr_options, include = FALSE}
+knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
+```
 
@@
 Many people who don't use Git unwittingly re-invent a poor man's version of it.
-Figure X depicts a hypothetical analysis of the iris data, captured in a single R source file.
+Figure \ref{fig:diy-vs-git} depicts a hypothetical analysis of the iris data, captured in a single R source file.
 With informal version control, contributors create derivative copies of `iris.R`, decorating the file name with initials, dates, and other descriptors.
-Even when working alone, this leads to multiple versions of `iris.R` of indeterminate relatedness (Fig Xa).
-In collaborative settings based on email distribution, the original file swiftly becomes the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve (Fig Xb).
+Even when working alone, this leads to multiple versions of `iris.R` of indeterminate relatedness (Figure \ref{fig:diy-vs-git}A).
+In collaborative settings based on email distribution, the original file swiftly becomes the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve Figure \ref{fig:diy-vs-git}B).
 
-```{r echo = FALSE, out.width="100%"}
+```{r diy-vs-git, fig.cap="\\label{fig:diy-vs-git}A: Solo work with DIY version control via filename.
+B: Collaborative work with DIY version control.
+C: Solo work with Git.
+D: Collaborative work with Git."}
 knitr::include_graphics("diy-vs-git-solo-vs-collab.png")
@@
 
-```{r eval = FALSE, echo = FALSE, results = 'asis'}
-cat("*here's one way I see phylogeny in my mind*\n")
-cat("![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)")
-
-cat("\n\n*make such a phylogeny with a set of crazy filenames like this*\n") 
-cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")
-
-cat("\n\n*or riff on the famous PhD Comic*") 
-cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
-```
-
 The Git way is to track the evolution of `iris.R`, through a series of commits,
-each equipped with an explanatory message and a nickname.
-Figure Xc depicts this sort of linear *in situ* development.
-Figure Xd shows the same history for a common collaborative Git workflow, where contributors work independently but sync regularly to a common version.
+each equipped with an explanatory message.
+Figure \ref{fig:diy-vs-git}C depicts this linear, *in situ* development process.
+Figure \ref{fig:diy-vs-git}D shows the same history for a common collaborative Git workflow, where contributors work independently but sync regularly to a common version.
 Especially important
@@
 is some pain in adopting the formalism of Git, but it is worth it.
-
-```{r include = FALSE, echo = FALSE, out.width="100%"}
-knitr::include_graphics("bartlett-copies-vs-commits.png")
-```
 
@@
 GitHub is a hosting service that provides a Git-aware home for such projects on the internet.
-It is like DropBox or Google Drive, but more structured, powerful, and programmatic.
+These relationships are shown in Figure \ref{fig:yours-theirs-github}.
+GitHub is like DropBox or Google Drive, but more structured, powerful, and programmatic.
 
@@
 
-```{r echo = FALSE, out.width="100%"}
+```{r yours-theirs-github, fig.cap = "\\label{fig:yours-theirs-github}With Git, all contributors have a copy of the repo, with all files and the full history.
+It is typical to stay in sync through the use of a central remote repo, such as GitHub.
+Hosted remotes like GitHub also provide access to the repo through a web browser."}
 knitr::include_graphics("your-repo-their-repo-central-remote-repo.png")
@@
 Under the hood, that is not exactly how Git implements things.
-Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.
+Although mental models don't have to be accurate in order to be useful, in this case it helps to align the two.
 
-```{r echo = FALSE, out.width="100%"}
+```{r commit-diff-sha-tag, fig.cap="\\label{fig:commit-diff-sha-tag}Partial commit history for our iris example."}
 knitr::include_graphics("commit-diff-sha-tag.png")
@@
 
-Figure X is another look at our fictional analysis of the iris data, focusing on the evolution of its script `iris.R`.
+Figure \ref{fig:commit-diff-sha-tag} is another look at our fictional analysis of the iris data, focusing on the evolution of its script, `iris.R`.
 Consider version A of this file and a modified version, version B.
@@
 When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the __history__, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.
-Figure X shows the messages associated with the last three commits.
+Figure \ref{fig:commit-diff-sha-tag} shows the messages associated with the last three commits.
 
@@
 Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice.
-The commit messages in Figure X are prefixed by such abbreviated SHAs.
+The commit messages in Figure \ref{fig:commit-diff-sha-tag}  are prefixed by such truncated SHAs.
 You can also designate certain snapshots as special with a __tag__, which is a name of your choosing.
@@
 For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.
-Figure X shows a tag, "draft-01", associated with the last commit.
+Figure \ref{fig:commit-diff-sha-tag} shows a tag, "draft-01", associated with the last commit.
 
@@
 
+```{r eval = FALSE, echo = FALSE, results = 'asis'}
+cat("*here's one way I see phylogeny in my mind*\n")
+cat("![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)")
+
+cat("\n\n*make such a phylogeny with a set of crazy filenames like this*\n") 
+cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")
+
+cat("\n\n*or riff on the famous PhD Comic*") 
+cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
+```
+
 ```{r eval = is_github_document, echo = FALSE, results = 'asis'}

### 2017-06-29 README.tex
Figure infrastructure: captions, references, etc.

@@
 Many people who don't use Git unwittingly re-invent a poor man's version
-of it. Figure X depicts a hypothetical analysis of the iris data,
-captured in a single R source file. With informal version control,
-contributors create derivative copies of \texttt{iris.R}, decorating the
-file name with initials, dates, and other descriptors. Even when working
-alone, this leads to multiple versions of \texttt{iris.R} of
-indeterminate relatedness (Fig Xa). In collaborative settings based on
-email distribution, the original file swiftly becomes the root of a
+of it. Figure \ref{fig:diy-vs-git} depicts a hypothetical analysis of
+the iris data, captured in a single R source file. With informal version
+control, contributors create derivative copies of \texttt{iris.R},
+decorating the file name with initials, dates, and other descriptors.
+Even when working alone, this leads to multiple versions of
+\texttt{iris.R} of indeterminate relatedness (Figure
+\ref{fig:diy-vs-git}A). In collaborative settings based on email
+distribution, the original file swiftly becomes the root of a
 complicated phylogeny that no amount of ``Track changes'' and good
-intentions can resolve (Fig Xb).
+intentions can resolve Figure \ref{fig:diy-vs-git}B).
 
-\includegraphics[width=1\linewidth]{diy-vs-git-solo-vs-collab}
+\begin{figure}
+\includegraphics[width=1\linewidth]{diy-vs-git-solo-vs-collab} \caption{\label{fig:diy-vs-git}A: Solo work with DIY version control via filename.
+B: Collaborative work with DIY version control.
+C: Solo work with Git.
+D: Collaborative work with Git.}\label{fig:diy-vs-git}
+\end{figure}
 
 The Git way is to track the evolution of \texttt{iris.R}, through a
-series of commits, each equipped with an explanatory message and a
-nickname. Figure Xc depicts this sort of linear \emph{in situ}
-development. Figure Xd shows the same history for a common collaborative
-Git workflow, where contributors work independently but sync regularly
-to a common version. Especially important versions get a human-readable
-tag, to signal a meaningful milestone. Yes, there is some pain in
-adopting the formalism of Git, but it is worth it.
+series of commits, each equipped with an explanatory message. Figure
+\ref{fig:diy-vs-git}C depicts this linear, \emph{in situ} development
+process. Figure \ref{fig:diy-vs-git}D shows the same history for a
+common collaborative Git workflow, where contributors work independently
+but sync regularly to a common version. Especially important versions
+get a human-readable tag, to signal a meaningful milestone. Yes, there
+is some pain in adopting the formalism of Git, but it is worth it.
 
@@
 to a set of files. GitHub is a hosting service that provides a Git-aware
-home for such projects on the internet. It is like DropBox or Google
+home for such projects on the internet. These relationships are shown in
+Figure \ref{fig:yours-theirs-github}. GitHub is like DropBox or Google
 Drive, but more structured, powerful, and programmatic.
@@
 
-\includegraphics[width=1\linewidth]{your-repo-their-repo-central-remote-repo}
+\begin{figure}
+\includegraphics[width=1\linewidth]{your-repo-their-repo-central-remote-repo} \caption{\label{fig:yours-theirs-github}With Git, all contributors have a copy of the repo, with all files and the full history.
+It is typical to stay in sync through the use of a central remote repo, such as GitHub.
+Hosted remotes like GitHub also provide access to the repo through a web browser.}\label{fig:yours-theirs-github}
+\end{figure}
 
@@
 files in the repo, at a specific moment. Under the hood, that is not
-exactly how Git implements things. Mental models don't have to be
-accurate in order to be useful, but in this case there's some value in
-aligning the two.
+exactly how Git implements things. Although mental models don't have to
+be accurate in order to be useful, in this case it helps to align the
+two.
 
-\includegraphics[width=1\linewidth]{commit-diff-sha-tag}
+\begin{figure}
+\includegraphics[width=1\linewidth]{commit-diff-sha-tag} \caption{\label{fig:commit-diff-sha-tag}Partial commit history for our iris example.}\label{fig:commit-diff-sha-tag}
+\end{figure}
 
-Figure X is another look at our fictional analysis of the iris data,
-focusing on the evolution of its script \texttt{iris.R}. Consider
-version A of this file and a modified version, version B. Assume that
-version A was part of one Git commit and version B was part of the next
-commit. The set of differences between A and B is called a ``diff'' and
-Git users contemplate diffs a lot. Diff inspection is how you re-explain
-to yourself how version A differs from version B. Diff inspection is not
-limited to adjacent commits. You can inspect the diffs between any two
-commits.
+Figure \ref{fig:commit-diff-sha-tag} is another look at our fictional
+analysis of the iris data, focusing on the evolution of its script,
+\texttt{iris.R}. Consider version A of this file and a modified version,
+version B. Assume that version A was part of one Git commit and version
+B was part of the next commit. The set of differences between A and B is
+called a ``diff'' and Git users contemplate diffs a lot. Diff inspection
+is how you re-explain to yourself how version A differs from version B.
+Diff inspection is not limited to adjacent commits. You can inspect the
+diffs between any two commits.
 
@@
 \textbf{history}, by reading commit messages and skimming through diffs,
-is an extremely efficient way to get up to speed. Figure X shows the
-messages associated with the last three commits.
+is an extremely efficient way to get up to speed. Figure
+\ref{fig:commit-diff-sha-tag} shows the messages associated with the
+last three commits.
 
@@
 when you do, usually the first 7 characters suffice. The commit messages
-in Figure X are prefixed by such abbreviated SHAs. You can also
-designate certain snapshots as special with a \textbf{tag}, which is a
-name of your choosing. In a software project, it is typical to tag a
-release with its version, e.g., ``v1.0.3''. For a manuscript or
-analytical project, you might tag the version submitted to a journal or
-transmitted to external collaborators. Figure X shows a tag,
-``draft-01'', associated with the last commit.
+in Figure \ref{fig:commit-diff-sha-tag} are prefixed by such truncated
+SHAs. You can also designate certain snapshots as special with a
+\textbf{tag}, which is a name of your choosing. In a software project,
+it is typical to tag a release with its version, e.g., ``v1.0.3''. For a
+manuscript or analytical project, you might tag the version submitted to
+a journal or transmitted to external collaborators. Figure
+\ref{fig:commit-diff-sha-tag} shows a tag, ``draft-01'', associated with
+the last commit.
 

### 2017-06-29 README.Rmd
Work on first figure

@@
 
-```{r commit-diff-sha-tag, fig.cap="\\label{fig:commit-diff-sha-tag}Partial commit history for our iris example."}
+```{r commit-diff-sha-tag, fig.cap="\\label{fig:commit-diff-sha-tag}Partial commit history for our iris example, highlighting diffs, commit messages, SHAs, and tags."}
 knitr::include_graphics("commit-diff-sha-tag.png")

### 2017-06-29 README.tex
Work on first figure

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{commit-diff-sha-tag} \caption{\label{fig:commit-diff-sha-tag}Partial commit history for our iris example.}\label{fig:commit-diff-sha-tag}
+\includegraphics[width=1\linewidth]{commit-diff-sha-tag} \caption{\label{fig:commit-diff-sha-tag}Partial commit history for our iris example, highlighting diffs, commit messages, SHAs, and tags.}\label{fig:commit-diff-sha-tag}
 \end{figure}

### 2017-06-30 README.Rmd
Add figure of R/Rmd to GitHub-flavored Markdown

@@
 Markdown is especially holy for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code.
+Figure \ref{fig:report-from-r-or-rmd}A shows a simple `.Rmd` document for our iris example.
 Again, do not regard R Markdown as something you must clear your schedule to learn.
@@
 This process is made especially easy in RStudio, but is by no means limited to users of that application.
+Any R user can call `rmarkdown::render("foo.Rmd")`.
 
-*this too is crying out for a visual Rmd -> md -> rendered thing looking good*
+```{r report-from-r-or-rmd, fig.cap = "\\label{fig:report-from-r-or-rmd}A: R Markdown.
+B: R script, equivalent to the R Markdown in A.
+C: The same rendered result is produced by A or B.
+This is how the Markdown will look on GitHub."}
+knitr::include_graphics("report-from-r-or-rmd.png")
+```
 
@@
 Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, __and view the results of running that code__, including figures.
+Figure \ref{fig:report-from-r-or-rmd}C shows how the `.Rmd` document from Figure \ref{fig:report-from-r-or-rmd}A looks when rendered to `.md` and pushed to GitHub.
 It is the best of all worlds, because the code is revealed and, by definition, is the code that produced the results.
@@
 to go from `.R` to `.md`.
-Again, RStudio makes this especially easy, but this is not limited to RStudio.
+Figure \ref{fig:report-from-r-or-rmd}B shows an R script that produces exactly the same rendered report, i.e.
+Figure \ref{fig:report-from-r-or-rmd}C, as the R markdown in Figure \ref{fig:report-from-r-or-rmd}B.
+Again, RStudio makes this especially easy, but is not required.
 Once the Markdown file is pushed to GitHub, it is as if the reader has run your code or is able to look over your shoulder at your R session.
@@
 
-*figure showing the yaml needed for this? so the use of `github_document` or `keep_md = TRUE`; try to make a figure that does double or triple duty, i.e. illustrates this and points made elsewhere*
-
 Before we move on, I want to zoom out and revisit R Markdown and the rmarkdown package more generally.
 Note that R Markdown can be rendered to many more formats than Markdown, including HTML, PDF, and Microsoft Word, and can incorporate code chunks in many languages other than R.
-Markdown is emphasized here, because it is extraordinarily useful in GitHub-hosted projects.
-Even when targetting another output format, the Markdown must be created as a necessary intermediate.
-Sometimes it is all you neeed, such as on GitHub, and in other contexts it may even be discarded.
+Markdown is emphasized here because it is extraordinarily useful in GitHub-hosted projects.
+Many people are used to targetting other output formats and overlook this special synergy between Markdown and GitHub.
 

### 2017-06-30 README.tex
Add figure of R/Rmd to GitHub-flavored Markdown

@@
 \href{http://rmarkdown.rstudio.com}{R Markdown}, which is just Markdown
-that includes chunks of R code. Again, do not regard R Markdown as
-something you must clear your schedule to learn. If you can write email
-and a bit of R code, you can write R Markdown. The
+that includes chunks of R code. Figure \ref{fig:report-from-r-or-rmd}A
+shows a simple \texttt{.Rmd} document for our iris example. Again, do
+not regard R Markdown as something you must clear your schedule to
+learn. If you can write email and a bit of R code, you can write R
+Markdown. The
 \href{https://CRAN.R-project.org/package=rmarkdown}{rmarkdown package}
@@
 especially easy in RStudio, but is by no means limited to users of that
-application.
+application. Any R user can call \texttt{rmarkdown::render("foo.Rmd")}.
 
-\emph{this too is crying out for a visual Rmd -\textgreater{} md
--\textgreater{} rendered thing looking good}
+\begin{figure}
+\includegraphics[width=1\linewidth]{report-from-r-or-rmd} \caption{\label{fig:report-from-r-or-rmd}A: R Markdown.
+B: R script, equivalent to the R Markdown in A.
+C: The same rendered result is produced by A or B.
+This is how the Markdown will look on GitHub.}\label{fig:report-from-r-or-rmd}
+\end{figure}
 
@@
 on GitHub can read the prose, study the R code, \textbf{and view the
-results of running that code}, including figures. It is the best of all
-worlds, because the code is revealed and, by definition, is the code
-that produced the results. And yet a reader can gaze upon the product in
-a web browser, without needing to download the code, install all
-necessary dependencies, and run it.
+results of running that code}, including figures. Figure
+\ref{fig:report-from-r-or-rmd}C shows how the \texttt{.Rmd} document
+from Figure \ref{fig:report-from-r-or-rmd}A looks when rendered to
+\texttt{.md} and pushed to GitHub. It is the best of all worlds, because
+the code is revealed and, by definition, is the code that produced the
+results. And yet a reader can gaze upon the product in a web browser,
+without needing to download the code, install all necessary
+dependencies, and run it.
 
@@
 prepare a rendered version of an R script, i.e.~to go from \texttt{.R}
-to \texttt{.md}. Again, RStudio makes this especially easy, but this is
-not limited to RStudio. Once the Markdown file is pushed to GitHub, it
-is as if the reader has run your code or is able to look over your
+to \texttt{.md}. Figure \ref{fig:report-from-r-or-rmd}B shows an R
+script that produces exactly the same rendered report, i.e.~Figure
+\ref{fig:report-from-r-or-rmd}C, as the R markdown in Figure
+\ref{fig:report-from-r-or-rmd}B. Again, RStudio makes this especially
+easy, but is not required. Once the Markdown file is pushed to GitHub,
+it is as if the reader has run your code or is able to look over your
 shoulder at your R session. This provides a lightweight system for
@@
 
-\emph{figure showing the yaml needed for this? so the use of
-\texttt{github\_document} or \texttt{keep\_md\ =\ TRUE}; try to make a
-figure that does double or triple duty, i.e.~illustrates this and points
-made elsewhere}
-
 Before we move on, I want to zoom out and revisit R Markdown and the
@@
 Word, and can incorporate code chunks in many languages other than R.
-Markdown is emphasized here, because it is extraordinarily useful in
-GitHub-hosted projects. Even when targetting another output format, the
-Markdown must be created as a necessary intermediate. Sometimes it is
-all you neeed, such as on GitHub, and in other contexts it may even be
-discarded.
+Markdown is emphasized here because it is extraordinarily useful in
+GitHub-hosted projects. Many people are used to targetting other output
+formats and overlook this special synergy between Markdown and GitHub.
 

### 2017-06-30 README.Rmd
Add push-pull figure; move random notes elsewhere

@@
 The files in a project arise in different ways and play different roles.
-A critical issue for workflow happiness is to figure out how you want to handle different file types with respect to Git.
+A critical issue for workflow happiness is to decide how to handle different file types with respect to Git.
 You can direct Git to ignore specific files or file types, such as autosaves created by your editor.
@@
 A file that Git does not ignore is said to be __tracked__.
-Here's a useful framework for thinking about what to track:
+Here's a useful framework for deciding what to track:
 
@@
 
-**Intermediates:** These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTex workflows).
+__Intermediates:__ These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTex workflows).
 
@@
 A large binary file that changes often is, therefore, the worst of all worlds.
-This implies that adoption of Git/GitHub rewards a pivot away from `.docx`, `.xlsx`, and `.pdf` as primary file formats and towards `.Rmd`, `.md`, and `.csv`, at least during periods of rapid development.
+This implies that adoption of Git/GitHub suggests a pivot away from `.docx`, `.xlsx`, and `.pdf` as primary file formats and towards `.Rmd`, `.md`, and `.csv`, at least during periods of rapid development.
 
@@
 branches.
-The killer feature is that Git/GitHub enable regular and structured reconciliation of all versions of all the files in the project.
+The killer feature is that Git/GitHub enables regular and structured reconciliation of all versions of all the files in the project.
 It is definitely more complicated than collaborating on a Google Doc, but also more powerful.
@@
 You also push regularly to GitHub, to return the favor, and to maintain its status as the comprehensive, authoritative version of the project.
+Figure \ref{fig:push-pull} depicts this process.
+
+```{r push-pull, fig.cap = "\\label{fig:push-pull}One contributor has made two new commits and updates the master copy on GitHub with a push.
+Another contributor stays up-to-date with a pull from GitHub."}
+knitr::include_graphics("push-pull.png")
+```
 
@@
 So make lots of small commits, sync regularly with GitHub, and only track binary files with good reason.
-
-*inspiration for possible figures on collaboration/merging and the role of a remote, from the barlett talk*
-
-```{r echo = FALSE, out.width="100%"}
-knitr::include_graphics("bartlett-merge-commit.png")
-```
-
-```{r echo = FALSE, out.width="100%"}
-knitr::include_graphics("bartlett-pull.png")
-```
 
@@
 
+*inspiration for possible figure on collaboration/mergin, from the barlett talk*
+
+```{r echo = FALSE, out.width="100%"}
+knitr::include_graphics("bartlett-merge-commit.png")
+```
+
 ## Conclusion
@@
 
-## Random things lying around
-
-Consider links to GitHub repos that exemplify certain points and are very unlikely to disappear any time soon.n
-
-possible github explainer
-https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/
-
-http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst
-
-https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem
-
-https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3
-
-A Quick Introduction to Version Control with Git and GitHub
-http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004668
-
-The active R package development community on GitHub.
-
-  * If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub.
-You can watch the repository to get notified of major activity.
-You can fork it to keep your own copy.
-You can modify your fork to add features or fix bugs and send them back to the owner as a proposed change.
-
-The read-only mirrors of R source and all of CRAN. Coupled with GitHub search features, you can answer alot of your own questions this way.
-
-instructions re: supplementary materials
-http://amstat.tfjournals.com/supplementary-materials/
-
-TAS author instructions
-http://www.tandfonline.com/action/authorSubmission?journalCode=utas20&page=instructions
-
-http://mynameismichelle.com/git-frost/
-
-*These two points were originally written as comparisons with Google Docs, but now there are intervening paragraphs. I think I will cut them.*
-
-*Manage multiple files*.
-A Git repository is inherently multi-file and therefore well suited to projects comprised of many files, evolving in a coordinated fashion.
-Examples include a data analysis, a course website, a blog, an R package, or a book.
-If there is any way to proactively check or enforce their joint functionality, this is something you could verify manually prior to a commit or at certain milestones.
-In the case of a website, you might choose to rebuild the site prior to a commit.
-In the case of an R package, `R CMD check` is one of the easiest things to automate on GitHub.
-
-*Diffs and time travel*.
-Google Docs are fantastic for simple collaborative work, when you don't need detailed access to the history.
-But the version control offered by Google Drive is very limited compared to Git.
-You can't compare versions at arbitrary points in time, temporarily checkout previous versions, or maintain two lines of development.
-
-*May need a segue ... depends on the fate of the two paragraphs above.*
-
-```{r eval = FALSE, echo = FALSE, results = 'asis'}
-cat("*here's one way I see phylogeny in my mind*\n")
-cat("![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)")
-
-cat("\n\n*make such a phylogeny with a set of crazy filenames like this*\n") 
-cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")
-
-cat("\n\n*or riff on the famous PhD Comic*") 
-cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
-```
-
 ```{r eval = is_github_document, echo = FALSE, results = 'asis'}

### 2017-06-30 README.tex
Add push-pull figure; move random notes elsewhere

@@
 The files in a project arise in different ways and play different roles.
-A critical issue for workflow happiness is to figure out how you want to
-handle different file types with respect to Git. You can direct Git to
-ignore specific files or file types, such as autosaves created by your
-editor. This reduces noise and clutter: Git will not pester you to
-commit changes to these files and they will not appear in your GitHub
+A critical issue for workflow happiness is to decide how to handle
+different file types with respect to Git. You can direct Git to ignore
+specific files or file types, such as autosaves created by your editor.
+This reduces noise and clutter: Git will not pester you to commit
+changes to these files and they will not appear in your GitHub
 repository. A file that Git does not ignore is said to be
-\textbf{tracked}. Here's a useful framework for thinking about what to
-track:
+\textbf{tracked}. Here's a useful framework for deciding what to track:
 
@@
 file that changes often is, therefore, the worst of all worlds. This
-implies that adoption of Git/GitHub rewards a pivot away from
+implies that adoption of Git/GitHub suggests a pivot away from
 \texttt{.docx}, \texttt{.xlsx}, and \texttt{.pdf} as primary file
@@
 there can be independent lines of development, i.e.~branches. The killer
-feature is that Git/GitHub enable regular and structured reconciliation
+feature is that Git/GitHub enables regular and structured reconciliation
 of all versions of all the files in the project. It is definitely more
@@
 GitHub, to return the favor, and to maintain its status as the
-comprehensive, authoritative version of the project.
+comprehensive, authoritative version of the project. Figure
+\ref{fig:push-pull} depicts this process.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{push-pull} \caption{\label{fig:push-pull}One contributor has made two new commits and updates the master copy on GitHub with a push.
+Another contributor stays up-to-date with a pull from GitHub.}\label{fig:push-pull}
+\end{figure}
 
@@
 and only track binary files with good reason.
-
-\emph{inspiration for possible figures on collaboration/merging and the
-role of a remote, from the barlett talk}
-
-\includegraphics[width=1\linewidth]{bartlett-merge-commit}
-
-\includegraphics[width=1\linewidth]{bartlett-pull}
 
@@
 
+\emph{inspiration for possible figure on collaboration/mergin, from the
+barlett talk}
+
+\includegraphics[width=1\linewidth]{bartlett-merge-commit}
+
 \subsection{Conclusion}\label{conclusion}
@@
 
-\subsection{Random things lying
-around}\label{random-things-lying-around}
-
-Consider links to GitHub repos that exemplify certain points and are
-very unlikely to disappear any time soon.n
-
-possible github explainer
-\url{https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/}
-
-\url{http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst}
-
-\url{https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem}
-
-\url{https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3}
-
-A Quick Introduction to Version Control with Git and GitHub
-\url{http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004668}
-
-The active R package development community on GitHub.
-
-\begin{itemize}
-\tightlist
-\item
-  If you care deeply about someone else's project, such as an R package
-  you use heavily, you can track its development on GitHub. You can
-  watch the repository to get notified of major activity. You can fork
-  it to keep your own copy. You can modify your fork to add features or
-  fix bugs and send them back to the owner as a proposed change.
-\end{itemize}
-
-The read-only mirrors of R source and all of CRAN. Coupled with GitHub
-search features, you can answer alot of your own questions this way.
-
-instructions re: supplementary materials
-\url{http://amstat.tfjournals.com/supplementary-materials/}
-
-TAS author instructions
-\url{http://www.tandfonline.com/action/authorSubmission?journalCode=utas20\&page=instructions}
-
-\url{http://mynameismichelle.com/git-frost/}
-
-\emph{These two points were originally written as comparisons with
-Google Docs, but now there are intervening paragraphs. I think I will
-cut them.}
-
-\emph{Manage multiple files}. A Git repository is inherently multi-file
-and therefore well suited to projects comprised of many files, evolving
-in a coordinated fashion. Examples include a data analysis, a course
-website, a blog, an R package, or a book. If there is any way to
-proactively check or enforce their joint functionality, this is
-something you could verify manually prior to a commit or at certain
-milestones. In the case of a website, you might choose to rebuild the
-site prior to a commit. In the case of an R package,
-\texttt{R\ CMD\ check} is one of the easiest things to automate on
-GitHub.
-
-\emph{Diffs and time travel}. Google Docs are fantastic for simple
-collaborative work, when you don't need detailed access to the history.
-But the version control offered by Google Drive is very limited compared
-to Git. You can't compare versions at arbitrary points in time,
-temporarily checkout previous versions, or maintain two lines of
-development.
-
-\emph{May need a segue \ldots{} depends on the fate of the two
-paragraphs above.}
-
 \bibliographystyle{agsm}

### 2017-06-30 cutting-room-floor.Rmd
Add push-pull figure; move random notes elsewhere

---
title: "Excuse me, do you have a moment to talk about version control?"
authors:
- name: Jennifer Bryan
  affiliation: |
    RStudio and the Department of Statistics, University of British Columbia
  thanks: |
    The author gratefully acknowledges the constructive feedback from reviewers
    Nicholas Horton, Colin Rundel, and Hadley Wickham.
blinded: 0
output:
  github_document:
    toc: yes
  rticles::asa_article: null
keywords: version control
abstract: Abstract abstract abstract.
bibliography: git-github-for-stats.bib
---

```{r knitr_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
```

## Random notes that used to live at the end of the main document

Consider links to GitHub repos that exemplify certain points and are very unlikely to disappear any time soon.n

possible github explainer
https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/

http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst

https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem

https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3

A Quick Introduction to Version Control with Git and GitHub
http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004668

The active R package development community on GitHub.

  * If you care deeply about someone else's project, such as an R package you use heavily, you can track its development on GitHub. You can watch the repository to get notified of major activity. You can fork it to keep your own copy. You can modify your fork to add features or fix bugs and send them back to the owner as a proposed change.

The read-only mirrors of R source and all of CRAN. Coupled with GitHub search features, you can answer alot of your own questions this way.

instructions re: supplementary materials
http://amstat.tfjournals.com/supplementary-materials/

TAS author instructions
http://www.tandfonline.com/action/authorSubmission?journalCode=utas20&page=instructions

http://mynameismichelle.com/git-frost/

*These two points were originally written as comparisons with Google Docs, but now there are intervening paragraphs. I think I will cut them.*

*Manage multiple files*. A Git repository is inherently multi-file and therefore well suited to projects comprised of many files, evolving in a coordinated fashion. Examples include a data analysis, a course website, a blog, an R package, or a book. If there is any way to proactively check or enforce their joint functionality, this is something you could verify manually prior to a commit or at certain milestones. In the case of a website, you might choose to rebuild the site prior to a commit. In the case of an R package, `R CMD check` is one of the easiest things to automate on GitHub.

*Diffs and time travel*. Google Docs are fantastic for simple collaborative work, when you don't need detailed access to the history. But the version control offered by Google Drive is very limited compared to Git. You can't compare versions at arbitrary points in time, temporarily checkout previous versions, or maintain two lines of development.

*May need a segue ... depends on the fate of the two paragraphs above.*

```{r eval = FALSE, echo = FALSE, results = 'asis'}
cat("*here's one way I see phylogeny in my mind*\n")
cat("![](https://upload.wikimedia.org/wikipedia/commons/2/2b/Phylogeny_of_Retroviruses.jpg)")

cat("\n\n*make such a phylogeny with a set of crazy filenames like this*\n") 
cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")

cat("\n\n*or riff on the famous PhD Comic*") 
cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
```

### 2017-06-30 README.Rmd
Work on branches and pull requests

@@
 Branches and pull requests are an extremely powerful feature of Git/GitHub and should be your first foray beyond the basics of commit, push, and pull.
-A branch is an independent line of development within a repo, with the notion it will eventually be merged back into the main branch, a.k.a.
-"master".
+A branch is an independent line of development within a repo, where the intent is usually to merge it into the master branch when ready.
 In a course website, you might work in a branch to update a series of lessons, while leaving the current version intact in the meantime.
-A pull request is a specific and formal GitHub process for merging one branch into another.
-You can make a pull request between two branches in the same repo or between two copies of a repo.
+A pull request is a GitHub-specific way to propose changes to a repo that augments Git's regular branch and merge workflow with facilities for structured review.
+A pull request can be made between two branches in the same repo, such as `master` and `bugfix`, or between two independent copies of a repo, an original and a so-called "fork".
 This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages *link to a search?*.

### 2017-06-30 README.tex
Work on branches and pull requests

@@
 push, and pull. A branch is an independent line of development within a
-repo, with the notion it will eventually be merged back into the main
-branch, a.k.a. ``master''. In a course website, you might work in a
-branch to update a series of lessons, while leaving the current version
-intact in the meantime. A pull request is a specific and formal GitHub
-process for merging one branch into another. You can make a pull request
-between two branches in the same repo or between two copies of a repo.
-This is the mechanism for making and accepting contributions to open
-source software projects on GitHub, including many popular R packages
-\emph{link to a search?}.
+repo, where the intent is usually to merge it into the master branch
+when ready. In a course website, you might work in a branch to update a
+series of lessons, while leaving the current version intact in the
+meantime. A pull request is a GitHub-specific way to propose changes to
+a repo that augments Git's regular branch and merge workflow with
+facilities for structured review. A pull request can be made between two
+branches in the same repo, such as \texttt{master} and \texttt{bugfix},
+or between two independent copies of a repo, an original and a so-called
+``fork''. This is the mechanism for making and accepting contributions
+to open source software projects on GitHub, including many popular R
+packages \emph{link to a search?}.
 

### 2017-06-30 README.Rmd
Add link suggested by @nicholasjhorton; fixes #13

@@
 output:
-  rticles::asa_article: null
   github_document:
     toc: yes
+  rticles::asa_article: null
 keywords: version control
@@
 All the major hosting providers offer private repositories with flexible control over who can read or write to the repo.
-There are many ways to get private repositories for low or no cost, especially for academics.
+There are many ways to get private repositories for low or no cost, [especially for academics](https://help.github.com/articles/discounted-organization-accounts/).
 If you outgrow this initial arrangement, you can throw some combination of technical savvy and money at the problem.

### 2017-07-06 README.Rmd
Move figure fodder to img/
Cite R

@@
   * The marginal effort required to create a web presence for a project is negligible.
-  * GitHub makes a fantastic course management system for courses that use R. You can exchange actual working code with your students and explore the associated results.
+  * GitHub makes a fantastic course management system for courses that use R [@R]. You can exchange actual working code with your students and explore the associated results.
   * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.
@@
 ```{r echo = FALSE, out.width="100%"}
-knitr::include_graphics("bartlett-merge-commit.png")
+knitr::include_graphics("img/bartlett-merge-commit.png")
 ```

### 2017-07-06 README.tex
Add link suggested by @nicholasjhorton; fixes #13
Re-render PDF
Cite R

@@
   GitHub makes a fantastic course management system for courses that use
-  R. You can exchange actual working code with your students and explore
-  the associated results.
+  R \citep{R}. You can exchange actual working code with your students
+  and explore the associated results.
 \item
@@
 flexible control over who can read or write to the repo. There are many
-ways to get private repositories for low or no cost, especially for
-academics. If you outgrow this initial arrangement, you can throw some
-combination of technical savvy and money at the problem. You can either
-pay for a higher level of service, self-host one of these platforms, or
-advocate for organization-wide solutions.
+ways to get private repositories for low or no cost,
+\href{https://help.github.com/articles/discounted-organization-accounts/}{especially
+for academics}. If you outgrow this initial arrangement, you can throw
+some combination of technical savvy and money at the problem. You can
+either pay for a higher level of service, self-host one of these
+platforms, or advocate for organization-wide solutions.
 
@@
 
-\includegraphics[width=1\linewidth]{bartlett-merge-commit}
+\includegraphics[width=1\linewidth]{img/bartlett-merge-commit}
 

### 2017-07-06 README.Rmd
Typos
Clarify flow of time and development in first figure

@@
 each equipped with an explanatory message.
-Figure \ref{fig:diy-vs-git}C depicts this linear, *in situ* development process.
+Figure \ref{fig:diy-vs-git}C depicts this linear, *in situ* development process, with development and time flowing from bottom to top.
 Figure \ref{fig:diy-vs-git}D shows the same history for a common collaborative Git workflow, where contributors work independently but sync regularly to a common version.
-Especially important
-versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
-is some pain in adopting the formalism of Git, but it is worth it.
+Especially important versions get a human-readable tag, e.g., "draft-01", to signal a meaningful milestone.
+Yes, there is some pain in adopting the formalism of Git, but it is worth it.
 
@@
 
-We've explored Git's powerful structure for file management, so where does GitHub fit in?
+We've introduced Git's powerful structure for file management, so where does GitHub fit in?
 [GitHub](https://github.com) complements Git by providing a slick user interface and distribution mechanism for Git repositories.
@@
 
-The remote host acts as the clearinghouse for a Git-managed project. This allows
+The remote host acts as the distributor for a Git-managed project. This allows
 others to browse project files, explore their history, sync up with the current version, and perhaps even propose or make changes.
@@
 If I've convinced you to experiment with Git and GitHub, you need to do some initial setup.
-These first steps happen one or, for some steps, once per computer.
+These first steps happen once or, for some steps, once per computer.
 This is excerpted from [Happy Git and GitHub for the useR](http://happygitwithr.com), which holds battle-tested instructions honed over several years in [STAT 545](http://stat545.com) at the University of British Columbia.

### 2017-07-06 README.tex
Typos
Clarify flow of time and development in first figure

@@
 \ref{fig:diy-vs-git}C depicts this linear, \emph{in situ} development
-process. Figure \ref{fig:diy-vs-git}D shows the same history for a
-common collaborative Git workflow, where contributors work independently
-but sync regularly to a common version. Especially important versions
-get a human-readable tag, to signal a meaningful milestone. Yes, there
-is some pain in adopting the formalism of Git, but it is worth it.
+process, with development and time flowing from bottom to top. Figure
+\ref{fig:diy-vs-git}D shows the same history for a common collaborative
+Git workflow, where contributors work independently but sync regularly
+to a common version. Especially important versions get a human-readable
+tag, e.g., ``draft-01'', to signal a meaningful milestone. Yes, there is
+some pain in adopting the formalism of Git, but it is worth it.
 
@@
 
-We've explored Git's powerful structure for file management, so where
+We've introduced Git's powerful structure for file management, so where
 does GitHub fit in? \href{https://github.com}{GitHub} complements Git by
@@
 
-The remote host acts as the clearinghouse for a Git-managed project.
-This allows others to browse project files, explore their history, sync
-up with the current version, and perhaps even propose or make changes.
+The remote host acts as the distributor for a Git-managed project. This
+allows others to browse project files, explore their history, sync up
+with the current version, and perhaps even propose or make changes.
 GitHub's well-designed web interface is a dramatic improvement over
@@
 If I've convinced you to experiment with Git and GitHub, you need to do
-some initial setup. These first steps happen one or, for some steps,
+some initial setup. These first steps happen once or, for some steps,
 once per computer. This is excerpted from

### 2017-07-06 README.Rmd
Rationalize citation / linking of RStudio, Git

@@
 
-Now the bad news: Git was built neither for the exact usage described here, nor for broad usability.
-You will undoubtedly notice this, so it's best to know in advance.
-Happily, there are many helpful tools that mediate your interactions with Git.
-GitHub itself is a fine example, as is [RStudio](https://www.rstudio.com/products/rstudio/).
-In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.
+Now the bad news: Git was built neither for the exact usage described here, nor
+for broad usability. You will undoubtedly notice this, so it's best to know in
+advance. Happily, there are many helpful tools that mediate your interactions
+with Git. GitHub itself is a fine example, as is
+[RStudio](https://www.rstudio.com/products/rstudio/) [@rstudio]. In addition to
+pointing out tools that soften Git's sharpest edges, I recommend specific habits
+and attitudes that reduce frustration.
 
@@
 
-[Git](http://git-scm.com) is a __version control system__. Its original purpose 
-was to help groups of developers work collaboratively on big software projects. 
-Git manages the evolution of a set of files -- called a __repository__ or 
-__repo__ -- in a sane, highly structured way. It is like the "Track 
-Changes" feature from Microsoft Word, but more rigorous, powerful, and scaled up
-to multiple files.
+Git is a __version control system__. Its original purpose was to help groups of
+developers work collaboratively on big software projects. Git manages the
+evolution of a set of files -- called a __repository__ or __repo__ -- in a sane,
+highly structured way. It is like the "Track Changes" feature from Microsoft
+Word, but more rigorous, powerful, and scaled up to multiple files.
 
@@
 Even when working alone, this leads to multiple versions of `iris.R` of indeterminate relatedness (Figure \ref{fig:diy-vs-git}A).
-In collaborative settings based on email distribution, the original file swiftly becomes the root of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve Figure \ref{fig:diy-vs-git}B).
+In collaborative settings based on email distribution, the original file swiftly becomes part of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve Figure \ref{fig:diy-vs-git}B).
 
@@
 We've introduced Git's powerful structure for file management, so where does GitHub fit in?
-[GitHub](https://github.com) complements Git by providing a slick user interface and distribution mechanism for Git repositories.
+GitHub complements Git by providing a slick user interface and distribution mechanism for Git repositories.
 Git is the software you will use locally to record changes to a set of files.
@@
 * Install a local Git client, *optional but highly recommended*. A Git client 
-provides a graphical user interface for Git, which is otherwise
-command-line only. If you are an R user, you will find that 
-[RStudio](https://www.rstudio.com/products/rstudio-desktop/) provides a great 
-deal of this functionality. There are some notable gaps, however, so you might 
-still choose to install a dedicated and comprehensive Git client such as 
+provides a graphical user interface for Git, which is otherwise command-line
+only. If you are an R user, you will find that RStudio provides a great deal of
+this functionality. There are some notable gaps, however, so you might still
+choose to install a dedicated and comprehensive Git client such as 
 [SourceTree](https://www.sourcetreeapp.com) or 
@@
 
-I targeted [GitHub](https://github.com) -- not [Bitbucket](https://bitbucket.org) or [GitLab](https://about.gitlab.com) -- for the sake of specificity.
-However, all the big-picture principles and even some mechanics will carry over to these alternative hosting platforms.
-I am advocating for the use of hosted version control as a general concept, with GitHub being the best and most common provider today.
-I note that many companies and even universities are starting to make GitHub Enterprise or GitLab available internally.
-For example, we host our own instance of GitHub Enterprise at UBC to support our Master of Data Science program
+I targeted GitHub -- not [Bitbucket](https://bitbucket.org) or
+[GitLab](https://about.gitlab.com) -- for the sake of specificity. However, all
+the big-picture principles and even some mechanics will carry over to these
+alternative hosting platforms. I am advocating for the use of hosted version
+control as a general concept, with GitHub being the best and most common
+provider today. I note that many companies and even universities are starting to
+make GitHub Enterprise or GitLab available internally. For example, we host our
+own instance of GitHub Enterprise at UBC to support our Master of Data Science
+program
 

### 2017-07-06 README.tex
Rationalize citation / linking of RStudio, Git

@@
 mediate your interactions with Git. GitHub itself is a fine example, as
-is \href{https://www.rstudio.com/products/rstudio/}{RStudio}. In
-addition to pointing out tools that soften Git's sharpest edges, I
-recommend specific habits and attitudes that reduce frustration.
+is \href{https://www.rstudio.com/products/rstudio/}{RStudio}
+\citep{rstudio}. In addition to pointing out tools that soften Git's
+sharpest edges, I recommend specific habits and attitudes that reduce
+frustration.
 
@@
 
-\href{http://git-scm.com}{Git} is a \textbf{version control system}. Its
-original purpose was to help groups of developers work collaboratively
-on big software projects. Git manages the evolution of a set of files --
-called a \textbf{repository} or \textbf{repo} -- in a sane, highly
-structured way. It is like the ``Track Changes'' feature from Microsoft
-Word, but more rigorous, powerful, and scaled up to multiple files.
+Git is a \textbf{version control system}. Its original purpose was to
+help groups of developers work collaboratively on big software projects.
+Git manages the evolution of a set of files -- called a
+\textbf{repository} or \textbf{repo} -- in a sane, highly structured
+way. It is like the ``Track Changes'' feature from Microsoft Word, but
+more rigorous, powerful, and scaled up to multiple files.
 
@@
 \ref{fig:diy-vs-git}A). In collaborative settings based on email
-distribution, the original file swiftly becomes the root of a
-complicated phylogeny that no amount of ``Track changes'' and good
-intentions can resolve Figure \ref{fig:diy-vs-git}B).
+distribution, the original file swiftly becomes part of a complicated
+phylogeny that no amount of ``Track changes'' and good intentions can
+resolve Figure \ref{fig:diy-vs-git}B).
 
@@
 We've introduced Git's powerful structure for file management, so where
-does GitHub fit in? \href{https://github.com}{GitHub} complements Git by
-providing a slick user interface and distribution mechanism for Git
-repositories. Git is the software you will use locally to record changes
-to a set of files. GitHub is a hosting service that provides a Git-aware
-home for such projects on the internet. These relationships are shown in
-Figure \ref{fig:yours-theirs-github}. GitHub is like DropBox or Google
-Drive, but more structured, powerful, and programmatic.
+does GitHub fit in? GitHub complements Git by providing a slick user
+interface and distribution mechanism for Git repositories. Git is the
+software you will use locally to record changes to a set of files.
+GitHub is a hosting service that provides a Git-aware home for such
+projects on the internet. These relationships are shown in Figure
+\ref{fig:yours-theirs-github}. GitHub is like DropBox or Google Drive,
+but more structured, powerful, and programmatic.
 
@@
   otherwise command-line only. If you are an R user, you will find that
-  \href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio}
-  provides a great deal of this functionality. There are some notable
-  gaps, however, so you might still choose to install a dedicated and
-  comprehensive Git client such as
+  RStudio provides a great deal of this functionality. There are some
+  notable gaps, however, so you might still choose to install a
+  dedicated and comprehensive Git client such as
   \href{https://www.sourcetreeapp.com}{SourceTree} or
@@
 
-I targeted \href{https://github.com}{GitHub} -- not
-\href{https://bitbucket.org}{Bitbucket} or
+I targeted GitHub -- not \href{https://bitbucket.org}{Bitbucket} or
 \href{https://about.gitlab.com}{GitLab} -- for the sake of specificity.

### 2017-07-07 README.Rmd
Abstract, conclusion, Header 1 not 2, spellcheck

@@
   rticles::asa_article: null
-keywords: version control
-abstract: Abstract abstract abstract.
+keywords: |
+    data science, workflow, Git, GitHub, R language, R Markdown, reproducibility
+abstract: |
+    Data analysis, statistical research, and teaching statistics have at least
+    one thing in common: these activities all produce many files! There may be
+    data files, source code, figures, tables, prepared reports, and much more.
+    These files generally evolve over the course of a project and often need to
+    be shared with others, for reading or edits, as a project unfolds. Without
+    explicit and structured management, project organization can easily descend
+    into chaos, taking time away from the primary work and reducing the quality
+    of the final product. This unhappy result can be avoided by repurposing
+    tools and workflows from the software development world, namely, distributed
+    version control. This article describes the use of the version control
+    system Git and and the hosting site GitHub for statistical and data
+    scientific workflows. Special attention is given to projects that use the
+    statistical language R and, optionally, R Markdown documents.
 bibliography: git-github-for-stats.bib
@@
 knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
-```
-
-```{r include = FALSE}
-## I temporarily embed some figures from the internet and LaTeX just. can't.
-## deal. with. it.
 is_github_document <-
@@
 
-## Why Git?
+# Why Git?
 
@@
 
-## What is Git?
+# What is Git?
 
@@
 
-## Who should read this and what to expect
+# Who should read this and what to expect
 
@@
 
-## What is GitHub?
+# What is GitHub?
 
@@
 
-## Initial system setup
+# Initial system setup
 
@@
   * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
-  * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.
+  * Realize that no one is giving out Git style points. It's OK to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.
 
-## Repositories and workflow
+# Repositories and workflow
 
@@
 
-## Commits, diffs, and tags
+# Commits, diffs, and tags
 
@@
 
-## Markdown is special on GitHub
+# Markdown is special on GitHub
 
@@
 
-## Markdown is special for R users
+# Markdown is special for R users
 
@@
 Markdown is emphasized here because it is extraordinarily useful in GitHub-hosted projects.
-Many people are used to targetting other output formats and overlook this special synergy between Markdown and GitHub.
+Many people are used to targeting other output formats and overlook this special synergy between Markdown and GitHub.
 
-## Which files to commit
+# Which files to commit
 
@@
 
-__Intermediates:__ These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTex workflows).
+__Intermediates:__ These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTeX workflows).
 
@@
 
-## Collaboration
+# Collaboration
 
@@
 
-## GitHub as web presence
+# GitHub as web presence
 
@@
 
-## More resources
+# Where to go next?
 
-I've tried to convey the main points about the use of Git and GitHub in statistical and data analytical settings, but I've had to leave many things out.
-There are more advanced topics that will come up once your use of Git becomes more sophisticated and there are topics that are only relevant to certain types of reader.
+I've tried to convey the main points about the use of Git and GitHub in statistical and data analytical settings, but I've left many things out.
+There are more advanced topics that will come up as your use of Git becomes more sophisticated and there are topics that are only relevant to certain types of reader.
 
@@
 A branch is an independent line of development within a repo, where the intent is usually to merge it into the master branch when ready.
+In fact, Figure \ref{fig:diy-vs-git}D depicts exactly such a branch-and-merge workflow, which can be used in both collaborative and solo projects.
 In a course website, you might work in a branch to update a series of lessons, while leaving the current version intact in the meantime.
-A pull request is a GitHub-specific way to propose changes to a repo that augments Git's regular branch and merge workflow with facilities for structured review.
+A pull request is a GitHub-specific way to propose changes to a repo that overlays Git's regular branch and merge workflow with facilities for structured review.
 A pull request can be made between two branches in the same repo, such as `master` and `bugfix`, or between two independent copies of a repo, an original and a so-called "fork".
-This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages *link to a search?*.
+This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages.
 
-*inspiration for possible figure on collaboration/mergin, from the barlett talk*
+# Call to action
 
-```{r echo = FALSE, out.width="100%"}
-knitr::include_graphics("img/bartlett-merge-commit.png")
-```
-
-## Conclusion
-
-*NEEDS ONE ... BUT SHORT ... SOME FODDER*
-
-Transparency about process and product is increasingly important in science.
-The SOMETHING for reproducibility is well accepted.
-A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.
-
-make statistical thought and implementation available
+Statistics is an important part of Data Science, though there is lively debate about the exact relationship, both past and future [@donoho].
+I strongly disagree with glib claims that "Data Science is just Statistics", which I've heard in academia.
+For reasons both genuine and pragmatic, Statistics departments want to participate in the growth of Data Science at their respective universities.
+So far, their success has been decidedly mixed [@donoho].
+In the era of Data Science, staying current can't be limited to mathematical and methodological developments.
+The holistic work style described here is widely regarded as current best practice, yet we generally neither practice nor preach it in academic statistics.
+It is but one example of the many ways in which Data Science is **not** just statistics.
+By acknowledging the importance of modern workflows and tooling, Statistics can solidify its claim to being one of the core disciplines of Data Science.
 
@@
 ## surely there is a better way?
-cat("## References\n")
+cat("# References\n")
 ```

### 2017-07-07 README.tex
Abstract, conclusion, Header 1 not 2, spellcheck

@@
 \begin{abstract}
-Abstract abstract abstract.
+Data analysis, statistical research, and teaching statistics have at
+least one thing in common: these activities all produce many files!
+There may be data files, source code, figures, tables, prepared reports,
+and much more. These files generally evolve over the course of a project
+and often need to be shared with others, for reading or edits, as a
+project unfolds. Without explicit and structured management, project
+organization can easily descend into chaos, taking time away from the
+primary work and reducing the quality of the final product. This unhappy
+result can be avoided by repurposing tools and workflows from the
+software development world, namely, distributed version control. This
+article describes the use of the version control system Git and and the
+hosting site GitHub for statistical and data scientific workflows.
+Special attention is given to projects that use the statistical language
+R and, optionally, R Markdown documents.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} version control
+{\it Keywords:} data science, workflow, Git, GitHub, R language, R Markdown,
+reproducibility
 \vfill
@@
 
-\subsection{Why Git?}\label{why-git}
+\section{Why Git?}\label{why-git}
 
@@
 
-\subsection{What is Git?}\label{what-is-git}
+\section{What is Git?}\label{what-is-git}
 
@@
 
-\subsection{Who should read this and what to
+\section{Who should read this and what to
 expect}\label{who-should-read-this-and-what-to-expect}
@@
 
-\subsection{What is GitHub?}\label{what-is-github}
+\section{What is GitHub?}\label{what-is-github}
 
@@
 
-\subsection{Initial system setup}\label{initial-system-setup}
+\section{Initial system setup}\label{initial-system-setup}
 
@@
 \item
-  Realize that no one is giving out Git style points. It's ok to
+  Realize that no one is giving out Git style points. It's OK to
   ``power-cycle'', i.e.~re-initialize the Git repository, to get
@@
 
-\subsection{Repositories and workflow}\label{repositories-and-workflow}
+\section{Repositories and workflow}\label{repositories-and-workflow}
 
@@
 
-\subsection{Commits, diffs, and tags}\label{commits-diffs-and-tags}
+\section{Commits, diffs, and tags}\label{commits-diffs-and-tags}
 
@@
 
-\subsection{Markdown is special on
+\section{Markdown is special on
 GitHub}\label{markdown-is-special-on-github}
@@
 
-\subsection{Markdown is special for R
+\section{Markdown is special for R
 users}\label{markdown-is-special-for-r-users}
@@
 Markdown is emphasized here because it is extraordinarily useful in
-GitHub-hosted projects. Many people are used to targetting other output
+GitHub-hosted projects. Many people are used to targeting other output
 formats and overlook this special synergy between Markdown and GitHub.
 
-\subsection{Which files to commit}\label{which-files-to-commit}
+\section{Which files to commit}\label{which-files-to-commit}
 
@@
 serve a temporary purpose, but are not intrinsically valuable (e.g.,
-\texttt{.aux} and \texttt{.log} in LaTex workflows).
+\texttt{.aux} and \texttt{.log} in LaTeX workflows).
 
@@
 
-\subsection{Collaboration}\label{collaboration}
+\section{Collaboration}\label{collaboration}
 
@@
 
-\subsection{GitHub as web presence}\label{github-as-web-presence}
+\section{GitHub as web presence}\label{github-as-web-presence}
 
@@
 
-\subsection{More resources}\label{more-resources}
+\section{Where to go next?}\label{where-to-go-next}
 
 I've tried to convey the main points about the use of Git and GitHub in
-statistical and data analytical settings, but I've had to leave many
-things out. There are more advanced topics that will come up once your
-use of Git becomes more sophisticated and there are topics that are only
-relevant to certain types of reader.
+statistical and data analytical settings, but I've left many things out.
+There are more advanced topics that will come up as your use of Git
+becomes more sophisticated and there are topics that are only relevant
+to certain types of reader.
 
@@
 repo, where the intent is usually to merge it into the master branch
-when ready. In a course website, you might work in a branch to update a
-series of lessons, while leaving the current version intact in the
+when ready. In fact, Figure \ref{fig:diy-vs-git}D depicts exactly such a
+branch-and-merge workflow, which can be used in both collaborative and
+solo projects. In a course website, you might work in a branch to update
+a series of lessons, while leaving the current version intact in the
 meantime. A pull request is a GitHub-specific way to propose changes to
-a repo that augments Git's regular branch and merge workflow with
+a repo that overlays Git's regular branch and merge workflow with
 facilities for structured review. A pull request can be made between two
@@
 to open source software projects on GitHub, including many popular R
-packages \emph{link to a search?}.
+packages.
 
-\emph{inspiration for possible figure on collaboration/mergin, from the
-barlett talk}
+\section{Call to action}\label{call-to-action}
 
-\includegraphics[width=1\linewidth]{img/bartlett-merge-commit}
-
-\subsection{Conclusion}\label{conclusion}
-
-\emph{NEEDS ONE \ldots{} BUT SHORT \ldots{} SOME FODDER}
-
-Transparency about process and product is increasingly important in
-science. The SOMETHING for reproducibility is well accepted. A more
-underappreciated benefit is democratization of our field, as this
-affords a much broader audience a clear view of how scientists and
-programmers work.
-
-make statistical thought and implementation available
+Statistics is an important part of Data Science, though there is lively
+debate about the exact relationship, both past and future
+\citep{donoho}. I strongly disagree with glib claims that ``Data Science
+is just Statistics'', which I've heard in academia. For reasons both
+genuine and pragmatic, Statistics departments want to participate in the
+growth of Data Science at their respective universities. So far, their
+success has been decidedly mixed \citep{donoho}. In the era of Data
+Science, staying current can't be limited to mathematical and
+methodological developments. The holistic work style described here is
+widely regarded as current best practice, yet we generally neither
+practice nor preach it in academic statistics. It is but one example of
+the many ways in which Data Science is \textbf{not} just statistics. By
+acknowledging the importance of modern workflows and tooling, Statistics
+can solidify its claim to being one of the core disciplines of Data
+Science.
 

### 2017-07-07 README.Rmd
https

@@
 
-Why would a statistician use a version control system, such as [Git](http://git-scm.com) [@git]?
-And what is the point of hosting your work online, e.g., on [GitHub](https://github.com) [@github]?
+Why would a statistician use a version control system, such as [Git](https://git-scm.com) [@git]?
+And what is the point of hosting your work online, e.g., on [GitHub](https://github.com)[@github]?
 Could the gains possibly justify the inevitable pain?

### 2017-07-07 appendix.Rmd
First crack at appendix

---
title: "Appendix: Tutorials and examples"
output:
  github_document:
    toc: yes
  pdf_document: null
---

Supplement for the article:

Excuse me, do you have a moment to talk about version control?  
by Jennifer Bryan

This appendix provides annotated links to step-by-step instructions, examples, and other resources, as promised in the main article. It is organized using the same
section headers.

Many of these resources live primarily on the web and, therefore, are more perishable than the content main text. As time goes on, it is possible that URLs may change, and it might be necessary to use a site's navigation or search features to re-locate the resource.

# Why Git?

Software:

  * [Git](https://git-scm.com) is a version control system.
  * [RStudio](https://www.rstudio.com/products/rstudio/) is an integrated development environment for R.

Website:

  *  [GitHub](https://github.com) is a web hosting service for Git repositories.

# What is Git?

Comic relief:

  * ["FINAL.doc"](http://phdcomics.com/comics/archive.php?comicid=1531) from [PhD](http://phdcomics.com), pokes gentle fun at DIY filename-based version control.

# Who should read this and what to expect

*no links*

# What is GitHub?

Example of utility: Drill down into the code of any R package on CRAN by clicking around this read-only mirror of CRAN stewarded by Gábor Csárdi.

  * <https://github.com/cran> Just click on a package or use the search box to find the packages you use most.
  * Compare this to the alternative of downloading the `*.tar.gz` from CRAN, unpacking it, and exploring the code locally. And repeating that process whenever the package changes.

Example of utility: See the files behind the book ["R for Data Science"](http://r4ds.had.co.nz) by Garrett Grolemund and Hadley Wickham:

  * <https://github.com/hadley/r4ds>
  * Click on files to see their current state right now.
  * Look at the [commit history](https://github.com/hadley/r4ds/commits/master) to see how things have evolved.
  * Use the search box in upper left corner to search for specific words, like `join` or `factor`.
  * See a [summary of what the >100 contributors](https://github.com/hadley/r4ds/graphs/contributors) have added to this book: this is how readers submit fixes and small improvements.
  * Look at the [issues](https://github.com/hadley/r4ds/issues), such as [the one where I discuss error handling](https://github.com/hadley/r4ds/issues/323) in R Markdown documents, which is linked to the [exact commit](https://github.com/hadley/r4ds/commit/64ce2fa3579a96cdfd97adb69554d8c0de73177c) where the authors addressed my suggestion.

Comic relief

  * [xkcd 1597](https://xkcd.com/1597/) is about Git and strategies for coping with its famously unfriendly user interface.

# Initial system setup

[Happy Git and GitHub for the useR](http://happygitwithr.com) holds concrete setup and early usage information, developed through several years of Git/GitHub use in [STAT 545](http://stat545.com) at the University of British Columbia.

  * [Register a free GitHub account](http://happygitwithr.com/github-acct.html)
  * [Install or update R and RStudio](http://happygitwithr.com/install-r-rstudio.html)
  * [Install Git](http://happygitwithr.com/install-git.html) **We have specific reasons for our installation methods, so seriously consider using them!**
  * [Introduce yourself to Git](http://happygitwithr.com/hello-git.html)
  * [Install a Git client](http://happygitwithr.com/git-client.html). Good options -- you might want more than one!
    - [RStudio](https://www.rstudio.com/products/rstudio/) has basic Git client functionality.
    - [SourceTree](https://www.sourcetreeapp.com) is free and excellent.
    - [GitKraken](https://www.gitkraken.com) is also well-regarded and works across Mac OS, Windows, and Linux.
  * [Connect to GitHub](http://happygitwithr.com/push-pull-github.html) will affirm that your system is set up properly.

# Repositories and workflow

Continuing to link to [Happy Git and GitHub for the useR](http://happygitwithr.com).

  * [New project, GitHub first](http://happygitwithr.com/new-github-first.html) is a good introductory workflow for starting new projects.
  * [Existing project, GitHub first](http://happygitwithr.com/existing-github-first.html) shows one way to take an existing project and make it a Git/GitHub repo and RStudio Project.
  * [Existing project, GitHub last](http://happygitwithr.com/existing-github-last.html) shows another way to get an existing project onto GitHub.

# Commits, diffs, and tags

Comic relief

  * [xkcd 1296](https://xkcd.com/1296/) is about the challenge of writing highly informative commit messages all the time.

Repo, commit, and diff. [dplyr](http://dplyr.tidyverse.org) is a widely used R package whose development is managed with Git and on GitHub.

  * Visit its GitHub repo: <https://github.com/tidyverse/dplyr>.
  * Look at some of the recent commits: <https://github.com/tidyverse/dplyr/commits/master>.
  * Click on an individual commit to see exactly what changed, possibly across multiple files, from one snapshot to the next.
  * CRAN releases are all marked with tags: <https://github.com/tidyverse/dplyr/releases>, which make it easy to see the exact state of the source code for any specific version.

# Markdown is special on GitHub

[Markdown](https://daringfireball.net/projects/markdown/syntax) is widely used to author content for the web:

  * WordPress: <https://en.support.wordpress.com/markdown/>
  * Stackoverflow: <https://stackoverflow.com/editing-help>
  * GitHub: <https://help.github.com/categories/writing-on-github/>

# Markdown is special for R users

R Markdown = R + Markdown:

  * <http://rmarkdown.rstudio.com> is a landing page for a huge array of resources about R Markdown.
  * [Happy Git and GitHub for the useR](http://happygitwithr.com) has tutorials for:
    - [Test drive R Markdown](http://happygitwithr.com/rmd-test-drive.html)
    - [Render an R script](http://happygitwithr.com/r-test-drive.html)

# Which files to commit

[Happy Git and GitHub for the useR](http://happygitwithr.com) deals with some of these choices in [Make a GitHub repo browsable](http://happygitwithr.com/repo-browsability.html).

# Collaboration

[How Google Solved the Version Control Problem](https://www.onshape.com/cad-blog/how-google-solved-the-version-control-problem) is an excellent presentation of the fundamental problem with DIY version control and how cloud-based systems, like Google Drive and GitHub, offer a better way.

GitHub has many excellent resources for collaborative workflows, using branches, issues, and pull requests, including how to tackle merge conflicts:

  * [Collaborating with issues and pull requests](https://help.github.com/categories/collaborating-with-issues-and-pull-requests/)

# GitHub as web presence

[Happy Git and GitHub for the useR](http://happygitwithr.com) covers this in [Make a GitHub repo browsable](http://happygitwithr.com/repo-browsability.html).

If you decide to make a proper website (vs just making your GitHub repo more browsable), go here:

  * [GitHub Pages](https://help.github.com/categories/github-pages-basics/) is the basic platform.
  * [GitHub Pages can create a simple website](https://github.com/blog/2289-publishing-with-github-pages-now-as-easy-as-1-2-3) automatically.
  * [Jekyll static site generator](https://jekyllrb.com) for maximum flexibility.
  * R- and R Markdown-specific frameworks:
    - [R Markdown websites](http://rmarkdown.rstudio.com/rmarkdown_websites.html)
    - [bookdown](https://bookdown.org)
    - [blogdown](https://bookdown.org/yihui/blogdown/)

# Where to go next?

Alternatives to GitHub:

  * [Bitbucket](https://bitbucket.org)
  * [GitLab](https://about.gitlab.com)

# Call to action

Donoho's "50 Years of Data Science" examines Statistics' slowness to embrace a broader definition of scholarly statistical activity and training:

  * <http://courses.csail.mit.edu/18.337/2015/docs/50YearsDataScience.pdf>

### 2017-07-07 README.Rmd
Edits and re-render

--- 2017-07-07 README.Rmd
@@
 
-Why would a statistician use a version control system, such as [Git](http://git-scm.com) [@git]?
+Why would a statistician use a version control system, such as [Git](https://git-scm.com) [@git]?
 And what is the point of hosting your work online, e.g., on [GitHub](https://github.com) [@github]?
@@
 This has an implication for selecting your first Git projects: you will enjoy the most gain for your pain if you pick a project that involves sharing rapidly evolving files with others.
-It is tempting to pick a quiet, private project, but you risk missing out on the main benefits of formal version control.
+It is tempting to pick a quiet, private project.
+But if you do, you may never find the benefits of formal version control compelling enough to cement the new habit.
 
@@
 Statistics is an important part of Data Science, though there is lively debate about the exact relationship, both past and future [@donoho].
-I strongly disagree with glib claims that "Data Science is just Statistics", which I've heard in academia.
+I strongly disagree with glib claims that "Data Science is just Statistics", which can be heard in academia.
 For reasons both genuine and pragmatic, Statistics departments want to participate in the growth of Data Science at their respective universities.

### 2017-07-07 README.tex
Edits and re-render

@@
 Why would a statistician use a version control system, such as
-\href{http://git-scm.com}{Git} \citep{git}? And what is the point of
+\href{https://git-scm.com}{Git} \citep{git}? And what is the point of
 hosting your work online, e.g., on \href{https://github.com}{GitHub}
@@
 sharing rapidly evolving files with others. It is tempting to pick a
-quiet, private project, but you risk missing out on the main benefits of
-formal version control.
+quiet, private project. But if you do, you may never find the benefits
+of formal version control compelling enough to cement the new habit.
 
@@
 \citep{donoho}. I strongly disagree with glib claims that ``Data Science
-is just Statistics'', which I've heard in academia. For reasons both
+is just Statistics'', which can be heard in academia. For reasons both
 genuine and pragmatic, Statistics departments want to participate in the

### 2017-07-07 appendix.Rmd
Flesh out appendix

@@
   pdf_document: null
+urlcolor: blue    
 ---
@@
 
-Many of these resources live primarily on the web and, therefore, are more perishable than the content main text.
+Many of these resources live primarily on the web and, therefore, are more perishable than the content of the article.
 As time goes on, it is possible that URLs may change, and it might be necessary to use a site's navigation or search features to re-locate the resource.
@@
   * [Git](https://git-scm.com) is a version control system.
-  * [RStudio](https://www.rstudio.com/products/rstudio/) is an integrated development environment for R.
+  * [RStudio](https://www.rstudio.com/products/rstudio/) is an integrated development environment for R. It can help you use Git and GitHub. More about that below.
 
@@
   * ["FINAL.doc"](http://phdcomics.com/comics/archive.php?comicid=1531) from [PhD](http://phdcomics.com), pokes gentle fun at DIY filename-based version control.
+
+I cite it in the article, but really want to underscore the accesibility of Alice Bartlett's talk "Git for Humans".
+Highly recommended.
+[Slides on Speakerdeck](https://speakerdeck.com/alicebartlett/git-for-humans).
+
+<https://github.com/jennybc/excuse-me-iris> is the actual Git repository that lead to the screenshots in the figure.
 
@@
     - [SourceTree](https://www.sourcetreeapp.com) is free and excellent.
-    - [GitKraken](https://www.gitkraken.com) is also well-regarded and works across Mac OS, Windows, and Linux.
+    - [GitKraken](https://www.gitkraken.com) is also well-regarded. It is notable in that it runs on Linux, in addition to Mac OS and Windows.
   * [Connect to GitHub](http://happygitwithr.com/push-pull-github.html) will affirm that your system is set up properly.
@@
 
+<https://github.com/jennybc/excuse-me-iris> is the actual Git repository depicted in this figure.
+
 Comic relief
@@
   * CRAN releases are all marked with tags: <https://github.com/tidyverse/dplyr/releases>, which make it easy to see the exact state of the source code for any specific version.
+
+General Git books:
+
+  * [Git in Practice](https://www.manning.com/books/git-in-practice) by Mike McQuaid is my favorite book for day-to-day Git operations.
+  * [Pro Git](https://git-scm.com/book/en/v2) by Scott Chacon and Ben Straub is an excellent and comprehensive resource. You can read it for free online.
 

### 2017-07-07 README.Rmd
Refer to supplement in main article

@@
     RStudio and the Department of Statistics, University of British Columbia
-  thanks: |
+  thanks: >
     The author gratefully acknowledges the constructive feedback from reviewers
@@
     data science, workflow, Git, GitHub, R language, R Markdown, reproducibility
-abstract: |
+abstract: >
     Data analysis, statistical research, and teaching statistics have at least
@@
     scientific workflows. Special attention is given to projects that use the
-    statistical language R and, optionally, R Markdown documents.
+    statistical language R and, optionally, R Markdown documents. Supplementary
+    materials include an annotated set of links to step-by-step tutorials,
+    real world examples, and other useful learning resources.
 bibliography: git-github-for-stats.bib
@@
 
+# Supplementary Materials
+
+__Appendix: Tutorials, examples, resources__ Provide an annotated set of links for each section of the article, including step-by-step tutorials, real world examples, and other useful learning resources.
+
 ```{r eval = is_github_document, echo = FALSE, results = 'asis'}

### 2017-07-07 README.tex
Refer to supplement in main article

@@
 Special attention is given to projects that use the statistical language
-R and, optionally, R Markdown documents.
+R and, optionally, R Markdown documents. Supplementary materials include
+an annotated set of links to step-by-step tutorials, real world
+examples, and other useful learning resources.
 \end{abstract}
@@
 
+\section{Supplementary Materials}\label{supplementary-materials}
+
+\textbf{Appendix: Tutorials, examples, resources} Provide an annotated
+set of links for each section of the article, including step-by-step
+tutorials, real world examples, and other useful learning resources.
+
 \bibliographystyle{agsm}

### 2017-07-08 README.Rmd
URL encode the colon in the abstract

@@
     Data analysis, statistical research, and teaching statistics have at least
-    one thing in common: these activities all produce many files! There may be
-    data files, source code, figures, tables, prepared reports, and much more.
-    These files generally evolve over the course of a project and often need to
-    be shared with others, for reading or edits, as a project unfolds. Without
-    explicit and structured management, project organization can easily descend
-    into chaos, taking time away from the primary work and reducing the quality
-    of the final product. This unhappy result can be avoided by repurposing
-    tools and workflows from the software development world, namely, distributed
-    version control. This article describes the use of the version control
-    system Git and and the hosting site GitHub for statistical and data
+    one thing in common&#58; these activities all produce many files! There may
+    be data files, source code, figures, tables, prepared reports, and much
+    more. These files generally evolve over the course of a project and often
+    need to be shared with others, for reading or edits, as a project unfolds.
+    Without explicit and structured management, project organization can easily
+    descend into chaos, taking time away from the primary work and reducing the
+    quality of the final product. This unhappy result can be avoided by
+    repurposing tools and workflows from the software development world, namely,
+    distributed version control. This article describes the use of the version
+    control system Git and and the hosting site GitHub for statistical and data
     scientific workflows. Special attention is given to projects that use the

### 2017-07-08 README.Rmd
Last proofread; restore missing citation to good-enough

@@
     Data analysis, statistical research, and teaching statistics have at least
-    one thing in common&#58; these activities all produce many files! There may
-    be data files, source code, figures, tables, prepared reports, and much
-    more. These files generally evolve over the course of a project and often
+    one thing in common&#58; these activities all produce many files! There are
+    data files, source code, figures, tables, prepared reports, and much
+    more. Most of these files evolve over the course of a project and often
     need to be shared with others, for reading or edits, as a project unfolds.
@@
 each equipped with an explanatory message.
-Figure \ref{fig:diy-vs-git}C depicts this linear, *in situ* development process, with development and time flowing from bottom to top.
+Figure \ref{fig:diy-vs-git}C depicts this linear, *in situ* process, with development and time flowing from bottom to top.
 Figure \ref{fig:diy-vs-git}D shows the same history for a common collaborative Git workflow, where contributors work independently but sync regularly to a common version.
@@
 Small changes, integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you.
-The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work.
+The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work [@good-enough].
 The presence of frequently-changing binary files also increases the burden.
@@
 
-__Appendix: Tutorials, examples, resources__ Provide an annotated set of links for each section of the article, including step-by-step tutorials, real world examples, and other useful learning resources.
+__Appendix: Tutorials, examples, resources__ Provides an annotated set of links for each section of the article, including step-by-step tutorials, real world examples, and other useful learning resources.
 

### 2017-07-08 README.tex
Last proofread; restore missing citation to good-enough

--- 2017-07-07 README.tex
@@
 least one thing in common: these activities all produce many files!
-There may be data files, source code, figures, tables, prepared reports,
-and much more. These files generally evolve over the course of a project
+There are data files, source code, figures, tables, prepared reports,
+and much more. Most of these files evolve over the course of a project
 and often need to be shared with others, for reading or edits, as a
@@
 series of commits, each equipped with an explanatory message. Figure
-\ref{fig:diy-vs-git}C depicts this linear, \emph{in situ} development
-process, with development and time flowing from bottom to top. Figure
+\ref{fig:diy-vs-git}C depicts this linear, \emph{in situ} process, with
+development and time flowing from bottom to top. Figure
 \ref{fig:diy-vs-git}D shows the same history for a common collaborative
@@
 by you) is proportional to the evolutionary distance between two lines
-of work. The presence of frequently-changing binary files also increases
-the burden. So make lots of small commits, sync regularly with GitHub,
-and only track binary files with good reason.
+of work \citep{good-enough}. The presence of frequently-changing binary
+files also increases the burden. So make lots of small commits, sync
+regularly with GitHub, and only track binary files with good reason.
 
@@
 
+\section{Supplementary Materials}\label{supplementary-materials}
+
+\textbf{Appendix: Tutorials, examples, resources} Provides an annotated
+set of links for each section of the article, including step-by-step
+tutorials, real world examples, and other useful learning resources.
+
 \bibliographystyle{agsm}

### 2017-07-08 appendix.Rmd
Refer to supplement in main article
Last proofread; restore missing citation to good-enough

@@
 ---
-title: "Appendix: Tutorials and examples"
+title: "Appendix: Tutorials, examples, resources"
 output:
@@
 
-I cite it in the article, but really want to underscore the accesibility of Alice Bartlett's talk "Git for Humans".
+I cite it in the article, but really want to underscore the accessibility of Alice Bartlett's talk "Git for Humans".
 Highly recommended.

### 2017-07-08 README.Rmd
Blinded and unblinded versions

@@
 output:
+  rticles::asa_article: null
   github_document:
     toc: yes
-  rticles::asa_article: null
 keywords: |

### 2017-08-28 README.Rmd
Cite "Infrastructure and tools for teaching computing throughout the statistical curriculum"; closes #15
cc @mine-cetinkaya-rundel @rundel

@@
   * The marginal effort required to create a web presence for a project is negligible.
-  * GitHub makes a fantastic course management system for courses that use R [@R]. You can exchange actual working code with your students and explore the associated results.
+  * GitHub makes a fantastic course management system for courses that use R [@R].
+You can exchange actual working code with your students and explore the associated results [@cetinkaya-rundel-dss-2017].
   * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.
@@
 In the era of Data Science, staying current can't be limited to mathematical and methodological developments.
-The holistic work style described here is widely regarded as current best practice, yet we generally neither practice nor preach it in academic statistics.
+The holistic work style described here is widely regarded as current best practice, yet we generally neither practice nor preach it in academic statistics (but see [@cetinkaya-rundel-dss-2017] for an exception to this rule).
 It is but one example of the many ways in which Data Science is **not** just statistics.

### 2017-08-28 README.tex
Cite "Infrastructure and tools for teaching computing throughout the statistical curriculum"; closes #15
cc @mine-cetinkaya-rundel @rundel

@@
   R \citep{R}. You can exchange actual working code with your students
-  and explore the associated results.
+  and explore the associated results \citep{cetinkaya-rundel-dss-2017}.
 \item
@@
 widely regarded as current best practice, yet we generally neither
-practice nor preach it in academic statistics. It is but one example of
-the many ways in which Data Science is \textbf{not} just statistics. By
-acknowledging the importance of modern workflows and tooling, Statistics
-can solidify its claim to being one of the core disciplines of Data
-Science.
+practice nor preach it in academic statistics (but see
+\citep{cetinkaya-rundel-dss-2017} for an exception to this rule). It is
+but one example of the many ways in which Data Science is \textbf{not}
+just statistics. By acknowledging the importance of modern workflows and
+tooling, Statistics can solidify its claim to being one of the core
+disciplines of Data Science.
 

### 2017-08-28 README.Rmd
Get an email address onto the front page for PeerJ

@@
 authors:
-- name: Jennifer Bryan
+- name: Jennifer Bryan ^[jenny@rstudio.com]
   affiliation: |

### 2017-08-28 README.tex
Get an email address onto the front page for PeerJ

@@
   \author{
-        Jennifer Bryan \thanks{The author gratefully acknowledges the constructive feedback from
+        Jennifer Bryan \footnote{\href{mailto:jenny@rstudio.com}{\nolinkurl{jenny@rstudio.com}}} \thanks{The author gratefully acknowledges the constructive feedback from
 reviewers Nicholas Horton, Colin Rundel, and Hadley Wickham.} \\

### 2017-10-05 README.Rmd
Delete "yes"

@@
 @git-for-humans; @ten-simple-rules-git].
-We use it to manage the motley collection of files that make up typical data analytical projects, which consist of data, figures, reports, and, yes, source code.
+We use it to manage the motley collection of files that make up typical data analytical projects, which consist of data, figures, reports, and source code.
 Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.

### 2017-10-05 README.Rmd
Make 2 sentences into 1

@@
 A lone ranger, working on a single computer, can benefit from adopting version 
-control. But not nearly enough to justify the pain of installation and workflow 
+control, But not nearly enough to justify the pain of installation and workflow 
 upheaval. There are much easier ways to get versioned back ups of files, if 

### 2017-10-05 README.Rmd
Add missing open paren

@@
 Even when working alone, this leads to multiple versions of `iris.R` of indeterminate relatedness (Figure \ref{fig:diy-vs-git}A).
-In collaborative settings based on email distribution, the original file swiftly becomes part of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve Figure \ref{fig:diy-vs-git}B).
+In collaborative settings based on email distribution, the original file swiftly becomes part of a complicated phylogeny that no amount of "Track changes" and good intentions can resolve (Figure \ref{fig:diy-vs-git}B).
 

### 2017-10-05 README.Rmd
Remove "Yes, "
Replace a chatty word ("slick") with a more formal word ("polished")

@@
 Especially important versions get a human-readable tag, e.g., "draft-01", to signal a meaningful milestone.
-Yes, there is some pain in adopting the formalism of Git, but it is worth it.
+There is some pain in adopting the formalism of Git, but it is worth it.
 
@@
 We've introduced Git's powerful structure for file management, so where does GitHub fit in?
-GitHub complements Git by providing a slick user interface and distribution mechanism for Git repositories.
+GitHub complements Git by providing a polished user interface and distribution mechanism for Git repositories.
 Git is the software you will use locally to record changes to a set of files.

### 2017-10-05 README.Rmd
Use less chatty word: "damage" vs "goof up"
Toning this down: delete "extraordinary"

@@
 1. When you are new with Git (or, frankly, even when you're not), it's common to
-goof up the Git infrastructure for a project. Note that your files can be intact
+damage the Git infrastructure for a project. Note that your files can be intact
 and safe, even while the Git tracking is a bit confused. Of course there are 
@@
 "simulation-study", or "final-exam". Coupled with the ability to cross-link 
-issues and the project files or file changes, you have extraordinary power to 
+issues and the project files or file changes, you have the power to 
 document why things have happened in the past and to organize what needs to 

### 2017-10-05 README.Rmd
Convert a set of bullet points to a paragraph

@@
 These first steps happen once or, for some steps, once per computer.
-This is excerpted from [Happy Git and GitHub for the useR](http://happygitwithr.com), which holds battle-tested instructions honed over several years in [STAT 545](http://stat545.com) at the University of British Columbia.
+These instructions are excerpted from [Happy Git and GitHub for the useR](http://happygitwithr.com), which holds battle-tested instructions honed over several years in [STAT 545](http://stat545.com) at the University of British Columbia.
 
@@
 Once this setup is done, you are ready to start using Git and GitHub with your 
-projects. Some general recommendations for agony reduction:
-
-  * I repeat: consider using a graphical front-end for Git, a.k.a. a Git client, versus restricting yourself to the command line interface.
-  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
-  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
-  * Realize that no one is giving out Git style points. It's OK to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.
+projects.
+Here are some general recommendations for agony reduction.
+As suggested above, it may be worth using a graphical front-end for Git, a.k.a.
+a Git client, versus restricting yourself to the command line interface.
+It is fruitful to establish confidence in basic operations (e.g.
+make a change, commit it, push it) before wading into more advanced usage (e.g.
+branching).
+It is helpful to begin with a project that will provide sustained practice over several months.
+For example, usage in a course is great, because it provides a relentless stream of small deadlines.
+Finally, it is liberating to realize that no one is giving out Git style points.
+It is acceptable to "power-cycle", i.e.
+re-initialize the Git repository, to get unstuck.
 

### 2017-10-05 README.Rmd
Convert a set of bullet points to a paragraph

@@
 
-Here is the daily workflow:
-
-  * Go about your usual business, writing R scripts or authoring reports in LaTeX or R Markdown.
+The daily workflow is probably not dramatically different from what you do currently.
+You work in the usual way, writing R scripts or authoring reports in LaTeX or R Markdown.
 But instead of only *saving* individual files, periodically you make a __commit__, which takes a snapshot of all the files in the entire project.
-    - If you have ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531), you have effectively made a commit, albeit only for a single file.
+If you have ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531), you have effectively made a commit, albeit only for a single file.
 It is a version that is significant to you and that you might want to inspect or revert to later.
-  * Push commits to GitHub periodically.
-    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment.
+Periodically, you push commits to GitHub.
+This is like sharing a document with colleagues on DropBox or sending it out as an email attachment.
 By pushing to GitHub, you make your work and all your accumulated progress accessible to others.

### 2017-10-05 README.Rmd
Remove "yes"

@@
 The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features.
-Markdown is in wide use on sites like [WordPress](https://en.support.wordpress.com/markdown/), [StackOverflow](https://stackoverflow.com/editing-help), and, yes, [GitHub](https://help.github.com/categories/writing-on-github/).
+Markdown is in wide use on sites such as [WordPress](https://en.support.wordpress.com/markdown/), [StackOverflow](https://stackoverflow.com/editing-help), and [GitHub](https://help.github.com/categories/writing-on-github/).
 These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting.

### 2017-10-05 README.Rmd
"create" instead of "author"

@@
 Markdown is in wide use on sites such as [WordPress](https://en.support.wordpress.com/markdown/), [StackOverflow](https://stackoverflow.com/editing-help), and [GitHub](https://help.github.com/categories/writing-on-github/).
-These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting.
+These sites use Markdown because it allows a diverse population of site users to create decent-looking web content, with hyperlinks and some formatting.
 Do not build this up into some heroic, LaTeX-level learning task, for it is not.

### 2017-10-05 README.Rmd
Combine sections re: Markdown and Markdown for R users
Remove "holy"

@@
 
-# Markdown is special on GitHub
+# Markdown is special on GitHub and for R users
 
@@
 The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features.
-Markdown is in wide use on sites such as [WordPress](https://en.support.wordpress.com/markdown/), [StackOverflow](https://stackoverflow.com/editing-help), and [GitHub](https://help.github.com/categories/writing-on-github/).
+Markdown is in wide use on sites such as [WordPress](https://en.support.wordpress.com/markdown/), [Stack Overflow](https://stackoverflow.com/editing-help), and [GitHub](https://help.github.com/categories/writing-on-github/).
 These sites use Markdown because it allows a diverse population of site users to create decent-looking web content, with hyperlinks and some formatting.
@@
 
-# Markdown is special for R users
-
-Markdown is especially holy for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code.
+Markdown is special for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code.
 Figure \ref{fig:report-from-r-or-rmd}A shows a simple `.Rmd` document for our iris example.

### 2017-10-05 README.Rmd
Do not bold this text

@@
 These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described above.
-Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, __and view the results of running that code__, including figures.
+Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, and view the results of running that code, including figures.
 Figure \ref{fig:report-from-r-or-rmd}C shows how the `.Rmd` document from Figure \ref{fig:report-from-r-or-rmd}A looks when rendered to `.md` and pushed to GitHub.

### 2017-10-05 response-to-review.Rmd
Response to TAS review

---
title: "Response to review"
output: html_document
---

I am currently unable to get back into the Scholar One TAS system, since my link has expired. TAS editor Hadley Wickham was able to forward me a marked up PDF and I have based my edits on the annotations and comments I see there.

I made all the small copy edits that I could make sense of. I was 100% sure what some of the suggestions were, however, so a few loci remain 'as is'. I think it is possible that Preview doesn't display all the mark up as intended? I did my best.

The series of changes that I made can be seen as a sequence of commits in this branch on the repository in which I've written this article:

https://github.com/dsscollection/git-github-for-stats/commits/TAS

Below are a few specific notes about changes (or non-changes).

Added citations for the iris data.

> The flow of the paper is somewhat disrupted by repeated use of bullet points.  Is there a way to fold some of these into complete paragraphs?  (This may not be one of the ones to modify, but do search out instances throughout where you can collapse these lists to have better text flow.)

I converted every set of bullet points to a paragraph that I could. Some I have left because I feel being set off or presented as individual steps is actually conducive to understanding.

> Replace "deltas" with "differences" or "incremental changes" (etc.)

"Deltas" are actually a technical term in Git. There are such things as "delta-chains", which describe all the incremental changes necessary to transform one version of a repo to another. I use this word intentionally. I think it's best to keep it. Finally, "delta" is used in alot in math to describe differences, so I suspect this audience will understand it.

> This last part may be extraneous detail that our readers don't need to know.

This comment is about SHA-1 checksums. Again, one doesn't have to do much Googling about Git to be confronted with this vocabulary and these concepts. I feel like the goal of this article is to demystify Git and it's best to introduce the key terms.

### 2017-10-05 README.tex
Successfully creating PDFs for TAS

@@
 \usepackage{natbib}
-\usepackage{listings}
 \usepackage{textcomp}
@@
 \addtolength{\topmargin}{-.8in}%
+
+%% load any required packages here
 
@@
 manage the motley collection of files that make up typical data
-analytical projects, which consist of data, figures, reports, and, yes,
-source code. Even those who identify more as statistician than data
-scientist generally have a similar mix of files that are the artifacts
-of a project.
+analytical projects, which consist of data, figures, reports, and source
+code. Even those who identify more as statistician than data scientist
+generally have a similar mix of files that are the artifacts of a
+project.
 
 A lone ranger, working on a single computer, can benefit from adopting
-version control. But not nearly enough to justify the pain of
+version control, But not nearly enough to justify the pain of
 installation and workflow upheaval. There are much easier ways to get
@@
 of it. Figure \ref{fig:diy-vs-git} depicts a hypothetical analysis of
-the iris data, captured in a single R source file. With informal version
-control, contributors create derivative copies of \texttt{iris.R},
-decorating the file name with initials, dates, and other descriptors.
-Even when working alone, this leads to multiple versions of
-\texttt{iris.R} of indeterminate relatedness (Figure
-\ref{fig:diy-vs-git}A). In collaborative settings based on email
-distribution, the original file swiftly becomes part of a complicated
-phylogeny that no amount of ``Track changes'' and good intentions can
-resolve Figure \ref{fig:diy-vs-git}B).
+the iris data \citep[\citet{anderson}]{fisher}, captured in a single R
+source file. With informal version control, contributors create
+derivative copies of \texttt{iris.R}, decorating the file name with
+initials, dates, and other descriptors. Even when working alone, this
+leads to multiple versions of \texttt{iris.R} of indeterminate
+relatedness (Figure \ref{fig:diy-vs-git}A). In collaborative settings
+based on email distribution, the original file swiftly becomes part of a
+complicated phylogeny that no amount of ``Track changes'' and good
+intentions can resolve (Figure \ref{fig:diy-vs-git}B).
 
@@
 to a common version. Especially important versions get a human-readable
-tag, e.g., ``draft-01'', to signal a meaningful milestone. Yes, there is
-some pain in adopting the formalism of Git, but it is worth it.
+tag, e.g., ``draft-01'', to signal a meaningful milestone. There is some
+pain in adopting the formalism of Git, but it is worth it.
 
@@
 We've introduced Git's powerful structure for file management, so where
-does GitHub fit in? GitHub complements Git by providing a slick user
+does GitHub fit in? GitHub complements Git by providing a polished user
 interface and distribution mechanism for Git repositories. Git is the
@@
   When you are new with Git (or, frankly, even when you're not), it's
-  common to goof up the Git infrastructure for a project. Note that your
+  common to damage the Git infrastructure for a project. Note that your
   files can be intact and safe, even while the Git tracking is a bit
@@
 ability to cross-link issues and the project files or file changes, you
-have extraordinary power to document why things have happened in the
-past and to organize what needs to happen in the future.
+have the power to document why things have happened in the past and to
+organize what needs to happen in the future.
 
@@
 some initial setup. These first steps happen once or, for some steps,
-once per computer. This is excerpted from
+once per computer. These instructions are excerpted from
 \href{http://happygitwithr.com}{Happy Git and GitHub for the useR},
@@
 Once this setup is done, you are ready to start using Git and GitHub
-with your projects. Some general recommendations for agony reduction:
-
-\begin{itemize}
-\tightlist
-\item
-  I repeat: consider using a graphical front-end for Git, a.k.a. a Git
-  client, versus restricting yourself to the command line interface.
-\item
-  Establish confidence in the basics (e.g.~make a change, commit it,
-  push it) before wading into more advanced usage (e.g.~branching).
-\item
-  Commit yourself to Git usage on a project that will provide sustained
-  practice over several months. Usage in a course is great, because it
-  provides a relentless stream of small deadlines.
-\item
-  Realize that no one is giving out Git style points. It's OK to
-  ``power-cycle'', i.e.~re-initialize the Git repository, to get
-  unstuck.
-\end{itemize}
+with your projects. Here are some general recommendations for agony
+reduction. As suggested above, it may be worth using a graphical
+front-end for Git, a.k.a. a Git client, versus restricting yourself to
+the command line interface. It is fruitful to establish confidence in
+basic operations (e.g.~make a change, commit it, push it) before wading
+into more advanced usage (e.g.~branching). It is helpful to begin with a
+project that will provide sustained practice over several months. For
+example, usage in a course is great, because it provides a relentless
+stream of small deadlines. Finally, it is liberating to realize that no
+one is giving out Git style points. It is acceptable to ``power-cycle'',
+i.e.~re-initialize the Git repository, to get unstuck.
 
@@
 
-Here is the daily workflow:
-
-\begin{itemize}
-\tightlist
-\item
-  Go about your usual business, writing R scripts or authoring reports
-  in LaTeX or R Markdown. But instead of only \emph{saving} individual
-  files, periodically you make a \textbf{commit}, which takes a snapshot
-  of all the files in the entire project.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    If you have ever versioned a file
+The daily workflow is probably not dramatically different from what you
+do currently. You work in the usual way, writing R scripts or authoring
+reports in LaTeX or R Markdown. But instead of only \emph{saving}
+individual files, periodically you make a \textbf{commit}, which takes a
+snapshot of all the files in the entire project. If you have ever
+versioned a file
 \href{http://www.phdcomics.com/comics/archive.php?comicid=1531}{by
-    adding your initials or the date}, you have effectively made a
-    commit, albeit only for a single file. It is a version that is
-    significant to you and that you might want to inspect or revert to
-    later.
-  \end{itemize}
-\item
-  Push commits to GitHub periodically.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    This is like sharing a document with colleagues on DropBox or
-    sending it out as an email attachment. By pushing to GitHub, you
-    make your work and all your accumulated progress accessible to
-    others.
-  \end{itemize}
-\end{itemize}
+adding your initials or the date}, you have effectively made a commit,
+albeit only for a single file. It is a version that is significant to
+you and that you might want to inspect or revert to later. Periodically,
+you push commits to GitHub. This is like sharing a document with
+colleagues on DropBox or sending it out as an email attachment. By
+pushing to GitHub, you make your work and all your accumulated progress
+accessible to others.
 
@@
 
-\section{Markdown is special on
-GitHub}\label{markdown-is-special-on-github}
+\section{Markdown is special on GitHub and for R
+users}\label{markdown-is-special-on-github-and-for-r-users}
 
@@
 content, but also to prioritize human-readability, even at the cost of
-fancy features. Markdown is in wide use on sites like
+fancy features. Markdown is in wide use on sites such as
 \href{https://en.support.wordpress.com/markdown/}{WordPress},
-\href{https://stackoverflow.com/editing-help}{StackOverflow}, and, yes,
+\href{https://stackoverflow.com/editing-help}{Stack Overflow}, and
 \href{https://help.github.com/categories/writing-on-github/}{GitHub}.
 These sites use Markdown because it allows a diverse population of site
-users to author decent-looking web content, with hyperlinks and some
+users to create decent-looking web content, with hyperlinks and some
 formatting. Do not build this up into some heroic, LaTeX-level learning
@@
 
-\section{Markdown is special for R
-users}\label{markdown-is-special-for-r-users}
-
-Markdown is especially holy for R users because of
+Markdown is special for R users because of
 \href{http://rmarkdown.rstudio.com}{R Markdown}, which is just Markdown
@@
 \texttt{.Rmd} file has been rendered to \texttt{.md}, anyone viewing it
-on GitHub can read the prose, study the R code, \textbf{and view the
-results of running that code}, including figures. Figure
+on GitHub can read the prose, study the R code, and view the results of
+running that code, including figures. Figure
 \ref{fig:report-from-r-or-rmd}C shows how the \texttt{.Rmd} document
